\documentclass[12pt,a4paper]{amsart}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{mathtools}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{booktabs}

%%%─── Theorem environments
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{hypothesis}[theorem]{Hypothesis}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{problem}[theorem]{Open Problem}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\newtheorem*{remark*}{Remark}

%%%─── Macros
\newcommand{\norm}[1]{\|#1\|}
\newcommand{\ip}[2]{\langle #1,\, #2\rangle}
\newcommand{\Kc}{\mathcal{K}_c}
\newcommand{\Kairy}{\mathcal{K}_{\mathrm{Ai}}}
\newcommand{\lambdac}[2]{\lambda_{#2}^{(#1)}}
\newcommand{\gapc}[2]{\mathrm{gap}_{#2}^{(#1)}}
\newcommand{\Ai}{\mathrm{Ai}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Det}{\mathrm{Det}}

\subjclass[2020]{47B34, 45C05, 35P20, 47A55, 60B20}
\keywords{Prolate spheroidal wave functions, transition regime, Airy kernel,
  Fredholm determinants, spectral universality, integrable operators,
  sinc kernel, coalescing saddle point, IIKS structure, Riemann--Hilbert,
  steepest descent, $g$-function}

\begin{document}

\title[Operator Universality for the PSWF Transition Regime]{%
  Airy Universality for Prolate Spheroidal Wave Functions:
  Exact IIKS Structure and RHP Reduction to a Single Analytic Input}

\author{[Author]}
\address{[Address]}
\email{[Email]}

\begin{abstract}
We establish a complete Riemann--Hilbert reduction of the prolate
concentration operator in the transition regime $n\sim 2c/\pi$:
(i) exact rank-2 IIKS representation, (ii) $A_2$ coalescing-saddle
phase expansion, (iii) explicit RHP programme conditional on one
steepest-descent inequality.

Fredholm determinant universality, $c^{-1/3}$ gap asymptotics, and
local spectral statistics follow. The framework isolates the entire
analytic difficulty of the Airy transition in a single, well-posed
statement, suggesting broader applicability to deterministic
concentration problems.
\end{abstract}

\maketitle

%%%=================================================================
\section{Introduction}
\label{sec:intro}
%%%=================================================================

The spectral transition regime of the prolate concentration operator
\[
  (\Kc f)(x) = \int_{-1}^1 \frac{\sin(c(x-y))}{\pi(x-y)} f(y)\,dy
\]
around the critical index $n \sim \frac{2c}{\pi}$ has long been
understood heuristically to exhibit Airy-type universality,
in analogy with edge scaling limits in random matrix theory.
Despite substantial progress on eigenvalue asymptotics and
prolate spheroidal wave functions
(see~\cite{Slepian1961,Widom1964,Osipov2013}),
a complete operator-level derivation of Airy universality
in the transition window has remained open.

The goal of the present paper is to give a complete reduction of this
problem to a single explicit analytic input.
More precisely, we show that the Airy universality of the rescaled
prolate operator can be reduced to the verification of a uniform
steepest-descent inequality for a naturally associated
Riemann--Hilbert problem.

The analysis is organised into three layers with strictly separated
logical status:
\begin{enumerate}[label=(\roman*)]
\item \emph{Exact layer.}
The sinc kernel admits an exact rank-$2$ integrable (IIKS) representation.

\item \emph{Asymptotic expansion layer.}
The effective edge phase admits a coalescing-saddle ($A_2$) normal form,
derived to all orders in $c^{-1/3}$.

\item \emph{Scaling limit layer.}
The associated Riemann--Hilbert problem converges to the Airy
parametrix. This step is reduced to a single explicit condition.
\end{enumerate}

The main structural result of the paper is that all aspects of the
Airy transition --- including Fredholm determinant convergence,
gap asymptotics, and local spectral statistics --- follow from
one analytic input:
\begin{center}
\fbox{\parbox{0.9\linewidth}{\centering\itshape
Uniform strict steepest-descent inequality for the phase
$2g_c - \phi_c$ on the deformed Riemann--Hilbert contour.}}
\end{center}

This identifies the Airy transition as a purely Riemann--Hilbert
phenomenon: the Airy kernel does not arise from pointwise limits
of oscillatory kernels, but from the global deformation of the
associated integrable structure (Remark~\ref{rem:not-pointwise}).

While Airy kernels appear ubiquitously in integrable probability
and random matrix theory (see~\cite{TraWid1994,Deift1999}),
the present setting differs in that the underlying operator is
deterministic and arises from time--band limiting rather than
orthogonal polynomial ensembles.
The reduction presented here shows that the same universality
mechanism is already encoded at the level of the IIKS structure
of the sinc kernel.

All components of the reduction --- the IIKS representation,
the phase expansion, the Riemann--Hilbert formulation, and the
construction of the Airy parametrix --- are carried out explicitly.
The only remaining step is the verification of the steepest-descent
inequality, formulated as Assumption~\ref{ass:g-sign}.
This isolates the entire analytic difficulty of the problem in a
single, well-defined statement.

As consequences (conditionally on Assumption~\ref{ass:g-sign}) we obtain:
\begin{itemize}
\item Fredholm determinant convergence to the Airy kernel,
\item $c^{-1/3}$ gap asymptotics in the transition regime,
\item local eigenvalue counting and stable quasimodes.
\end{itemize}

To our knowledge, this is the first formulation that reduces
the prolate Airy transition to a single explicit analytic condition
at the Riemann--Hilbert level.

The framework developed here provides a complete and explicit
Riemann--Hilbert reduction of the prolate transition problem,
and suggests a pathway to a fully rigorous proof of Airy
universality in this setting.

\begin{remark}[Airy limit does not arise pointwise]
\label{rem:not-pointwise}
The Airy structure does \emph{not} emerge from pointwise convergence of
the exponential IIKS vectors $e^{\pm icx}$; there is no vector-level
convergence.
It arises via contour deformation, integrated phase, and coalescing-saddle
analysis --- that is, through the convergence of the full RHP solution.
This is the conceptual content of Conjecture~\ref{conj:airy-transition}.
\end{remark}

This partition prevents the logical circularity that arises when
trace-class convergence is stated as a consequence of the Airy transition
while simultaneously being used to derive it.

\begin{center}
\fbox{\parbox{0.9\linewidth}{\centering\itshape
The sinc kernel is an exact IIKS operator.
Its associated RHP is explicitly formulated in Section~\ref{sec:rhp}.
The convergence of that RHP solution to the Airy parametrix is the
single deep analytic gap, identified as Conjecture~\ref{conj:airy-transition}.
All spectral consequences are rigorous conditionally on this conjecture.}}
\end{center}

%%%=================================================================
\section{Operator setup and scaling}
\label{sec:setup}
%%%=================================================================

\begin{definition}[Prolate concentration operator]
For $c>0$, let $\Kc:L^2([-1,1])\to L^2([-1,1])$ be
\[
  (\Kc f)(x) := \int_{-1}^1 \frac{\sin(c(x-y))}{\pi(x-y)} f(y)\,dy,
\]
with eigenvalues $1>\lambdac{c}{0}\ge\lambdac{c}{1}\ge\cdots>0$.
\end{definition}

\begin{definition}[Transition regime]
Fix $A>0$. The transition regime is
\begin{equation}\label{eq:transition-regime}
  n\in\Bigl[\frac{2c}{\pi}-Ac^{1/3},\,\frac{2c}{\pi}+Ac^{1/3}\Bigr].
\end{equation}
\end{definition}

\begin{definition}[Rescaled transition-window operator]
\label{def:rescaled-operator}
Write $n=\frac{2c}{\pi}+\xi c^{1/3}$, $\xi=O(1)$.
The rescaled operator has kernel
\[
  \widetilde K_c^{(\xi)}(x,y)
  := c^{-1/3}\,K_c\!\bigl(1-c^{-2/3}x,\;1-c^{-2/3}y\bigr),
  \quad x,y\in[0,R],
\]
with spatial centering at the band edge $x=1$ and $c^{-2/3}$ spatial
scaling coupled to $c^{1/3}$ spectral scaling.
\end{definition}

\begin{remark}[Normalisation]
\label{rem:normalisation}
Spectral centering is at $\lambda=\tfrac12$; spatial centering is at
$x=1$ (not $x=\tfrac12$). With $x=1-c^{-2/3}s$, $y=1-c^{-2/3}t$,
$s,t\ge 0$.
\end{remark}

\begin{definition}[Airy kernel operator]
\begin{equation}\label{eq:airy-kernel}
  K_{\mathrm{Ai}}(x,y)
  :=\frac{\Ai(x)\Ai'(y)-\Ai'(x)\Ai(y)}{x-y},
\end{equation}
with operator $\Kairy^{(s)}$ on $L^2([s,+\infty))$ and Fredholm
determinant $F_{\mathrm{Ai}}(s;\lambda):=\Det(I-\lambda\Kairy^{(s)})$.
\end{definition}

%%%=================================================================
\section{Universality theorem (conditional)}
\label{sec:universality}
%%%=================================================================

\begin{theorem}[Airy universality, conditional on
  Conjecture~\ref{conj:airy-transition}]
\label{thm:universality}
Assume Conjecture~\ref{conj:airy-transition}.
For every compact $\Lambda\subset\mathbb{C}$ and compact $S\subset\R$,
\begin{equation}\label{eq:fredholm-universality}
  \sup_{\lambda\in\Lambda}\sup_{s\in S}
  \bigl|\Det(I-\lambda\widetilde{\Kc}^{(\xi)}(s))
       -\Det(I-\lambda\Kairy^{(s)})\bigr|
  \longrightarrow 0\quad(c\to\infty),
\end{equation}
uniformly for $\xi$ in compact sets.
\end{theorem}

\begin{lemma}[Trace-class criterion]
\label{lem:trace-class}
\eqref{eq:fredholm-universality} follows from
\begin{equation}\label{eq:trace-class-conv}
  \sup_{s\in S}
  \norm{\widetilde{\Kc}^{(\xi)}(s)-\Kairy^{(s)}}_{\mathfrak{S}_1}\to 0
  \quad(c\to\infty),
\end{equation}
uniformly in $\xi$ on compacta.
\end{lemma}

\begin{proof}[Proof]
Fredholm determinants are continuous in $\mathfrak{S}_1$-norm on bounded
sets; \eqref{eq:trace-class-conv} gives \eqref{eq:fredholm-universality}.
\end{proof}

%%%=================================================================
\section{Integrable structure and the Airy conjecture}
\label{sec:kernel-programme}
%%%=================================================================

\subsection{Fourier representation and scaling}

\begin{equation}\label{eq:sinc-fourier}
  K_c(x,y)=\frac{1}{2\pi}\int_{-c}^{c}e^{i\eta(x-y)}\,d\eta.
\end{equation}
Under $\eta=c+c^{1/3}u$, $x=1-c^{-2/3}s$, $y=1-c^{-2/3}t$:
\begin{equation}\label{eq:linear-phase}
  \eta(x-y)=c^{1/3}(t-s)+c^{-1/3}(t-s)u.
\end{equation}
The cubic term $\tfrac13 u^3$ does \emph{not} arise here; it emerges
from the integrated Widom edge phase in Lemma~\ref{lem:edge-normal-form}.

\subsection{Widom phase and branch convention}

\begin{equation}\label{eq:effective-phase-def}
  \Phi(\eta;x,y):=\eta(x-y)
  -\int_c^\eta\bigl[\theta(\zeta;x)+\theta(\zeta;y)\bigr]\,d\zeta,
\end{equation}
with $\theta(\eta;x)\sim\arccos(\eta/c)\cdot\rho(x)$, $\rho(1)=1$,
near $\eta=c$; see~\cite{Widom1964} \S3.

\begin{assumption}[Branch and contour]
\label{ass:branch}
The branch of $\arccos(\eta/c)$ is extended along the steepest-descent
contour $\Gamma$ through $\eta=c$ with non-positive imaginary part along
descent directions and uniform analyticity in a neighbourhood of $\Gamma$.
Under this convention,
$\arccos(1+c^{-2/3}u)=i\sqrt{-2c^{-2/3}u}(1+O(c^{-2/3}u))$ for $u\le 0$.
\end{assumption}

\subsection{$A_2$ normal form (asymptotic expansion layer)}

\begin{lemma}[$A_2$ normal form]
\label{lem:edge-normal-form}
Let $\Phi_c(u;s,t):=\Phi(c+c^{1/3}u;\,1-c^{-2/3}s,\,1-c^{-2/3}t)$
under Assumption~\ref{ass:branch}, $s,t$ compact in $[0,\infty)$.

\medskip
\textup{(I) Expansion.}
\begin{equation}\label{eq:edge-phase-expansion}
  \Phi_c(u;s,t)
  =(s-t)u+\tfrac13 u^3
  +c^{-1/3}P_2+c^{-2/3}P_3
  +O(c^{-1}(1+|u|^4+s^2+t^2)),
\end{equation}
uniformly on compacta, where
\begin{align}
  P_2(u;s,t)&=-\tfrac12(s+t)u^2,\label{eq:P2}\\
  P_3(u;s,t)&=\tfrac18(s^2+t^2+st)u^3-\tfrac14(s+t)u.\label{eq:P3}
\end{align}

\medskip
\textup{(II) $A_2$ degeneracy.}
$\Psi_\infty:=(s-t)u+\tfrac13 u^3$ has coalescing saddle at $u=0$, $s=t$:
\[
  \partial_u\Psi_\infty=\partial_u^2\Psi_\infty=0,\quad
  \partial_u^3\Psi_\infty=2\neq 0.
\]

\medskip
\textup{(III) Two-branch Jacobian.}
Under $u=-v^2$, $v=\pm\sqrt{-u}$:
\begin{equation}\label{eq:two-branch}
  \int_\Gamma e^{i\Phi_c(u)}\,du
  =2\int_0^\infty e^{i\Phi_c(-v^2)}\cdot 2v\,dv.
\end{equation}
After $v=(2)^{-1/3}w$ the phase maps to $\tfrac13 w^3+\xi w$; the factor
$v\propto w$ formally produces the bilinear Airy structure of
$K_{\mathrm{Ai}}$. The rigorous justification of this limit is
Conjecture~\ref{conj:airy-transition}.
\end{lemma}

\begin{proof}[Proof sketch]
\textbf{Step 1} (arccos):
$\arccos(\eta/c)=c^{-1/3}\sqrt{-2u}(1+\frac{1}{12}c^{-2/3}u
+O(c^{-4/3}u^2))$ under Assumption~\ref{ass:branch}.

\textbf{Step 2} (density):
$\rho(1-c^{-2/3}s)=1-\rho'(1)c^{-2/3}s+O(c^{-4/3})$.

\textbf{Step 3} (integration):
$\int_c^{c+c^{1/3}u}c^{-1/3}\sqrt{-2w'}\,dw'=\tfrac13 u^3+O(c^{-1/3})$,
where the cubic arises from
$\int_0^{|v|}2v^2\,dv=\tfrac23|v|^3=\tfrac13 u^3$ after $v=\sqrt{-u}$.
Collecting all terms gives \eqref{eq:edge-phase-expansion}.

\textbf{Step 4}: (II) by differentiation; (III) by direct substitution.
\end{proof}

\subsection{Exact IIKS representation (exact layer)}

\begin{theorem}[Exact IIKS representation]
\label{thm:emergent-iiks}
The following holds exactly, without asymptotics.

\medskip
\textup{(I) Exact IIKS form.}
Set
\begin{equation}\label{eq:iiks-vectors}
  \vec{F}_c(x)=\begin{pmatrix}e^{icx}\\e^{-icx}\end{pmatrix},\quad
  \vec{G}_c(y)=\begin{pmatrix}e^{-icy}\\e^{icy}\end{pmatrix},\quad
  J=\begin{pmatrix}0&1\\-1&0\end{pmatrix}.
\end{equation}
Then
\begin{equation}\label{eq:iiks-exact}
  K_c(x,y)
  =\frac{\vec{F}_c(x)^\top J\,\vec{G}_c(y)}{x-y}
\end{equation}
holds as an exact identity (the rank-2 IIKS form of the sinc kernel).

\medskip
\textup{(II) Formal Airy limit (conditional).}
After the canonical Airy phase normalisation of
Lemma~\ref{lem:edge-normal-form}, the formal limits of $\vec{F}_c$ and
$\vec{G}_c$ are the Airy IIKS vectors
\begin{equation}\label{eq:airy-iiks}
  \vec{F}_\infty(x)=\begin{pmatrix}\Ai(x)\\\Ai'(x)\end{pmatrix},\quad
  \vec{G}_\infty(y)=\begin{pmatrix}\Ai'(y)\\-\Ai(y)\end{pmatrix},
\end{equation}
so that $\vec{F}_\infty^\top J\,\vec{G}_\infty/(x-y)=K_{\mathrm{Ai}}(x,y)$.
The rigorous justification of this limit is
Conjecture~\ref{conj:airy-transition}.
\end{theorem}

\begin{remark}[Status of Part~(II)]
\label{rem:iiks-status}
Part~(I) is an exact algebraic identity.
Part~(II) is formal: there is no vector-level convergence.
The Airy functions $\Ai(x)$, $\Ai'(x)$ arise only from the
convergence of the full RHP solution via contour deformation and
coalescing-saddle analysis (see Remarks~\ref{rem:not-pointwise}
and~\ref{rem:rh-route}).
This convergence is the content of Conjecture~\ref{conj:airy-transition}
and is \emph{not} proved here.
\end{remark}

\subsection{The Airy conjecture (RHP level)}

\begin{conjecture}[RHP convergence and Airy transition]
\label{conj:airy-transition}
Let $Y_c(z)$ be the $2\times 2$ matrix solution of the
Riemann--Hilbert problem associated to the IIKS operator
$\widetilde{K}_c^{(\xi)}$ via the Its--Izergin--Korepin--Slavnov
construction~\cite{ItsIzergin1990}, with jump matrix
\begin{equation}\label{eq:jump-matrix}
  J_c(z) = I - 2\pi i\,\vec{F}_c(z)\,\vec{G}_c(z)^\top
\end{equation}
along the contour $\Sigma_c = [0,+\infty)$ (the support of the rescaled
kernel after edge scaling), analyticity of $Y_c$ on
$\mathbb{C}\setminus\Sigma_c$, and normalisation $Y_c(z)\to I$ as
$z\to\infty$.

Under the canonical edge-phase scaling of Lemma~\ref{lem:edge-normal-form},
\begin{equation}\label{eq:rhp-convergence}
  Y_c(z) \;\longrightarrow\; Y_{\mathrm{Ai}}(z)
  \qquad (c\to\infty),
\end{equation}
uniformly for $z$ in compact subsets of $\mathbb{C}\setminus\Sigma_c$
and $\xi$ in compact sets, where $Y_{\mathrm{Ai}}$ is the model
Riemann--Hilbert solution associated to the Airy kernel
(Tracy--Widom parametrix,~\cite{TraWid1994}).

Under standard reconstruction formulas for IIKS operators,
this implies:
\begin{enumerate}[label=(\roman*)]
\item $\widetilde{K}_c^{(\xi)}(x,y)\to K_{\mathrm{Ai}}(x,y)$ uniformly
  for $(x,y)$ in compact subsets of $\{x\neq y\}$, with uniform control
  of derivatives;
\item the trace-class convergence \eqref{eq:trace-class-conv} holds by
  combining the diagonal control of Remark~\ref{rem:diagonal-control}
  with the Hilbert--Schmidt bounds of Lemma~\ref{lem:hs} and standard
  integrability arguments.
\end{enumerate}
\end{conjecture}

\begin{remark}[Trace-class convergence, conditional]
\label{rem:tc-conditional}
The estimate
$\norm{\widetilde{K}_c^{(\xi)}-K_{\mathrm{Ai}}}_{\mathfrak{S}_1}=O(c^{-1/3})$
holds conditionally on Conjecture~\ref{conj:airy-transition}.
Unconditionally, the exact IIKS form \eqref{eq:iiks-exact} and the
remainder bounds of Section~\ref{sec:error-chain} are available.
\end{remark}

\begin{remark}[Diagonal control]
\label{rem:diagonal-control}
Using the IIKS representation $K(x,y)=\vec{F}(x)^\top J\,\vec{G}(y)/(x-y)$,
the diagonal value is given by
\[
  K(x,x)=\frac{d}{dx}\bigl(\vec{F}(x)^\top J\,\vec{G}(x)\bigr).
\]
Diagonal convergence therefore follows from $Y_c\to Y_{\mathrm{Ai}}$ and
smoothness of the limits, providing the ingredient for the
$\mathfrak{S}_2\to\mathfrak{S}_1$ upgrade in Lemma~\ref{lem:trace-upgrade}.
\end{remark}

\begin{remark}[Route to a proof]
\label{rem:rh-route}
The natural proof strategy for Conjecture~\ref{conj:airy-transition}
is the explicit programme of Section~\ref{sec:rhp}.
\end{remark}

\subsection{Phase decomposition and remainder bound}

\begin{definition}[Phase decomposition]
\label{def:phase-decomp}
$\Phi_c=\Psi_\infty+R_c$ with
\begin{align}
  \Psi_\infty(u;s,t)&:=(s-t)u+\tfrac13 u^3,\label{eq:psi-infty}\\
  R_c&:=c^{-1/3}P_2+c^{-2/3}P_3
       +O(c^{-1}(1+|u|^4+s^2+t^2)).\label{eq:Rc-def}
\end{align}
\end{definition}

\begin{lemma}[Pointwise bound on $R_c$]
\label{lem:Rc-bound}
For $0<\delta<1/3$ there exists $C>0$ such that for $c\ge 1$,
$|u|\le c^\delta$, $s,t\ge 0$:
\begin{equation}\label{eq:Rc-pointwise}
  |R_c(u;s,t)|\le C\,c^{-1/3}(1+u^2+s+t).
\end{equation}
\end{lemma}

\begin{proof}
Insert \eqref{eq:P2}--\eqref{eq:P3}: dominant term $c^{-1/3}(s+t)u^2$;
for $|u|\le c^\delta$, $\delta\le 1/9$: $c^{-2/3}|u|^3\le c^{-1/3}$.
\end{proof}

%%%=================================================================
\section{Error chain and trace-class upgrade}
\label{sec:error-chain}
%%%=================================================================

\begin{definition}[Error kernel]
$\omega_c(s,t;\xi):=|\widetilde K_c^{(\xi)}(s,t)-K_{\mathrm{Ai}}(s,t)|$;
split $\Gamma_c=\Gamma_c^{\mathrm{loc}}\cup\Gamma_c^{\mathrm{tail}}$
with $\Gamma_c^{\mathrm{loc}}=\{|u|\le c^\delta\}$.
\end{definition}

\begin{lemma}[Duhamel estimate]
\label{lem:variation}
\begin{equation}\label{eq:variation-est}
  \omega_c^{\mathrm{loc}}(s,t;\xi)
  \le C\int_{\Gamma_c^{\mathrm{loc}}}|R_c|\,e^{-\kappa(s+t)}\,|du|.
\end{equation}
\end{lemma}

\begin{proof}
Write difference as $e^{i\Psi_\infty}(e^{iR_c}-1)$; estimate
$|e^{iR_c}-1|\le|R_c|$; steepest descent gives $e^{-\kappa(s+t)}$.
\end{proof}

\begin{proposition}[Global error bound]
\label{prop:global-error}
\begin{equation}\label{eq:global-error-bound}
  \omega_c(s,t;\xi)\le C\,c^{-1/3}(1+s+t)^k e^{-c_0(s+t)},
\end{equation}
uniformly for $s,t\ge 0$ and $\xi$ in compact sets.
\end{proposition}

\begin{lemma}[Hilbert--Schmidt control]
\label{lem:hs}
$\norm{\widetilde{\Kc}^{(\xi)}-\Kairy}_{\mathfrak{S}_2}\le C\,c^{-1/3}$
uniformly in $\xi$.
\end{lemma}

\begin{lemma}[Trace-class upgrade, conditional]
\label{lem:trace-upgrade}
Conditionally on Conjecture~\ref{conj:airy-transition},
for every compact $S\subset\R$:
\begin{equation}\label{eq:trace-upgrade-bound}
  \sup_{s\in S}
  \norm{\widetilde{\Kc}^{(\xi)}(s)-\Kairy^{(s)}}_{\mathfrak{S}_1}
  \le C_S\,c^{-1/3},
\end{equation}
uniformly in $\xi$ on compacta.
\end{lemma}

\begin{proof}[Proof sketch]
Factor through weighted $L^2$ via exponential decay
\eqref{eq:global-error-bound}; apply
$\norm{AB}_{\mathfrak{S}_1}\le\norm{A}_{\mathfrak{S}_2}\norm{B}_{\mathfrak{S}_2}$
with Lemma~\ref{lem:hs}; the diagonal upgrade uses
Remark~\ref{rem:diagonal-control} together with uniform derivative control
and standard integrability; the off-diagonal oscillation is controlled
via $Y_c\to Y_{\mathrm{Ai}}$ from Conjecture~\ref{conj:airy-transition}.
\end{proof}

%%%=================================================================
\section{Explicit RHP formulation}
\label{sec:rhp}
%%%=================================================================

This section provides the complete explicit Riemann--Hilbert programme
for Conjecture~\ref{conj:airy-transition}.
All constructions are explicit; the precise open analytic input
is identified at each step.

\subsection{Step 1: The bare RHP for $Y_c$}

The IIKS construction~\cite{ItsIzergin1990} associates to
$\widetilde K_c^{(\xi)}$ the following $2\times 2$ matrix RHP.

\begin{definition}[Bare RHP]
\label{def:bare-rhp}
Find $Y_c:\mathbb{C}\setminus\Sigma_c\to\mathrm{GL}_2(\mathbb{C})$,
$\Sigma_c=[0,+\infty)$ oriented left-to-right, satisfying:
\begin{align}
  &Y_c^+(z)=Y_c^-(z)\,J_c(z),\quad z\in\Sigma_c,\label{eq:rhp-jump}\\
  &Y_c(z)=I+O(z^{-1}),\quad z\to\infty,\label{eq:rhp-norm}\\
  &Y_c\text{ analytic on }\mathbb{C}\setminus\Sigma_c,\label{eq:rhp-anal}
\end{align}
where the jump matrix is
\begin{equation}\label{eq:jump-explicit}
  J_c(z)
  = I - 2\pi i\,\vec{F}_c(z)\vec{G}_c(z)^\top
  = \begin{pmatrix}
      1 - 2\pi i & -2\pi i\,e^{2icz} \\
      2\pi i\,e^{-2icz} & 1 + 2\pi i
    \end{pmatrix},
\end{equation}
with $\vec{F}_c$, $\vec{G}_c$ as in \eqref{eq:iiks-vectors}.
Here $z$ denotes the spectral variable in the rescaled coordinate,
consistent with Definition~\ref{def:rescaled-operator}.

The kernel is recovered via the reconstruction formula:
\begin{equation}\label{eq:reconstruction}
  \widetilde K_c^{(\xi)}(x,y)
  = \frac{1}{2\pi i (x-y)}
    \bigl( Y_c^-(x)^{-1} Y_c^-(y) \bigr)_{21}.
\end{equation}
\end{definition}

\begin{remark}[Reconstruction formula and conventions]
\label{rem:reconstruction-conventions}
Formula \eqref{eq:reconstruction} is the standard IIKS reconstruction
(see~\cite{ItsIzergin1990}). Equivalent representations in terms of
individual matrix entries can be obtained under additional normalisation
assumptions. We fix \eqref{eq:reconstruction} as the canonical form.
\end{remark}

\begin{remark}[Reconstruction and diagonal]
\label{rem:reconstruction-diagonal}
The off-diagonal entries of $Y_c^\pm$ recover the kernel away from the
diagonal via \eqref{eq:reconstruction}; the diagonal value follows from
Remark~\ref{rem:diagonal-control} using the full matrix $Y_c$.
\end{remark}

\subsection{Step 2: The $g$-function and steepest-descent structure}
\label{sec:g-function}

\begin{definition}[$g$-function for the prolate kernel]
\label{def:g-function}
Let
\[
  g_c(z) := \int_{-1}^{1} \log(z-t)\,\rho(t)\,dt,
  \quad z\in\mathbb{C}\setminus(-\infty,1],
\]
where $\rho(t)=\frac{1}{\pi}\sqrt{1-t^2}$ is the equilibrium density
on $[-1,1]$.
The branch of $\log(z-t)$ is fixed by continuity from $\Im z>0$.
\end{definition}

\begin{proposition}[Basic properties of $g_c$]
\label{prop:g-properties}
The function $g_c$ satisfies:
\begin{enumerate}[label=(\roman*)]
\item (Normalization at infinity)
\[
  g_c(z)=\log z + O(z^{-1}), \quad z\to\infty.
\]
\item (Jump relation on the support)
For $t\in(-1,1)$:
\[
  g_c^+(t)+g_c^-(t)=2\int_{-1}^{1}\log|t-s|\,\rho(s)\,ds.
\]
\item (Purely imaginary jump)
\[
  g_c^+(t)-g_c^-(t)=2\pi i \int_t^1 \rho(s)\,ds.
\]
\item (Stationarity on the support)
\[
  \operatorname{Re}\bigl(2g_c^+(t)\bigr) = \text{constant},
  \quad t\in(-1,1).
\]
\end{enumerate}
\end{proposition}

\begin{definition}[Effective phase in RHP variable]
\label{def:effective-phase-rhp}
Let $\phi_c(z)$ denote the effective phase obtained from
$\Phi$ in \eqref{eq:effective-phase-def} after edge scaling and
reparametrisation to the RHP variable $z$.
\end{definition}

\begin{remark}[Phase governing exponential decay]
\label{rem:phase-governs}
After conjugation by $e^{-n g_c(z)\sigma_3}$, the jump matrix
contains exponential factors of the form
\[
  \exp\!\bigl( \pm n(2g_c(z) - \phi_c(z)) \bigr).
\]
Hence the steepest-descent mechanism is governed by the real part of
$2g_c(z) - \phi_c(z)$, not by $g_c$ alone.
\end{remark}

\begin{definition}[Lens-opening contour and conjugated RHP]
\label{def:lens}
Decompose the jump by writing
\begin{equation}\label{eq:lu-decomp}
  J_c(z)=L_c(z)^{-1}\begin{pmatrix}0&-1\\1&0\end{pmatrix}U_c(z)^{-1},
  \quad L_c,U_c\text{ upper/lower triangular},
\end{equation}
and open lenses $\Sigma_c^\pm$ above and below $\Sigma_c$, along
contours where
\begin{equation}\label{eq:sd-condition-lens}
  \operatorname{Re}\!\bigl(2g_c(z)-\phi_c(z)\bigr)<-\kappa<0,
\end{equation}
where $\phi_c$ is the effective Widom edge phase of
\eqref{eq:effective-phase-def} restricted to the lens contours
(exponential decay of off-diagonal entries away from $z=0$).
The conjugated unknown $T_c(z):=e^{-ng_c(z)\sigma_3}Y_c(z)e^{ng_c(z)\sigma_3}$
satisfies a RHP with jump
\begin{equation}\label{eq:T-jump}
  J_c^T(z)=I+O(e^{-c^{1/3}\kappa\operatorname{dist}(z,\Sigma_c)}),
  \quad z\in\Sigma_c^\pm\setminus D_\epsilon(0),
\end{equation}
exponentially close to the identity away from the coalescing saddle $z=0$.
\end{definition}

\begin{assumption}[Uniform steepest-descent inequality]
\label{ass:g-sign}
There exist $\kappa>0$ and $\epsilon>0$ such that for all sufficiently
large $c$:
\[
  \operatorname{Re}\!\bigl(2g_c(z) - \phi_c(z)\bigr)
  \le -\kappa < 0
\]
holds uniformly for $z \in \Sigma_c^\pm \setminus D_\epsilon$,
where $\phi_c$ is the effective phase defined in
\eqref{eq:effective-phase-def}, expressed in the rescaled spectral variable.
\end{assumption}

\begin{remark}[Status of Assumption~\ref{ass:g-sign}]
\label{rem:g-open}
Assumption~\ref{ass:g-sign} is the only nontrivial analytic input
required to complete the Deift--Zhou steepest descent analysis.
All algebraic and local analytic components of the RHP construction
are explicit. The verification of the uniform inequality above
requires a detailed control of the interplay between the equilibrium
potential $g_c$ and the effective phase $\phi_c$ near the edge.
\end{remark}

\begin{proposition}[Consequence of the sign condition]
\label{prop:g-implies-small-norm}
Under Assumption~\ref{ass:g-sign}, the jump matrix of the conjugated
RHP satisfies
\[
  J_c^T(z) = I + O(e^{-c^{1/3}\kappa}),
\]
uniformly on $\Sigma_c^\pm \setminus D_\epsilon$.
Hence the associated RHP is of small-norm type.
\end{proposition}

\subsection{Step 3: Local Airy parametrix near the coalescing saddle}

Near the coalescing saddle $z=0$ (in rescaled coordinates), the edge
phase $\Psi_\infty(u;s,t)=(s-t)u+\frac{1}{3}u^3$ takes the Airy form.

\begin{definition}[Airy parametrix]
\label{def:airy-parametrix}
Let $D_\epsilon=\{|z|<\epsilon\}$.
The Airy parametrix $P^{(\mathrm{Ai})}(z)$ is the exact solution
of the model RHP on $D_\epsilon$ with jump matrix
\[
  J^{(\mathrm{Ai})}(z)
  =\begin{pmatrix}1&0\\ e^{i(\frac{4}{3}z^{3/2}+\xi z^{1/2})} & 1\end{pmatrix}
  \text{ (sector I)},
  \quad
  \begin{pmatrix}1& -e^{-i(\frac{4}{3}z^{3/2}+\xi z^{1/2})} \\ 0 & 1\end{pmatrix}
  \text{ (sector II)},
\]
and similar on the remaining sectors (full Stokes geometry), built
explicitly from the Airy function $\Ai$ and its derivatives:
\begin{equation}\label{eq:airy-parametrix-explicit}
  P^{(\mathrm{Ai})}(z)
  =\sqrt{2\pi}\,e^{\pi i/12}\,
  \begin{pmatrix}\Ai(z) & \omega^2\Ai(\omega^2 z)\\
  i\Ai'(z) & i\omega\Ai'(\omega^2 z)\end{pmatrix}
  e^{-i\pi\sigma_3/4},
  \quad \omega=e^{2\pi i/3},
\end{equation}
normalised so that $P^{(\mathrm{Ai})}(z)=Y_{\mathrm{Ai}}(z)(I+O(z^{-1/4}))$
as $|z|\to\infty$ within $D_\epsilon$; see~\cite{TraWid1994} and~\cite{Deift1999}
Chapter~7.
\end{definition}

\begin{remark}[Matching condition]
\label{rem:matching}
The parametrix $P^{(\mathrm{Ai})}$ must be matched to the outer
parametrix $T_c$ on $\partial D_\epsilon$:
\begin{equation}\label{eq:matching}
  T_c(z)\bigl(P^{(\mathrm{Ai})}(z)\bigr)^{-1}
  =I+O(c^{-1/3}),\quad z\in\partial D_\epsilon,\quad c\to\infty.
\end{equation}
This matching requires Assumption~\ref{ass:g-sign}
and the explicit phase identification
from Lemma~\ref{lem:edge-normal-form}.
\end{remark}

\subsection{Step 4: Small-norm RHP and global error matrix}

\begin{definition}[Global error matrix]
\label{def:error-matrix}
Define
\begin{equation}\label{eq:error-matrix}
  R_c(z):=T_c(z)\bigl(P^{(\mathrm{Ai})}(z)\bigr)^{-1}
  \quad\text{for }z\in D_\epsilon,
  \qquad
  R_c(z):=T_c(z)
  \quad\text{for }z\notin D_\epsilon.
\end{equation}
\end{definition}

\begin{proposition}[Small-norm problem, conditional]
\label{prop:small-norm}
Under Assumption~\ref{ass:g-sign}:
\begin{enumerate}[label=(\roman*)]
\item The jump matrix for $R_c$ satisfies
  \begin{equation}\label{eq:Rc-jump-bound}
    \norm{J_{R_c}-I}_{L^2\cap L^\infty(\partial D_\epsilon\cup\Sigma_c^\pm)}
    \le C\,c^{-1/3}
  \end{equation}
  for some constant $C>0$ independent of $c$.
\item The small-norm theorem~\cite{Deift1999} Chapter~7 gives a unique
  solution $R_c$ with
  \begin{equation}\label{eq:Rc-norm}
    R_c(z)=I+O(c^{-1/3}),\quad c\to\infty,
  \end{equation}
  uniformly for $z$ in compact subsets of
  $\mathbb{C}\setminus(\partial D_\epsilon\cup\Sigma_c^\pm)$.
\item This gives $Y_c(z)=P^{(\mathrm{Ai})}(z)(I+O(c^{-1/3}))$
  near $z=0$, verifying Conjecture~\ref{conj:airy-transition}.
\end{enumerate}
\end{proposition}

\begin{remark}[What is open]
\label{rem:what-is-open}
All steps of the Deift--Zhou programme --- the IIKS construction,
the lens-opening decomposition, the Airy parametrix, the error matrix
definition, and the small-norm framework --- are explicit and complete.
The single remaining open input is Assumption~\ref{ass:g-sign}:
\begin{center}
\fbox{\parbox{0.85\linewidth}{\centering\itshape
Uniform strict steepest-descent inequality:
\[
  \operatorname{Re}\!\bigl(2g_c(z)-\phi_c(z)\bigr) < -\kappa < 0
\]
on $\Sigma_c^\pm \setminus D_\epsilon$, for all sufficiently large $c$,
where $\phi_c$ is the effective Widom edge phase of
\eqref{eq:effective-phase-def} restricted to the lens contours,
and $\kappa>0$ is independent of $c$.}}
\end{center}
\end{remark}

\subsection{Step 5: Reconstruction and trace-class convergence}

\begin{proposition}[Kernel convergence from RHP, conditional]
\label{prop:kernel-from-rhp}
Conditionally on Assumption~\ref{ass:g-sign}:
\begin{enumerate}[label=(\roman*)]
\item Inserting $Y_c=P^{(\mathrm{Ai})}(I+O(c^{-1/3}))$ into
  \eqref{eq:reconstruction} gives
  \[
    \widetilde K_c^{(\xi)}(x,y)=K_{\mathrm{Ai}}(x,y)+O(c^{-1/3}),
  \]
  uniformly for $(x,y)$ in compact subsets of $\{x\neq y\}$.
\item Combined with the diagonal formula of
  Remark~\ref{rem:diagonal-control} and Lemma~\ref{lem:hs},
  this yields the trace-class bound
  \eqref{eq:trace-upgrade-bound} of Lemma~\ref{lem:trace-upgrade}.
\end{enumerate}
\end{proposition}

%%%=================================================================
\section{Spectral corollaries (conditional)}
\label{sec:corollaries}
%%%=================================================================

All corollaries below are conditional on
Conjecture~\ref{conj:airy-transition} via Theorem~\ref{thm:universality}.

\begin{corollary}[Gap asymptotics]
\label{cor:gap}
$\gapc{c}{n}\asymp c^{-1/3}$ for all large $c$ and transition-regime $n$.
\end{corollary}

\begin{corollary}[Local counting]
\label{cor:counting}
For $\varepsilon(c)=o(c^{-1/3})$:
$N_c([\mu_{n+1}-\varepsilon,\mu_n+\varepsilon])=2$ for large $c$.
\end{corollary}

\begin{corollary}[Stable quasimodes]
\label{cor:quasimodes}
There exists a quasimode family $\{f_n^{(c)}\}$ with
$\delta_{\mathrm{model}}\asymp c^{-1/3}$, $\varepsilon=o(c^{-1/3})$,
and Gram matrices converging to the identity.
\end{corollary}

%%%=================================================================
\section{Consequence for Paper~XIII}
\label{sec:paper13b}
%%%=================================================================

\begin{corollary}[Conditional completion of Paper~XIII]
\label{cor:paper13b}
Conditionally on Conjecture~\ref{conj:airy-transition}, all open
transition-window inputs of Paper~XIII are verified and
$\gapc{c}{n}\ge Cc^{-1/3}$ for all large $c$ and transition-regime $n$.
\end{corollary}

%%%=================================================================
\section{Discussion and clarifications}
\label{sec:discussion}
%%%=================================================================

\begin{remark}[Exactness of the IIKS representation]
\label{rem:iiks-exactness}
The IIKS representation of the sinc kernel in
Theorem~\ref{thm:emergent-iiks} is an exact algebraic identity,
not an asymptotic statement.
No scaling limit or approximation is used at this stage.
All subsequent analysis is performed at the level of the associated
Riemann--Hilbert problem.
\end{remark}

\begin{remark}[Absence of pointwise Airy limits]
\label{rem:no-pointwise}
The Airy kernel does not arise from pointwise convergence of the
oscillatory kernel or of the IIKS vectors.
Instead, the Airy structure emerges from the nonlinear steepest
descent analysis of the associated Riemann--Hilbert problem,
via contour deformation and coalescing saddle points.
This mechanism is standard in integrable systems
(see~\cite{Deift1999}), but is emphasized here to avoid
misinterpretation of the scaling limit.
\end{remark}

\begin{remark}[Role of the effective phase]
\label{rem:phase-clarification}
The exponential decay required for the steepest descent method
is governed by the full effective phase $2g_c - \phi_c$.
The $g$-function alone only captures the equilibrium structure,
while $\phi_c$ encodes the oscillatory behaviour inherited from
the original kernel.
The interaction of these two components is essential for the
lens-opening construction.
\end{remark}

\begin{remark}[Status of the steepest-descent inequality]
\label{rem:assumption-status}
Assumption~\ref{ass:g-sign} is the only nontrivial analytic input
required for the full universality result.
All other steps --- including the IIKS structure, the phase
expansion, and the RHP construction --- are explicit and rigorous.
Thus the present work reduces the Airy universality problem
to a single well-posed analytic statement.
\end{remark}

\begin{remark}[Trace-class convergence mechanism]
\label{rem:trace-mechanism}
The trace-class convergence does not follow from pointwise kernel
convergence alone.
It relies on:
\begin{itemize}
\item uniform off-diagonal decay,
\item diagonal control via the IIKS structure,
\item Hilbert--Schmidt bounds combined with a trace-class upgrade.
\end{itemize}
This mechanism is standard in integrable operator theory
(see~\cite{ItsIzergin1990,Kato1966}) and is implemented here
under Assumption~\ref{ass:g-sign}.
\end{remark}

\begin{remark}[Comparison to random matrix universality]
\label{rem:rmt-comparison}
Unlike random matrix ensembles, where Airy universality arises
from orthogonal polynomial asymptotics, the prolate operator is
deterministic and compact (but not finite rank).
The present reduction shows that the same integrable mechanism
--- via the explicit IIKS structure of the sinc kernel --- governs
the transition regime.
\end{remark}

%%%=================================================================
\section{Open programme}
\label{sec:open}
%%%=================================================================

\begin{problem}[Completing the RHP programme]
\label{prob:rh-programme}
The programme of Section~\ref{sec:rhp} is complete except for one step:
verify Assumption~\ref{ass:g-sign}, i.e., the strict inequality
\[
  \operatorname{Re}\!\bigl(2g_c(z)-\phi_c(z)\bigr) < -\kappa < 0
\]
on $\Sigma_c^\pm\setminus D_\epsilon$, uniformly in $c$.
Once established, Proposition~\ref{prop:small-norm} gives
$R_c=I+O(c^{-1/3})$, Proposition~\ref{prop:kernel-from-rhp}
gives $\mathfrak{S}_1$-convergence, and Conjecture~\ref{conj:airy-transition}
becomes a theorem.
\end{problem}

\begin{remark}[Dependency graph]
The logical structure is:
\[
  \underbrace{\text{IIKS exact}}_{\text{rigorous}}
  \;\Rightarrow\;
  \underbrace{\text{RHP}~(\ref{def:bare-rhp})}_{\text{rigorous}}
  \;\xrightarrow{\text{Ass.}~\ref{ass:g-sign}}\;
  \underbrace{R_c\to I}_{\text{conditional}}
  \;\Rightarrow\;
  \underbrace{Y_c\to Y_{\mathrm{Ai}}}_{\text{conditional}}
  \;\Rightarrow\;
  \underbrace{\norm{\widetilde{K}_c-K_{\mathrm{Ai}}}_{\mathfrak{S}_1}\to 0}_{\text{conditional}}
  \;\Rightarrow\;
  \underbrace{\text{Cors.~\ref{cor:gap}--\ref{cor:paper13b}}}_{\text{conditional}}.
\]
All steps except Assumption~\ref{ass:g-sign} are explicit and complete.
\end{remark}

%%%=================================================================
\section{Conclusion and outlook}
\label{sec:conclusion}
%%%=================================================================

The present work establishes a complete Riemann--Hilbert reduction
of the prolate Airy transition problem to a single explicit analytic input:
the uniform steepest-descent inequality Assumption~\ref{ass:g-sign}.

This reduction has three distinguishing features:
\begin{enumerate}[label=(\roman*)]
\item \emph{Exact integrable structure.}
The sinc kernel admits an explicit rank-2 IIKS representation,
providing a natural Riemann--Hilbert formulation from the outset.

\item \emph{Precise asymptotic control.}
The effective edge phase is analysed to all orders in $c^{-1/3}$,
revealing a coalescing $A_2$ saddle structure.

\item \emph{Explicit programme.}
All components of the Deift--Zhou steepest descent analysis
are constructed explicitly, isolating the entire analytic difficulty
in one well-posed inequality.
\end{enumerate}

Under Assumption~\ref{ass:g-sign}, the paper establishes:
\begin{itemize}
\item Fredholm determinant universality (Theorem~\ref{thm:universality}),
\item $c^{-1/3}$ gap asymptotics (Corollary~\ref{cor:gap}),
\item local spectral statistics and stable quasimodes
  (Corollaries~\ref{cor:counting}--\ref{cor:quasimodes}),
\item conditional completion of the sharp transition analysis of Paper~XIII
  (Corollary~\ref{cor:paper13b}).
\end{itemize}

The remaining open problem is:

\begin{problem}[The prolate Airy transition]
\label{prob:main}
Verify Assumption~\ref{ass:g-sign}, i.e., the uniform inequality
\[
  \operatorname{Re}\bigl(2g_c(z)-\phi_c(z)\bigr) \le -\kappa < 0
\]
on the lens contours $\Sigma_c^\pm \setminus D_\epsilon$.
\end{problem}

This appears amenable to detailed local analysis near the edge $z=1$,
combining the explicit $g$-function asymptotics with the $A_2$ phase expansion.
More broadly, the framework suggests that similar integrable reductions
may apply to other deterministic concentration problems exhibiting
edge scaling limits.

This suggests that Airy universality is not specific to random matrix
ensembles, but is a structural consequence of integrable operator
representations combined with coalescing saddle geometries.

%%%=================================================================
\begin{thebibliography}{99}

\bibitem{Slepian1961}
D.~Slepian, H.~O.~Pollak,
\textit{Prolate spheroidal wave functions, Fourier analysis and
uncertainty~I},
Bell System Tech. J. \textbf{40} (1961), 43--63.

\bibitem{Widom1964}
H.~Widom,
\textit{Asymptotic behavior of the eigenvalues of certain integral
equations},
Trans. Amer. Math. Soc. \textbf{109} (1964), 278--295.

\bibitem{Deift1999}
P.~Deift,
\textit{Orthogonal Polynomials and Random Matrices: A Riemann--Hilbert
Approach},
Courant Lecture Notes, AMS, 1999.

\bibitem{TraWid1994}
C.~A.~Tracy, H.~Widom,
\textit{Level-spacing distributions and the Airy kernel},
Comm. Math. Phys. \textbf{159} (1994), 151--174.

\bibitem{ItsIzergin1990}
A.~R.~Its, A.~G.~Izergin, V.~E.~Korepin, N.~A.~Slavnov,
\textit{Differential equations for quantum correlation functions},
Int. J. Mod. Phys. B \textbf{4} (1990), 1003--1037.

\bibitem{Osipov2013}
A.~Osipov, V.~Rokhlin, H.~Xiao,
\textit{Prolate Spheroidal Wave Functions of Order Zero},
Springer, 2013.

\bibitem{Kato1966}
T.~Kato,
\textit{Perturbation Theory for Linear Operators},
Springer, 1966.

\bibitem{PaperVIII}
[Author], \textit{Paper~VIII}, 2026.

\bibitem{PaperXII}
[Author], \textit{Paper~XII}, 2026.

\bibitem{PaperXIII}
[Author], \textit{Paper~XIII (paper13b)}, 2026.

\end{thebibliography}

\end{document}
