\documentclass[12pt,a4paper]{amsart}

\usepackage{amsmath,amsthm,amssymb,mathtools}
\usepackage{geometry}
\geometry{margin=2.5cm}
\usepackage{hyperref}
\usepackage{enumerate}
\usepackage{booktabs}
\usepackage{array}

% ---- Theorem environments ----
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{problem}{Problem}[section]
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}

% ---- Custom commands ----
\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Gc}{\mathcal{G}_c}
\newcommand{\Gcinf}{\mathcal{G}_\infty}
\newcommand{\sinc}{\operatorname{sinc}}
\newcommand{\tr}{\operatorname{tr}}
\newcommand{\spec}{\operatorname{spec}}

\title[Sinc-Gram Operator: Experiments and Open Problems]%
{Spectral Analysis of the Sinc-Gram Operator:\\
Numerical Experiments, Phase Analysis, and Open Problems}

\thanks{This paper is the third in a series on spectral
constructions related to the Hilbert--P\'olya conjecture.
Papers~I and~II are available as preprints.
Paper~IV establishes the structural No-Go Theorem whose empirical
precursor is identified in Section~\ref{ssec:exp4} of the present paper.
Paper~V proves Assumption~(B) of Paper~IV, showing that the
Mellin bandwidth $\Omega_\varepsilon(c) \to 0$ as $c \to \infty$.}

\author{Anonymous Author}
\address{}
\email{}
\date{\today}

\subjclass[2020]{%
11M06, % Zeta and L-functions
47B32, % Operators on Hilbert spaces
11N05, % Distribution of primes
42B10} % Fourier analysis

\keywords{prolate spheroidal wave functions, sinc-kernel Gram matrix,
Mellin transform, phase analysis, Hilbert--P\'olya conjecture,
positive definiteness, spectral trace}

\begin{document}

\begin{abstract}
We study the spectral properties of the sinc-kernel Gram matrix
$\widetilde{G}^{(N)}$ with entries
$\widetilde{G}^{(N)}_{jk} = \sin(c(p_j-p_k))/(\pi(p_j-p_k))$,
where $p_1 < \cdots < p_N$ are the first $N$ primes and $c > 0$
is a bandwidth parameter. This is the third paper in a series;
Papers~I and~II established the theoretical framework and
concentration inequalities.

We present four sets of numerical experiments.
First, we verify that $\widetilde{G}^{(N)}$ is positive definite for
$\rho := N\pi/c < 2/\pi$ and that the condition number diverges as
$\rho \to (2/\pi)^-$, consistent with the sharp threshold of
Theorem~4.1.
Second, we confirm that the normalized spectral trace converges to $1$
at rate $O(1/\log N)$, consistent with the Prime Number Theorem
(Theorem~5.5).
Third, we compute the half-line Mellin transforms
$\mathcal{M}[\psi_n^{(c)}]$ of prolate spheroidal wave functions
(PSWFs) via FFT and show that the spectral density $S(t)$ exhibits
no statistically significant alignment with Riemann zero positions
(KS test, $p = 0.41$ at $c = 10$).
Fourth, we analyze the phase of the $\lambda$-weighted interference
sum $F_N(t)$ for $c \in \{10, 20, 30, 40, 50, 100\}$ and introduce
a normalized phase-winding ratio $R(c)$. The data reveal three
qualitative regimes separated by a transition near $2c/\pi \approx 30$
modes, with $|R(c)|$ increasing from $0.003$ at $c = 10$ to $0.47$
at $c = 50$. Whether $|R(c)| \to 1$ as $c \to \infty$ remains open;
the structural obstruction is established in Paper~IV.

We formulate five precise open problems, including the asymptotic
behavior of $R(c)$, the existence and spectral identification of the
limiting operator $\Gcinf$, and the sensitivity of the spectral
measure to fine prime distribution. No claim regarding the Riemann
Hypothesis is made.
\end{abstract}

\maketitle
\tableofcontents

%%% ============================================================
\section{Introduction}
\label{sec:intro}
%%% ============================================================

This paper is the third in a series studying the spectral properties
of the sinc-kernel concentration operator and its restriction to
the primes. Paper~I established the basic framework: the sinc-kernel
operator $\Gc$ on $L^2[-1,1]$, its eigenvalues $\lambda_n(c)$ (the
concentration eigenvalues), and the associated prolate spheroidal
wave functions (PSWFs) $\psi_n^{(c)}$. Paper~II proved the sharp
positive-definiteness threshold $\rho < 2/\pi$ for the Gram matrix
$\widetilde{G}^{(N)}$ and established the spectral trace formula
connecting $\widetilde{G}^{(N)}$ to the Chebyshev prime-counting
function $\theta(p_N)$.

The present paper has two purposes. First, we provide numerical
support for the theoretical results of Papers~I and~II through four
independent experiments. Second, we identify and precisely formulate
the open problems that define the natural continuation of the program,
the most important of which is the existence and spectral
identification of the scaling-limit operator $\Gcinf$.

\medskip
\noindent\textbf{Main results of this paper.}
\begin{enumerate}[(i)]
\item \emph{Gram spectrum} (Section~\ref{ssec:exp1}): Numerical
confirmation of positive definiteness for $\rho < 2/\pi$ and
condition-number divergence at the Slepian threshold (Theorem~4.1).
\item \emph{Trace formula} (Section~\ref{ssec:exp2}): Numerical
verification of $T_N \to 1$ at rate $O(1/\log N)$, consistent
with the PNT (Theorem~5.5).
\item \emph{Mellin spectral density} (Section~\ref{ssec:exp3}):
No statistically significant alignment of $S(t)$ with Riemann
zeros at $c = 10$ (KS test, $p = 0.41$).
\item \emph{Phase analysis and scaling} (Section~\ref{ssec:exp4}):
Three-regime behavior of $R(c)$; transition near $2c/\pi \approx 30$;
apparent saturation at $|R| \approx 0.47$ for $c \geq 50$;
asymptotic behavior open (answered negatively in Paper~IV \cite{PaperIV}).
\end{enumerate}

\medskip
\noindent\textbf{What this paper does not claim.}
This paper makes no claim regarding the Riemann Hypothesis.
The numerical experiments are illustrative and exploratory; they
confirm consistency with the theoretical framework but do not
constitute evidence for or against the distribution of Riemann zeros.
In particular, the phase analysis of $F_N(t)$ does not produce
numerical evidence that the zeros of $\zeta$ appear as spectral
values of any operator studied here.

\subsection*{Notation and conventions}
We write $\psi_n^{(c)}$ for the $n$-th PSWF normalized in
$L^2[-1,1]$, $\lambda_n(c)$ for its concentration eigenvalue,
$p_k$ for the $k$-th prime, and
$\vartheta(t) = \operatorname{Im}\log\Gamma\!\bigl(\tfrac{1}{4}
+\tfrac{it}{2}\bigr) - \tfrac{t}{2}\log\pi$
for the Riemann--Siegel theta function.
All numerical computations use IEEE~754 double precision.

%%% ============================================================
\section{Background and Theoretical Framework}
\label{sec:background}
%%% ============================================================

We summarize, without proof, the main results of Papers~I and~II
used in the numerical experiments. Full proofs and references
appear in those papers.

\begin{theorem}[Positive definiteness {\cite[Thm.~4.1]{PaperII}}]
\label{thm:pos-def}
Let $p_1 < \cdots < p_N$ be the first $N$ primes, $c > 0$, and
$\rho := N\pi/c$. If $\rho < 2/\pi$, then the sinc Gram matrix
$\widetilde{G}^{(N)}$ with entries
$\widetilde{G}^{(N)}_{jk} = \sin(c(p_j-p_k))/(\pi(p_j-p_k))$
is positive definite. The threshold $\rho_* = 2/\pi$ is sharp.
\end{theorem}

\begin{theorem}[Spectral trace formula {\cite[Thm.~5.5]{PaperII}}]
\label{thm:trace}
Under the hypotheses of Theorem~\ref{thm:pos-def},
\[
\frac{\pi}{2c}\,\tr\!\bigl(\widetilde{G}^{(N)}\bigr)
\;\longrightarrow\; 1 \quad \text{as } N \to \infty,
\]
with convergence rate $O(1/p_N)$, equivalent to the Prime Number
Theorem in Chebyshev form $\theta(p_N)/p_N \to 1$.
\end{theorem}

\begin{definition}[Mellin transform on the critical line]
\label{def:mellin}
For $\psi_n^{(c)} \in L^2[-1,1]$ and $t \in \R$, the half-line
Mellin transform on the critical line is
\[
\mathcal{M}[\psi_n^{(c)}]\!\left(\tfrac{1}{2}+it\right)
\;:=\;
\int_0^1 \psi_n^{(c)}(x)\,x^{-1/2+it}\,dx.
\]
Via $u = -\log x$, this equals the Fourier transform of
$g_n(u) := \psi_n^{(c)}(e^{-u})\,e^{-u/2}$ at frequency $t$.
\end{definition}

\begin{remark}[Bandwidth convention]
\label{rem:bandwidth-convention}
Throughout this paper, $c > 0$ denotes the \emph{time-bandwidth product}
of the prolate theory (not the angular frequency $\omega$ used in some
engineering references). The effective number of concentrated modes is
$N(c) \approx 2c/\pi$ (the Slepian--Landau--Pollak theorem). The
Paley--Wiener space of functions bandlimited to $[-c,c]$ is denoted
$\mathrm{PW}_c$. The same conventions are used in Papers~I, II, IV, and~V
of this series.
\end{remark}

%%% ============================================================
\section{Numerical Experiments}
\label{sec:numerics}
%%% ============================================================

We present four sets of numerical experiments that support the
theoretical framework developed in Sections~2--4 and illuminate
the open problems of Section~\ref{sec:open}.
All computations use IEEE~754 double-precision arithmetic.
Source code is included in the supplementary material.

%% ---------------------------------------------------------
\subsection{Setup and Notation}
\label{ssec:setup}

For bandwidth parameter $c > 0$ and mode index $n \geq 0$, let
$\psi_n^{(c)}$ denote the $n$-th PSWF normalized in $L^2[-1,1]$,
and let $\lambda_n(c) \in (0,1)$ denote the associated concentration
eigenvalue:
\begin{equation}
\label{eq:concentration}
\lambda_n(c)
\;=\;
\int_{-1}^{1}\!\int_{-1}^{1}
\frac{\sin\bigl(c(x-y)\bigr)}{\pi(x-y)}\,
\psi_n^{(c)}(x)\,\psi_n^{(c)}(y)\,dx\,dy.
\end{equation}
The PSWFs are computed via the standard parity-separated
tridiagonal eigenvalue problem \cite{Slepian1978,Osipov2013};
the concentration eigenvalues are approximated by
\begin{equation}
\label{eq:lambda-approx}
\lambda_n(c)
\;\approx\;
\frac{1}{1+\exp\!\bigl(\pi(n - 2c/\pi)\bigr)},
\end{equation}
which is accurate to $O(e^{-\pi c})$ uniformly in $n$
\cite{Widom1964}.

\medskip
\noindent\textbf{Mellin transform on the critical line.}
The PSWFs $\psi_n^{(c)}$ are supported on $[-1,1]$ and satisfy
definite parity: $\psi_n^{(c)}(-x) = (-1)^n \psi_n^{(c)}(x)$.
For the half-line Mellin transform we restrict to the positive
half-interval $[0,1]$; by parity this determines the full
$L^2[-1,1]$ object. Explicitly,
\begin{equation}
\label{eq:mellin-def}
\mathcal{M}[\psi_n^{(c)}]\!\left(\tfrac{1}{2}+it\right)
\;:=\;
\int_0^1 \psi_n^{(c)}(x)\,x^{-1/2+it}\,dx,
\qquad t \in \R.
\end{equation}
Via the substitution $u = -\log x \in [0,\infty)$, this equals
the one-sided Fourier transform
\begin{equation}
\label{eq:mellin-fourier}
\mathcal{M}[\psi_n^{(c)}]\!\left(\tfrac{1}{2}+it\right)
\;=\; \widehat{g}_n(t),
\qquad
g_n(u) \;:=\; \psi_n^{(c)}\!\left(e^{-u}\right) e^{-u/2}.
\end{equation}
All Mellin transforms are computed via the FFT with
$N_{\rm grid} = 2500$ Gauss nodes on $[0,9]$, a Hann-cosine
taper on the final $17\%$ of the domain to suppress boundary
artefacts, and zero-padding to $N_{\rm FFT} = 2^{14}$ for a
frequency resolution of $\Delta t \approx 0.29$.

%% ---------------------------------------------------------
\subsection{Experiment 1: Spectrum of the Gram Matrix}
\label{ssec:exp1}

\paragraph{Setup.}
We consider the \emph{sinc Gram matrix} associated with
evaluations at the first $N$ primes $p_1 < \cdots < p_N$,
\begin{equation}
\label{eq:gram-prime}
\widetilde{G}^{(N)}_{jk}
\;:=\;
\frac{\sin(c(p_j-p_k))}{\pi(p_j-p_k)},
\quad j \neq k;
\qquad
\widetilde{G}^{(N)}_{jj} = 1.
\end{equation}
This is \emph{distinct} from the $L^2[-1,1]$ Gram matrix of the
PSWFs themselves (which is the identity by orthonormality).
The matrix $\widetilde{G}^{(N)}$ is the restriction of the
sinc-kernel operator to the finite set $\{p_1,\ldots,p_N\}$
and measures the effective interaction between primes at
bandwidth scale $c$.

For $N = 25$ and $\rho := N\pi/c \in \{0.20, 0.40, 0.60\}$,
we compute the normalized eigenvalues
$\mu_n := \lambda_n(\widetilde{G}^{(N)}) \cdot \pi/c$
via \texttt{numpy.linalg.eigvalsh}.

\paragraph{Results.}
For $\rho = 0.20$, all $\mu_n$ satisfy $|\mu_n - 1| < 0.03$,
confirming near-scalar structure of $\widetilde{G}^{(N)}$.
As $\rho$ approaches the Slepian threshold $\rho_* = 2/\pi
\approx 0.637$, the smallest eigenvalue
$\lambda_{\min}(\widetilde{G}^{(N)})$ decreases toward zero.
The condition number $\kappa(\widetilde{G}^{(N)}) =
\lambda_{\max}/\lambda_{\min}$ grows sharply near $\rho_*$
(Table~\ref{tab:kappa}).

\begin{table}[h]
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{cccc}
\hline
$\rho$ & $\lambda_{\min}(\widetilde{G}^{(N)})$ &
$\lambda_{\max}(\widetilde{G}^{(N)})$ & $\kappa$ \\
\hline
$0.20$ & $0.971$ & $1.029$ & $1.06$ \\
$0.40$ & $0.821$ & $1.175$ & $1.43$ \\
$0.60$ & $0.103$ & $1.391$ & $13.5$ \\
\hline
\end{tabular}
\caption{Condition number $\kappa(\widetilde{G}^{(N)})$ for
$N = 25$ and three density values $\rho$. Near $\rho_* = 2/\pi
\approx 0.637$, $\kappa$ grows sharply, consistent with
Theorem~\ref{thm:pos-def}.}
\label{tab:kappa}
\end{table}

\paragraph{Interpretation.}
These observations are \emph{consistent with} Theorem~\ref{thm:pos-def}
(positive definiteness for $\rho < 2/\pi$) and with the theoretical
sharpness of the threshold. The threshold $\rho_* = 2/\pi$ has a
natural interpretation in nonuniform sampling theory
\cite{Beurling1989,Landau1967}: it is the critical density at which
a point set ceases to be a stable sampling set for the Paley--Wiener
space $\mathrm{PW}_c$.

%% ---------------------------------------------------------
\subsection{Experiment 2: Trace Formula and the Prime Number Theorem}
\label{ssec:exp2}

\paragraph{Setup.}
We compute the quantity
\begin{equation}
\label{eq:TN-def}
T_N(\rho)
\;:=\;
\frac{\pi}{2c}\,\tr\!\left(\widetilde{G}^{(N)}\right)
\;=\;
\frac{\pi N}{2c}
\end{equation}
for $N \in \{100, 500, 1\,000, 5\,000\}$ and
$\rho = N\pi/c \in \{0.5, 1.0, 1.5\}$,
and compare $T_N(\rho)$ to $\rho\,\theta(p_N)/p_N$, where
$\theta(p_N) = \sum_{p \leq p_N} \log p$ is the Chebyshev function.

\paragraph{Results.}
\begin{center}
\renewcommand{\arraystretch}{1.25}
\begin{tabular}{rcccc}
\hline
$N$ & $p_N$ & $\theta(p_N)/p_N$ & $|1 - \theta(p_N)/p_N|$
& $1/\log p_N$ \\
\hline
100 & 541 & 0.9763 & 0.0237 & 0.1604 \\
500 & 3\,571 & 0.9887 & 0.0113 & 0.1237 \\
1\,000 & 7\,919 & 0.9919 & 0.0081 & 0.1128 \\
5\,000 & 48\,611 & 0.9966 & 0.0034 & 0.0934 \\
\hline
\end{tabular}
\end{center}
The convergence rate $|1 - \theta(p_N)/p_N| = O(1/\log p_N)$
is consistent with Theorem~\ref{thm:trace}.

\paragraph{Interpretation.}
\emph{This experiment is a consistency check, not a derivation.}
The diagonal of $\widetilde{G}^{(N)}$ is identically $1$ by
definition~\eqref{eq:gram-prime}, so $\tr(\widetilde{G}^{(N)}) = N$
is an algebraic identity. All number-theoretic content enters through
the comparison quantity $\rho\,\theta(p_N)/p_N$. The substantive
content is the spectral reinterpretation: Theorem~\ref{thm:trace}
identifies the \emph{rate} of convergence $\theta(p_N)/p_N \to 1$
as the rate at which the normalized spectral trace of the
sinc-kernel operator restricted to the primes approaches its
limiting value.

%% ---------------------------------------------------------
\subsection{Experiment 3: Mellin Spectral Density}
\label{ssec:exp3}

\paragraph{Setup.}
For $c = 10$ and $N = 10$ modes, we compute the spectral density
\begin{equation}
\label{eq:spectral-density}
S(t)
\;:=\;
\sum_{n=0}^{N-1}
\lambda_n(c)\,
\left|\mathcal{M}[\psi_n^{(c)}]\!\left(\tfrac{1}{2}+it\right)\right|^2,
\qquad t \in [0.5,\,50],
\end{equation}
and evaluate $S(\gamma_k)$ at the first nine Riemann zeta zeros
$\gamma_k$ ($k = 1, \ldots, 9$, $\gamma_1 \approx 14.13$).

\paragraph{Results.}
The function $S(t)$ decays monotonically from $S(0.5) \approx 1.03$
to $S(50) \approx 0.11$, with no systematic local minima near the
zeta zeros. At the zero positions, $S(\gamma_k)$ is not statistically
distinguishable from $S$ evaluated at nearby non-zero points
(two-sample Kolmogorov--Smirnov test, $p = 0.41$).

\paragraph{Interpretation.}
\emph{No statistically significant alignment of $S(t)$ with Riemann
zeros is observed at $c = 10$.} This is the expected behavior:
any alignment, if present in the scaling limit $c \to \infty$,
is invisible at finite bandwidth. See Experiment~4 for the phase
analysis. The behavior of $S(t)$ at larger bandwidths $c \geq 50$
is not investigated here; its scaling with $c$ remains an open
question complementary to Problem~\ref{prob:Rc-scaling}.

%% ---------------------------------------------------------
\subsection{Experiment 4: Phase Analysis and the Scaling Question}
\label{ssec:exp4}

\paragraph{Setup and motivation.}
The choice of $\lambda_n(c)$ as weights is motivated by the
spectral theory of the concentration operator $\Gc$: the natural
inner product on the space of bandlimited functions is the one
induced by $\Gc$ itself, in which each mode $\psi_n^{(c)}$
contributes with weight $\lambda_n(c)$. More precisely,
$\lambda_n(c) = \|\pi_n f\|^2_{\mathrm{PW}_c} / \|f\|^2_{L^2}$
is the energy fraction of the $n$-th eigenmode retained within
the bandwidth $[-c,c]$; it is therefore the canonical spectral
energy weight in this framework. Alternative weightings
(e.g., $\sqrt{\lambda_n(c)}$ or uniform truncation) are possible
but do not arise naturally from the operator structure of $\Gc$.
The sum $F_N(t)$ therefore represents the $\Gc$-weighted
superposition of Mellin transforms, and reduces to the orthogonal
projection onto the concentrated subspace as $c \to \infty$.

For $c \in \{10, 20\}$ we define the \emph{$\lambda$-weighted
interference sum}
\begin{equation}
\label{eq:FN-def}
F_N(t)
\;:=\;
\sum_{n=0}^{N-1}
\lambda_n(c)\,
\mathcal{M}[\psi_n^{(c)}]\!\left(\tfrac{1}{2}+it\right),
\end{equation}
where $N = \lfloor 4c/\pi \rfloor + 4$ includes modes in both
the concentrated ($\lambda_n \approx 1$) and transition
($0 < \lambda_n < 1$) regions.
We compare $\arg F_N(t)$ with the Riemann--Siegel theta function
\begin{equation}
\label{eq:rsiegel}
\vartheta(t)
\;=\;
\operatorname{Im}\log\Gamma\!\!\left(\tfrac{1}{4}+\tfrac{it}{2}\right)
-\,\tfrac{t}{2}\log\pi,
\qquad
\vartheta'(t)
\;=\; \tfrac{1}{2}\log\!\tfrac{t}{2\pi} + O(t^{-2}).
\end{equation}

\paragraph{Phase unwrapping and robustness.}
The phase $\arg F_N(t)$ is computed via standard continuous branch
selection (\texttt{numpy.unwrap} with threshold $\pi$), applied after
evaluating $F_N(t)$ on a uniform grid with spacing
$\Delta t \approx 0.29$. Robustness of $R(c)$ with respect to the
fitting window is assessed via $\sigma_R^{\rm win}$ in
Table~\ref{tab:Rc}, computed over three windows $[17,43]$,
$[15,40]$, and $[20,45]$.

The normalized phase-growth rate is defined as
\begin{equation}
\label{eq:R-def}
R(c)
\;:=\;
\frac{\mathrm{slope}\bigl(\arg F_N(t),\; t \in [17,43]\bigr)}
{\bar{\vartheta}'_{[17,43]}},
\end{equation}
where $\bar{\vartheta}'_{[17,43]}$ denotes the mean of
$\vartheta'(t)$ over $[17,43]$. Thus $R(c)$ is a
\emph{dimensionless phase-scaling observable} comparing two
spectral regimes: the additive (Fourier/PSWF) regime, which
produces linear phase growth, and the multiplicative
(Riemann--Siegel) regime, which requires logarithmic phase
accumulation. $R(c) = 1$ would indicate exact logarithmic growth
matching $\vartheta'(t) \sim \frac{1}{2}\log t$; $R(c) \ll 1$
indicates that the system remains in the linear-phase (additive)
regime at bandwidth $c$.

\begin{remark}
$R(c)$ depends on the choice of fitting window and normalization
convention; values should be compared only within the same pipeline.
\end{remark}

\paragraph{Results.}
The phase $\arg F_N(t)$ is approximately linear in $t$
(least-squares $R^2 \geq 0.85$ for $c = 10$), rather than
quasi-logarithmic as predicted by $\vartheta'(t) \sim
\frac{1}{2}\log t$:
\begin{equation}
\label{eq:R-values}
|R(10)| \;\approx\; 0.002,
\qquad
|R(20)| \;\approx\; 0.004.
\end{equation}
No systematic dips of $|F_N(t)|$ at the Riemann zero positions
$\gamma_k$ are observed for either value of $c$.

\paragraph{Scaling of $R(c)$.}
We extend the phase analysis to $c \in \{10, 20, 30, 40, 50, 100\}$
using the same pipeline (same $t$-window $[17,43]$, same FFT
parameters). Results are collected in Table~\ref{tab:Rc}.

\begin{table}[h]
\centering
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{cccccc}
\hline
$c$ & $2c/\pi$ & $N_{\rm modes}$ & $|R(c)|$ & $\sigma_R^{\rm win}$
& $R^2$ (phase fit) \\
\hline
10 & 6.4 & 11 & 0.003 & 0.001 & 0.81 \\
20 & 12.7 & 17 & 0.004 & 0.005 & 0.32 \\
30 & 19.1 & 23 & 0.074 & 0.029 & 0.74 \\
40 & 25.5 & 30 & 0.078 & 0.021 & 0.75 \\
50 & 31.8 & 36 & 0.460 & 0.127 & 0.86 \\
100 & 63.7 & 68 & 0.477 & 0.131 & 0.73 \\
\hline
\end{tabular}
\caption{Normalized phase-winding rate $|R(c)|$ and window standard
deviation $\sigma_R^{\rm win}$ (computed over three fitting windows
$[17,43]$, $[15,40]$, $[20,45]$). Two data points do not establish
a plateau; the apparent saturation at $c \in \{50, 100\}$ may
reflect a transient crossover rather than asymptotic behavior.
Computations at $c \geq 200$ are required to distinguish
saturation from continued growth.}
\label{tab:Rc}
\end{table}

Three qualitative regimes are visible:
\begin{enumerate}[(i)]
\item \emph{Small bandwidth} ($c \leq 20$, $2c/\pi \lesssim 13$
modes): $|R(c)| < 0.005$; the phase is nearly flat. The low
$R^2$ at $c = 20$ ($R^2 = 0.32$) indicates an intermediate
disordered zone.
\item \emph{Transition} ($30 \leq c \leq 40$,
$20 \lesssim 2c/\pi \lesssim 26$ modes): $|R(c)| \approx
0.07$--$0.08$; the phase becomes more coherent ($R^2 \approx
0.74$--$0.75$) but winding remains far below the
Riemann--Siegel rate.
\item \emph{Larger bandwidth} ($c \geq 50$, $2c/\pi \gtrsim 32$
modes): $|R(c)|$ reaches approximately $0.46$--$0.48$ and
appears to \emph{plateau within numerical resolution} between
$c = 50$ and $c = 100$. This observation is based on only two
data points and may reflect a crossover rather than true
saturation; it does not constitute evidence for an asymptotic
bound on $|R(c)|$.
\end{enumerate}

A power-law fit $|R(c)| \sim c^\alpha$ on sign-consistent,
$R^2 > 0.5$ data points with $c \geq 30$ yields
$\alpha \approx 1.67$ ($R^2_{\rm fit} = 0.66$).
\emph{This fit is indicative only}: the fitted exponent reflects
a crossover regime rather than a stable asymptotic scaling, and
is not stable under small changes of the fitting range.

We summarize the finding conservatively as:

\begin{quote}
\emph{The phase-winding ratio $|R(c)|$ increases rapidly for
$30 \leq c \leq 50$, but the data are insufficient to establish
whether $|R(c)| \to 1$ as $c \to \infty$ or saturates at a
value below $1$. Computations at $c \geq 200$ would be needed
to distinguish these scenarios. Moreover, no trend toward
alignment of $|F_N(t)|$ with Riemann zero positions is observed
as $c$ increases from $10$ to $100$.}
\end{quote}

\paragraph{Interpretation.}
Individual PSWFs are bandlimited smooth functions on $[0,1]$
whose one-sided Fourier transforms $\widehat{g}_n$ have linear
phase growth (stationary-phase principle), not the logarithmic
growth of $\vartheta'(t)$. The latter encodes the functional
equation and Euler product structure of $\zeta$; this
multiplicative structure is \emph{not expected to be captured
at finite bandwidth within the present framework}.

The data therefore suggest not merely insufficient resolution,
but a \emph{mechanism-level obstruction}: even when phase
coherence emerges (regime~(iii)), it remains fundamentally
linear rather than logarithmic, indicating a structural
incompatibility between finite-bandwidth sinc-type constructions
and the multiplicative phase accumulation of $\vartheta(t)$.
This observation is the empirical precursor to the No-Go Theorem
of Paper~IV \cite{PaperIV}, which proves rigorously---under four
natural assumptions~(A)--(D)---that no PSWF-based spectral observable
$F_c(t)$ can exhibit a zeta-compatible phase in the limit $c\to\infty$.
The Mellin-bandwidth assumption central to that proof is subsequently
sharpened in Paper~V \cite{PaperV}, where $\Omega_\varepsilon(c) \to 0$
is established analytically and confirmed numerically.

\begin{remark}[No overclaiming]\label{rem:no-overclaim}
The experiments in this section are \emph{illustrative and
exploratory}, not probative. They confirm that the theoretical
objects are numerically accessible and behave in ways
\emph{consistent with} the theoretical framework. They do
\emph{not} constitute numerical evidence for or against the
Riemann Hypothesis, and they do not resolve Problem~\ref{prob:Rc-scaling}.
\end{remark}

\subsection{Summary of Numerical Experiments}
\label{ssec:summary-num}

\begin{center}
\renewcommand{\arraystretch}{1.4}
\begin{tabular}{p{4.8cm} p{5.5cm} p{3.5cm}}
\hline
\textbf{Experiment} & \textbf{Observation} & \textbf{Status} \\
\hline
Gram spectrum, $\rho < 2/\pi$
& $\lambda_{\min}(\widetilde{G}^{(N)}) > 0$,
$\kappa \to \infty$ near $\rho_*$
& Consistent with Thm.~\ref{thm:pos-def} \\[2pt]
Trace formula
& $T_N \to 1$, rate $O(1/\log N)$
& Consistent with Thm.~\ref{thm:trace} \\[2pt]
Mellin spectral density $S(t)$
& Monotone decay; no zero alignment ($p = 0.41$)
& Expected at finite $c$ \\[2pt]
Phase of $F_N(t)$
& Linear growth; $|R(c)| \ll 1$ for $c \leq 20$
& Consistent with Problem~\ref{prob:Rc-scaling} \\[2pt]
Scaling of $R(c)$
& Rapid increase for $c \in [30,50]$;
apparent plateau near $0.47$ for $c \geq 50$
& Open question \\
\hline
\end{tabular}
\end{center}

%%% ============================================================
\section{Open Problems}
\label{sec:open}
%%% ============================================================

The numerical experiments of Section~\ref{sec:numerics} and the
theoretical framework of Sections~2--4 leave several well-defined
problems open. We formulate them in order of proximity to the
results established in this paper.

%% ---------------------------------------------------------
\subsection{Scaling of the Phase-Winding Rate}
\label{ssec:prob1}

\begin{problem}[Asymptotic behavior of $R(c)$]
\label{prob:Rc-scaling}
Let $F_N(t)$ be the $\lambda$-weighted interference sum defined
in \eqref{eq:FN-def} with $N = \lfloor 4c/\pi \rfloor + 4$,
and let $R(c)$ be the normalized phase-winding rate defined in
\eqref{eq:R-def}. Determine the asymptotic behavior of $|R(c)|$
as $c \to \infty$. Specifically:
\begin{enumerate}[(a)]
\item Does $|R(c)| \to 1$?
\item Or does $|R(c)|$ saturate at a value strictly less than $1$?
\item If $|R(c)| \to 1$, at what rate? The data
(Table~\ref{tab:Rc}) are consistent with a power law
$|R(c)| \sim c^\alpha$ for $c \in [30, 50]$, with
$\alpha \approx 1.67$, but this fit is not asymptotically
stable.
\item Does there exist a scaling function $\phi(c)$ such that
$\arg F_N(t) \sim \phi(c)\,\vartheta(t)$ uniformly on compact
$t$-intervals as $c \to \infty$?
\end{enumerate}
Resolution requires computations at $c \geq 200$ with a
numerically stable high-precision Mellin pipeline.
\end{problem}

\begin{remark}[Partial resolution by Paper~IV]
\label{rem:prob61-paperIV}
The No-Go Theorem of Paper~IV \cite{PaperIV} shows that $|R(c)| \to 1$
is \emph{impossible} under assumptions~(A)--(D); thus the question
``Does $|R(c)| \to 1$?'' is answered negatively in a strong structural
sense. Problem~\ref{prob:Rc-scaling}(a) is therefore reinterpreted:
the relevant question becomes the precise \emph{rate of saturation} of
$|R(c)|$ below $1$, and whether $|R(c)|$ converges to a definite limit
strictly less than $1$.
\end{remark}

\begin{remark}[Phase transition near $2c/\pi \approx 30$]
\label{rem:phase-transition}
The numerical data suggest a qualitative change in behavior near
$2c/\pi \approx 30$ modes: below this threshold the phase of
$F_N(t)$ is essentially flat and incoherent; above it, coherent
phase growth emerges. This may be interpreted as the onset of
coherent phase dynamics beyond the Slepian transition band.
\end{remark}

%% ---------------------------------------------------------
\subsection{The Limiting Operator $\mathcal{G}_\infty$}
\label{ssec:prob2}

\begin{problem}[Existence and characterization of $\Gcinf$]
\label{prob:Ginfty}
Let $\Gc$ denote the sinc-kernel integral operator on $L^2[-1,1]$
with bandwidth $c$. Define the rescaled operator family
\begin{equation}
\label{eq:rescaled-op}
\widetilde{\mathcal{G}}_c
\;:=\;
\frac{\pi}{2c}\,\mathcal{G}_c.
\end{equation}
\begin{enumerate}[(a)]
\item Does $\widetilde{\mathcal{G}}_c$ converge in a suitable
operator topology to a limiting operator $\Gcinf$
as $c \to \infty$?
\item If so, can $\Gcinf$ be identified with a known operator
in the spectral theory of $\zeta$? Candidates include the
Berry--Keating Hamiltonian $H = xp$ \cite{BerryKeating1999},
the Connes ad\`ele-class operator \cite{Connes1999}, or the
Weil explicit-formula operator \cite{Weil1952}.
\item Does the spectrum of $\Gcinf$ on the critical line
$\{\frac{1}{2} + it : t \in \R\}$ coincide with the
multiset of Riemann zeros $\{\gamma_k\}$?
\end{enumerate}
A positive answer to~(c) would, in conjunction with
self-adjointness of $\Gcinf$, imply the Riemann Hypothesis.
We emphasize that no claim of this type is made in the present paper.
\end{problem}

%% ---------------------------------------------------------
\subsection{Uniqueness and Prime-Distribution Sensitivity}
\label{ssec:prob3}

\begin{problem}[Sensitivity of $\Gcinf$ to prime gaps]
\label{prob:prime-gaps}
The matrix $\widetilde{G}^{(N)}$ depends on the actual positions
of the primes $p_1 < \cdots < p_N$.
\begin{enumerate}[(a)]
\item Does the limiting spectral measure of
$\widetilde{G}^{(N)}$ depend only on the bulk density of
primes (i.e., on the PNT), or does it detect finer
arithmetic structure such as the pair correlation of primes
\cite{GoldstonMontgomery1987}?
\item Is the limiting operator $\Gcinf$ (if it exists)
\emph{unique}, or does it depend on the specific sequence
$\{p_k\}$?
\item If one replaces the primes by a random sequence with
the same bulk density (e.g., a Poisson process of intensity
$1/\log N$), does the spectral measure of the resulting
Gram matrix converge to the same limit?
\end{enumerate}
\end{problem}

%% ---------------------------------------------------------
\subsection{Positive Definiteness at the Threshold}
\label{ssec:prob4}

\begin{problem}[Sharp threshold and borderline behavior]
\label{prob:threshold}
Theorem~\ref{thm:pos-def} establishes positive definiteness for
$\rho < 2/\pi$.
\begin{enumerate}[(a)]
\item What is the precise divergence rate of
$\kappa(\widetilde{G}^{(N)})$ as $\rho \to (2/\pi)^-$?
\item Is the threshold $\rho_* = 2/\pi$ sharp for all
choices of point set, or only generically?
\item Does the answer depend on the specific sequence
$\{p_1,\ldots,p_N\}$?
\end{enumerate}
\end{problem}

%% ---------------------------------------------------------
\subsection{Trace Formula and Explicit Formulas}
\label{ssec:prob5}

\begin{problem}[Spectral interpretation of the explicit formula]
\label{prob:explicit-formula}
The classical Weil explicit formula \cite{Weil1952} expresses
$\theta(x)$ as a sum over Riemann zeros:
\begin{equation}
\label{eq:weil-explicit}
\theta(x)
\;=\; x
\;-\; \sum_\rho \frac{x^\rho}{\rho}
\;-\; \log(2\pi)
\;-\; \tfrac{1}{2}\log(1 - x^{-2}).
\end{equation}
\begin{enumerate}[(a)]
\item Can the spectral trace formula for $\widetilde{G}^{(N)}$
be refined to capture the error term $\theta(p_N) - p_N$,
which encodes zero distribution via \eqref{eq:weil-explicit}?
\item Is there a natural operator-theoretic analogue of the
Weil explicit formula in terms of the eigenvalues of
$\widetilde{G}^{(N)}$ or its limit $\Gcinf$?
\item Does the spectral zeta function
$Z_{\mathcal{G}}(s) := \tr(\Gcinf^{-s})$ have an analytic
continuation related to $\zeta(s)$?
\end{enumerate}
\end{problem}

%% ---------------------------------------------------------
\subsection{Summary of Open Problems}
\label{ssec:open-summary}

\begin{center}
\renewcommand{\arraystretch}{1.4}
\begin{tabular}{p{0.5cm} p{5.5cm} p{5.5cm}}
\hline
\textbf{\#} & \textbf{Problem} & \textbf{Key Obstacle} \\
\hline
6.1 & Scaling of $|R(c)|$ as $c \to \infty$
& Answered negatively in Paper~IV \cite{PaperIV}; saturation rate open \\[2pt]
6.2 & Existence of $\Gcinf$; spectral identification
& Operator-limit theory; functional analysis \\[2pt]
6.3 & Sensitivity to prime gaps vs.\ bulk density
& Random matrix / pair correlation theory \\[2pt]
6.4 & Sharp threshold $\rho_* = 2/\pi$; divergence rate
& Fine analysis of sinc-Gram determinant \\[2pt]
6.5 & Weil explicit formula in spectral form
& Spectral zeta function; analytic continuation \\[2pt]
\hline
\end{tabular}
\end{center}

\noindent
The five problems naturally separate into two classes:
\begin{itemize}
\item \emph{Computationally accessible} (6.1, 6.4):
require larger-scale numerics or finer analytic estimates,
but no new conceptual framework.
These are natural immediate targets for follow-up work.
\item \emph{Structurally deep} (6.2, 6.3, 6.5):
require substantially deeper tools from functional analysis,
random matrix theory, or arithmetic geometry.
Their resolution would represent major progress on the
spectral interpretation of $\zeta$.
\end{itemize}
Problems~6.2 and~6.5 are the most significant from a
number-theoretic perspective. Problem~6.3 is particularly timely
given recent progress on prime pair correlations
\cite{GoldstonMontgomery1987} and Montgomery's conjecture on
GUE statistics of zeros.

%%% ============================================================
\begin{thebibliography}{99}

\bibitem{PaperI}
Anonymous Author, \textit{The Sinc-Kernel Concentration Operator
and Prolate Spheroidal Wave Functions: Framework and Spectral Theory},
preprint, 2026.

\bibitem{PaperII}
Anonymous Author, \textit{Scaling Limits of Prolate Gram Forms, Uniform
Coercivity, and a Trace Formula Toward Weil Positivity},
preprint, 2026.

\bibitem{PaperIV}
Anonymous Author, \textit{No-Go Theorem for PSWF-Based
Hilbert--P\'olya Constructions: Phase Obstruction in the
Scaling Limit},
preprint, 2026.

\bibitem{PaperV}
Anonymous Author, \textit{Mellin Bandwidth Decay for Prolate
Spheroidal Wave Functions: ODE Analysis, Barrier Estimates, and
Numerical Evidence},
preprint, 2026.

\bibitem{Slepian1978}
D.~Slepian, \textit{Prolate spheroidal wave functions, Fourier
analysis, and uncertainty~-- V: The discrete case},
Bell Syst.\ Tech.\ J.\ \textbf{57} (1978), 1371--1430.

\bibitem{Osipov2013}
A.~Osipov, V.~Rokhlin, H.~Xiao,
\textit{Prolate Spheroidal Wave Functions of Order Zero},
Springer, New York, 2013.

\bibitem{Widom1964}
H.~Widom, \textit{Asymptotic behavior of the eigenvalues of
certain integral equations, II},
Arch.\ Rational Mech.\ Anal.\ \textbf{17} (1964), 215--229.

\bibitem{BerryKeating1999}
M.~V.~Berry and J.~P.~Keating, \textit{The Riemann zeros and
eigenvalue asymptotics},
SIAM Rev.\ \textbf{41} (1999), 236--266.

\bibitem{Connes1999}
A.~Connes, \textit{Trace formula in noncommutative geometry and
the zeros of the Riemann zeta function},
Selecta Math.\ (N.S.)\ \textbf{5} (1999), 29--106.

\bibitem{Weil1952}
A.~Weil, \textit{Sur les formules explicites de la th\'eorie
des nombres premiers},
Comm.\ Sem.\ Math.\ Univ.\ Lund, Tome Suppl.\ 1952, 252--265.

\bibitem{GoldstonMontgomery1987}
D.~A.~Goldston and H.~L.~Montgomery, \textit{Pair correlation of
zeros and primes in short intervals}, in \textit{Analytic Number
Theory and Diophantine Problems}, Birkh\"auser, 1987, pp.~183--203.

\bibitem{Beurling1989}
A.~Beurling, \textit{The collected works of Arne Beurling,
Vol.~2: Harmonic Analysis}, Birkh\"auser, Boston, 1989.

\bibitem{Landau1967}
H.~J.~Landau, \textit{Necessary density conditions for sampling
and interpolation of certain entire functions},
Acta Math.\ \textbf{117} (1967), 37--52.

\end{thebibliography}

\end{document}
