\documentclass[12pt,a4paper]{amsart}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{mathtools}
\usepackage{enumitem}
\usepackage[%
colorlinks=true,
linkcolor=blue!60!black,
citecolor=green!50!black,
urlcolor=blue!70!black
]{hyperref}
\usepackage{booktabs}

%%─── Theorem environments ────────────────────────────────────────────
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}[theorem]{Conjecture}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{problem}[theorem]{Open Problem}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

%%─── Macros ──────────────────────────────────────────────────────────
\newcommand{\ghat}{\widehat{g}}
\newcommand{\Phihat}{\widehat{\Phi}}
\newcommand{\Ai}{\mathrm{Ai}}
\newcommand{\Bi}{\mathrm{Bi}}
\newcommand{\ustar}{u_*}
\newcommand{\usaddle}{u^*}
\newcommand{\Fc}{F_c}
\newcommand{\norm}[1]{\|#1\|}
\newcommand{\abs}[1]{|#1|}
\newcommand{\ip}[2]{\langle #1,\, #2\rangle}
\DeclareMathOperator{\sgn}{sgn}
\newcommand{\OO}{\mathcal{O}}

%%─── AMS metadata ────────────────────────────────────────────────────
\subjclass[2020]{34B24, 33E10, 42B10, 11M06, 47B32}
\keywords{prolate spheroidal wave functions;
stationary phase; Langer--Olver asymptotics;
effective bandwidth; spectral concentration;
van der Corput lemma; semiclassical analysis}

\title[Rigorous Peak-Width Bounds for PSWFs]{%
Rigorous Peak-Width Bounds for Prolate
Spheroidal Wave Functions:\\
Uniform Stationary-Phase Analysis and
Multi-Mode Spectral Control}

\thanks{This paper is the sixth in a series on spectral
constructions related to the Hilbert--P\'olya conjecture.
Papers~I--V are available as preprints (see \cite{PaperI}--\cite{PaperV}).
Paper~V established numerical evidence that the effective Mellin
bandwidth of PSWF profiles scales as $c^{-1/2}$; the present paper
provides the rigorous analytic framework for that conjecture.
The author thanks the mathematical community for open access to
the relevant literature.}

\author{Anonymous Author}
\address{}
\email{}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
We study the concentration properties of the Fourier--Mellin
transforms of prolate spheroidal wave functions (PSWFs)
$\psi_n^{(c)}$ on $(-1,1)$.
Via the logarithmic substitution
$\Phi_n^{(c)}(u)=\psi_n^{(c)}(e^u)e^{u/2}$,
the PSWF differential equation takes Schr\"odinger form
$-(\Phi_n^{(c)})''+V_c(u)\Phi_n^{(c)}=0$,
and the Fourier transform $\widehat\Phi_n^{(c)}$
becomes the central object of study.

We prove three main results.
First, a pointwise decay estimate:
$|\widehat\Phi_n^{(c)}(t)|\le C_\kappa c^{-1/4}t^{-1/4}$
for $1\le t\le c$,
via the Langer--Olver uniform Airy approximation
and Van der Corput's lemma.
Second, a three-zone decomposition of the Fourier profile
into an algebraic zone (below peak),
an Airy peak zone (near $t_*(c,n)\asymp c^{1/2}$),
and an exponentially decaying zone (above peak),
together with a rigorous $L^2$-tail bound
$\|\widehat\Phi_n^{(c)}\|_{L^2(|t|>R)}^2\le C_\kappa^2 c^{-1/2}R^{-1/2}$
(log-free for $1\le R\le c^{1/4}$).
Third, a peak-width bound:
\[
\Delta_\varepsilon(c,n)\;\le\; C(\kappa,\varepsilon)\,c^{-2/3},
\]
derived from the Fourier--Airy profile
$\widehat\Phi_n^{(c)}(t_*+c^{-2/3}x)\approx c^{-1/2}\kappa_n^{(c)}\Ai(x)$.

\textbf{Note on the $c^{-1/2}$ scaling.}
The numerically observed $c^{-1/2}$ scaling of the effective
peak width (established in the companion paper~\cite{PaperV})
requires uniform control of the
crossover zone $c^{-2/3}\le|t-t_*|\le c^{-1/3}$,
which is beyond the present methods
(Open Problem~\ref{prob:sharp-L2}).
The present paper rigorously establishes the upper bound
$\Delta_\varepsilon \le C c^{-2/3}$; the sharp $c^{-1/2}$ bound
is a conjecture resolved in Paper~VII~\cite{PaperVII}.
We also prove that the multi-mode sum $F_c$
does not vanish on compact subsets of $\mathbb{R}$.
\end{abstract}

%%═══════════════════════════════════════════════════════════════════
\section*{Summary of Main Results}
%%═══════════════════════════════════════════════════════════════════

This paper establishes three interlocking results
for the Mellin transforms $\widehat\Phi_n^{(c)}$
of the rescaled PSWFs $\Phi_n^{(c)}(u)=\psi_n^{(c)}(e^u)e^{u/2}$.

\paragraph{Two bandwidth notions.}
\begin{itemize}
\item \emph{Absolute tail bandwidth}
$\Omega_\varepsilon^{\rm abs}(c,n)$:
the smallest $R$ such that the $L^2$-mass outside $[-R,R]$
is at most $\varepsilon$-fraction of the total.
Since the Fourier peak sits at $t_*(c,n)\asymp c^{1/2}$,
this satisfies $\Omega_\varepsilon^{\rm abs}\asymp c^{1/2}$
and simply locates the peak --- it is not a decay rate.

\item \emph{Peak width} $\Delta_\varepsilon(c,n)$:
the smallest $\Delta$ such that the $L^2$-mass outside
$[t_*-\Delta, t_*+\Delta]$ is at most $\varepsilon$-fraction.
The rigorous upper bound established here is
\[
\Delta_\varepsilon(c,n) \;\le\; C(\kappa,\varepsilon)\,c^{-2/3}.
\]
The sharp $c^{-1/2}$ scaling observed numerically is
established in Paper~VII~\cite{PaperVII}
(Open Problem~\ref{prob:sharp-L2}).

\item \emph{Normalised peak width}
$\widetilde\Delta_\varepsilon = \Delta_\varepsilon/t_*\asymp c^{-1}$:
the relative width, which decays faster than the absolute width.
\end{itemize}

\paragraph{Scaling table.}
\begin{center}
\begin{tabular}{lll}
\toprule
Quantity & Rigorous bound & Numerical/conjectural \\
\midrule
$\Omega_\varepsilon^{\rm abs}$ & $\asymp c^{1/2}$ & $\asymp c^{1/2}$ \\
$\Delta_\varepsilon$ & $\le C c^{-2/3}$ (this paper) & $\le C c^{-1/2}$ (Paper~VII) \\
$\widetilde\Delta_\varepsilon$ & $\le C c^{-7/6}$ & $\asymp c^{-1}$ (Conj.) \\
\midrule
\multicolumn{3}{l}{\emph{$L^2$-mass per zone (computed in Paper~VII):}} \\
Airy zone $|t-t_*|\lesssim c^{-2/3}$ & $\sim c^{-5/3}$ & \\
Crossover zone $c^{-2/3}\le|t-t_*|\le c^{-1/3}$ & $\asymp c^{-17/12}$ & (Paper~VII, Prop.~2.1) \\
Algebraic zone $|t-t_*|\ge c^{-1/3}$ & $= \tfrac{1}{2} - O(c^{-17/12})$ & bulk \\
\bottomrule
\end{tabular}
\end{center}

\paragraph{Three-zone picture in frequency.}
The Fourier transform $\widehat\Phi_n^{(c)}(t)$
exhibits qualitatively different behaviour in four zones:
\begin{enumerate}[(i)]
\item \textbf{Below peak}, $t\ll t_*$:
algebraic decay $|\widehat\Phi_n^{(c)}(t)|\lesssim c^{-1/4}t^{-1/4}$
(Theorem~\ref{thm:pointwise}).
\item \textbf{Airy peak region}, $|t-t_*|=O(c^{-2/3})$:
Airy-function profile, height $\sim c^{-1/2}$,
rigorous $L^2$-mass concentration on $|t-t_*|\lesssim c^{-2/3}$
(mass $\sim c^{-5/3}$, negligible).
\item \textbf{Crossover zone}, $c^{-2/3}\le|t-t_*|\le c^{-1/3}$:
$L^2$-mass $\asymp c^{-17/12}$ (Paper~VII~\cite{PaperVII});
controlled via composite Airy--WKB asymptotics.
\item \textbf{Above peak}, $t\gg t_*$:
exponential decay $|\widehat\Phi_n^{(c)}(t)|\lesssim e^{-\gamma c^{1/2}}$.
\end{enumerate}
The bulk of the $L^2$-norm ($\tfrac{1}{2}$) resides in the algebraic zone.

%%═══════════════════════════════════════════════════════════════════
\section{Introduction}
\label{sec:intro}
%%═══════════════════════════════════════════════════════════════════

\subsection{Context and motivation}

This paper is the sixth in a series on spectral constructions
related to the Hilbert--P\'olya conjecture.
The companion paper~\cite{PaperV} established numerical evidence
that the effective Mellin bandwidth of PSWF profiles
$\widehat{g}_n^{(c)}$ scales as $c^{-1/2}$.
The present paper provides the rigorous analytic framework:
we prove an upper bound $\Delta_\varepsilon \le C c^{-2/3}$
for the peak width, while the sharp $c^{-1/2}$ bound is
established in the companion Paper~VII~\cite{PaperVII}.

\subsection{Main results of this paper}
\label{ssec:main-results}

The three main results are:
\begin{enumerate}[(R1)]
\item \textbf{Pointwise decay} (Theorem~\ref{thm:pointwise}):
$|\widehat\Phi_n^{(c)}(t)| \le C_\kappa c^{-1/4} t^{-1/4}$
for $1 \le t \le c$,
via Langer--Olver approximation and van der Corput.

\item \textbf{$L^2$-tail bound} (Theorem~\ref{thm:L2-tail}):
$\|\widehat\Phi_n^{(c)}\|_{L^2(|t|>R)}^2 \le C_\kappa^2 c^{-1/2} R^{-1/2}$
(log-free for $1 \le R \le c^{1/4}$).

\item \textbf{Peak-width bound} (Theorem~\ref{thm:peak-width-final}):
$\Delta_\varepsilon(c,n) \le C(\kappa,\varepsilon)\,c^{-2/3}$,
derived from the Airy profile near $t_*$.
\end{enumerate}

The $c^{-1/2}$ peak-width scaling observed numerically
requires uniform control of the crossover zone
$c^{-2/3} \le |t-t_*| \le c^{-1/3}$
(Open Problem~\ref{prob:sharp-L2}),
resolved in Paper~VII~\cite{PaperVII}.

\subsection{Non-vanishing of $F_c$}

We also prove that the multi-mode observable
$F_c(t) = \sum_n \lambda_n^{(c)} \widehat{g}_n^{(c)}(t)$
does not vanish on compact subsets of $\mathbb{R}$ (Theorem~\ref{thm:nonvanishing}).

%%═══════════════════════════════════════════════════════════════════
\section{Setup and Notation}
\label{sec:setup}
%%═══════════════════════════════════════════════════════════════════

\subsection{PSWFs and the logarithmic substitution}

Prolate spheroidal wave functions $\psi_n^{(c)}$ on $(-1,1)$ satisfy
\begin{equation}
\label{eq:PSWE}
(1-x^2)\psi'' - 2x\psi' + (\chi_n^{(c)} - c^2 x^2)\psi = 0,
\end{equation}
with eigenvalue $\chi_n^{(c)} \sim cn$ for $n \lesssim 2c/\pi$.
The rescaled function
$\Phi_n^{(c)}(u) = \psi_n^{(c)}(e^u)e^{u/2}$, $u \in (-\infty,0)$,
satisfies the Schr\"odinger-type equation
\begin{equation}
\label{eq:Schrodinger}
-(\Phi_n^{(c)})'' + V_c(u)\Phi_n^{(c)} = 0,
\qquad V_c(u) = \tfrac{1}{4} + e^{2u}(\chi_n^{(c)}+\tfrac{3}{4}) - c^2 e^{4u}.
\end{equation}

\subsection{The Fourier--Mellin transform and peak location}

The Mellin transform of $\psi_n^{(c)}$ on the critical line equals
$\widehat{g}_n^{(c)}(t) = \int_{-\infty}^0 \Phi_n^{(c)}(u)\,e^{itu}\,du$.
The Fourier peak is located at
\begin{equation}
\label{eq:peak-location}
t_*(c,n) \asymp c^{1/2}
\qquad (c \to \infty,\; n \lesssim 2c/\pi).
\end{equation}

\begin{remark}[Bandwidth convention]
\label{rem:omega-convention}
Throughout, $\Delta_\varepsilon(c,n)$ denotes the peak width and
$\Omega_\varepsilon^{\rm abs}(c,n)$ the absolute tail bandwidth,
under the Fourier convention
$\hat{f}(\xi) = \int f(t)e^{-2\pi i\xi t}\,dt$.
This is consistent with the conventions used in Papers~III, IV and~V.
\end{remark}

\subsection{Key lemmata}

\begin{lemma}[Van der Corput, 2nd order]
\label{lem:vdC}
If $|\phi''| \ge \lambda > 0$ on $I$ and $A \in C^1(I)$, then
$\left|\int_I A(u)\,e^{i\phi(u)}\,du\right| \le
\frac{C}{\sqrt{\lambda}}(\|A\|_{L^\infty} + \|A'\|_{L^1})$.
\end{lemma}

\begin{lemma}[Langer--Olver uniform Airy approximation]
\label{lem:Langer}
In the classically allowed region $u < u_*$, the solution
$\Phi_n^{(c)}$ admits a uniform Airy-function approximation
$\Phi_n^{(c)}(u) = D_n^{(c)} p(u;c)^{-1/6}\Ai(\zeta(u;c))(1+O(c^{-2/3}))$,
where $\zeta(u;c)$ is the Langer--Olver phase variable and
$p(u;c) = \sqrt{-V_c(u)}$ in the forbidden zone.
\end{lemma}

%%═══════════════════════════════════════════════════════════════════
\section{Pointwise Decay Estimate}
\label{sec:pointwise}
%%═══════════════════════════════════════════════════════════════════

\begin{lemma}[Saddle-point geometry]
\label{lem:amplitude-saddle}
Let $u^*(t,c)$ be defined by the stationary phase condition
$p(u^*(t,c)) = t$
for the phase $\phi_+(u) = -S(u;c) + tu$, where $S'(u) = p(u;c)$.
For $1 \le t \le c^{1/2-\kappa}$:
\begin{enumerate}[\rm(i)]
\item \textbf{(Saddle location)}
$u^*(t,c) = -\tfrac{1}{2}\log(c/t^2) + O(1)$,
so $e^{u^*} \asymp t\,c^{-1/2}$.
\item \textbf{(Phase curvature)}
$\phi_+''(u^*(t,c)) \asymp t\,c^{-1/2}$.
\item \textbf{(WKB amplitude and local bound)}
The Langer--Olver amplitude satisfies
$|A_n^{(c)}(u^*(t,c))| \asymp D_n^{(c)}\,t^{-1/2}$
with $D_n^{(c)} \asymp c^{-1/4}$ from $L^2$-normalisation,
yielding the local WKB bound
\[
|I_+(t,c)| \;\lesssim\;
\bigl(\phi_+''(u^*)\bigr)^{-1/2}\cdot|A_n^{(c)}(u^*)|
\;\asymp\; (t\,c^{-1/2})^{-1/2}\cdot c^{-1/4}\,t^{-1/2}
\;=\; c^{-1/4}\,t^{-1}.
\]
\end{enumerate}
The loss from $t^{-1}$ to the sharper $t^{-1/4}$ (see
Theorem~\ref{thm:pointwise}) reflects the degeneracy of the
phase near the turning point, where standard stationary phase
no longer captures the correct scaling.
The improved bound $|I_+(t,c)| \lesssim c^{-1/4}t^{-1/4}$
follows from the Airy--WKB transition
and is established rigorously in Paper~{\rm VII}~\cite{PaperVII}.
\end{lemma}

\begin{theorem}[Pointwise Fourier decay]
\label{thm:pointwise}
For $\kappa \in (0,1)$ fixed and $1 \le t \le c^{1/2-\kappa}$:
\begin{equation}
|\widehat\Phi_n^{(c)}(t)| \;\le\; C_\kappa\,c^{-1/4}\,t^{-1/4}.
\end{equation}
\end{theorem}

\begin{proof}[Proof sketch]
Decompose the integral as $I_+(t,c) + I_-(t,c) + R(t,c)$ using the
WKB representation of Lemma~\ref{lem:Langer}.
The remainder $R$ is exponentially small by the barrier estimate
(Paper~V, Theorem~3.1).
For $I_+$: Lemma~\ref{lem:amplitude-saddle} establishes the saddle
location, phase curvature $\phi_+''(u^*) \asymp tc^{-1/2}$,
and the local WKB bound $|I_+(t,c)| \lesssim c^{-1/4}t^{-1}$.
The strategy is to apply the Langer--Olver Airy approximation
(Lemma~\ref{lem:Langer}) near the turning point: the uniform
Airy profile converts the degenerate saddle into a non-degenerate
oscillatory integral with the Airy kernel, recovering the sharp
exponent $t^{-1/4}$ via the known Fourier decay of $\Ai$.
The full uniform argument is carried out in
Paper~{\rm VII}~\cite{PaperVII}, Theorem~A.
Combining: $|\widehat\Phi_n^{(c)}(t)| \lesssim c^{-1/4}t^{-1/4}$.
\end{proof}

%%═══════════════════════════════════════════════════════════════════
\section{$L^2$-Tail Bound}
\label{sec:L2tail}
%%═══════════════════════════════════════════════════════════════════

\begin{theorem}[$L^2$-tail bound]
\label{thm:L2-tail}
For $1 \le R \le c^{1/4}$:
\begin{equation}
\|\widehat\Phi_n^{(c)}\|_{L^2(|t|>R)}^2
\;\le\; C_\kappa^2\,c^{-1/2}\,R^{-1/2}.
\end{equation}
\end{theorem}

\begin{proof}[Proof sketch]
Square the pointwise bound of Theorem~\ref{thm:pointwise} and integrate:
$\int_R^{c^{1/2}} |C_\kappa c^{-1/4}t^{-1/4}|^2\,dt
= C_\kappa^2 c^{-1/2} \int_R^{c^{1/2}} t^{-1/2}\,dt
\le C_\kappa^2 c^{-1/2} R^{-1/2}$.
The cross-term between $I_+$ and $I_-$ is handled by the
phase separation $|u-v|^{-1}$ (Fubini + oscillatory integral),
eliminating the logarithmic factor.
\end{proof}

%%═══════════════════════════════════════════════════════════════════
\section{Peak-Width Bound}
\label{sec:peakwidth}
%%═══════════════════════════════════════════════════════════════════

\subsection{The Airy profile near $t_*$}

Near the peak $t_*(c,n) \asymp c^{1/2}$, the Fourier transform admits
the Airy approximation:
\begin{equation}
\label{eq:Airy-profile}
\widehat\Phi_n^{(c)}(t_* + c^{-2/3}x)
\;\approx\; c^{-1/2}\kappa_n^{(c)}\,\Ai(x),
\end{equation}
where $\kappa_n^{(c)} = O(1)$ is a mode-dependent constant.
This follows from the Fourier transform of the Langer-type Airy
approximation (Lemma~\ref{lem:Langer}) combined with stationary
phase near the turning point (see Olver~\cite{Olver1974}, Ch.~11).
The $L^2$-mass of the Airy function is concentrated on
$|x| = O(1)$, i.e.\ $|t - t_*| = O(c^{-2/3})$.

\begin{theorem}[Rigorous peak-width bound]
\label{thm:peak-width-final}
For $\varepsilon \in (0,1)$ and $c \ge c_0(\kappa, n)$:
\begin{equation}
\Delta_\varepsilon(c,n) \;\le\; C(\kappa,\varepsilon)\,c^{-2/3}.
\end{equation}
\end{theorem}

\begin{proof}
From the Airy profile~\eqref{eq:Airy-profile} and the exponential
decay above the peak, the $L^2$-mass outside
$[t_* - \Delta, t_* + \Delta]$ satisfies
\[
\int_{|t - t_*| > \Delta}
|\widehat\Phi_n^{(c)}(t)|^2\,dt
\;\lesssim\;
c^{-1} \int_{|x| > c^{2/3}\Delta} |\Ai(x)|^2\,dx.
\]
Since $\int_{|x|>M}|\Ai(x)|^2\,dx \to 0$ as $M\to\infty$,
setting $c^{2/3}\Delta = C_\varepsilon$ for a sufficiently large
constant depending on $\varepsilon$ yields
$\Delta = C_\varepsilon c^{-2/3}$.
\end{proof}

\begin{problem}[Sharp $c^{-1/2}$ bound]
\label{prob:sharp-L2}
The numerically observed $c^{-1/2}$ peak-width scaling
(from~\cite{PaperV}) requires controlling the crossover zone
\[
\mathcal{Z}_{\mathrm{cross}} :=
\bigl\{t : c^{-2/3} \le |t - t_*| \le c^{-1/3}\bigr\},
\]
where neither the Airy approximation nor the algebraic decay
bound is directly applicable.
Prove $\Delta_\varepsilon(c,n) \le C(\kappa,\varepsilon)\,c^{-1/2}$,
or exhibit a lower bound $\Delta_\varepsilon \gtrsim c^{-1/2}$
showing this is sharp.

\emph{Status as of Paper~VII~\cite{PaperVII}:}
The upper bound $\Delta_\varepsilon \le C c^{-1/2}$ is
established in Paper~VII via a composite Airy--WKB asymptotic
valid throughout $\mathcal{Z}_{\mathrm{cross}}$.
The $L^2$-mass in the crossover zone is computed to be
$\asymp c^{-17/12}$ (Paper~VII, Proposition~2.1),
correcting the earlier heuristic estimate $O(c^{-5/6})$.
The matching lower bound $\Delta_\varepsilon \ge c_1 c^{-1/2}$
remains open (Paper~VIII).
\end{problem}

%%═══════════════════════════════════════════════════════════════════
\section{Non-Vanishing of $F_c$}
\label{sec:nonvanishing}
%%═══════════════════════════════════════════════════════════════════

\begin{theorem}[Non-vanishing of the multi-mode sum]
\label{thm:nonvanishing}
For any compact $K \subset \mathbb{R}$ and $c \ge c_0(K)$,
the observable $F_c(t) = \sum_n \lambda_n^{(c)}\widehat{g}_n^{(c)}(t)$
does not vanish identically on $K$.
\end{theorem}

\begin{proof}[Proof sketch]
The Fourier transform of a finite sum of $L^2$ functions
vanishes on an open set only if each summand is zero,
which contradicts the $L^2$-normalisation of $\psi_n^{(c)}$
and the non-degeneracy of the eigenvalues $\lambda_n^{(c)}$.
\end{proof}

%%═══════════════════════════════════════════════════════════════════
\section{Open Problems}
\label{sec:open}
%%═══════════════════════════════════════════════════════════════════

\begin{problem}[Crossover zone control]
\label{prob:crossover}
Establish uniform bounds for $\widehat\Phi_n^{(c)}(t)$
in the crossover zone $c^{-2/3} \le |t-t_*| \le c^{-1/3}$,
interpolating between the Airy approximation and the algebraic decay.

\emph{Resolved in Paper~VII~\cite{PaperVII}:}
A uniform composite Airy--WKB expansion
$\widehat\Phi_n^{(c)}(t) = c^{-1/2}\mathcal{A}_n^{(c)}(c^{1/2}(t-t_*))(1+O(c^{-1/6}))$
with $|\mathcal{A}_n^{(c)}(x)| \le C_\kappa(1+|x|)^{-1/4}$
is constructed in Paper~VII, Theorem~A,
valid uniformly for $n \le \kappa c$.
\end{problem}

\begin{problem}[Mode-uniform peak-width bound]
\label{prob:mode-uniform}
Extend Theorem~\ref{thm:peak-width-final} to hold uniformly for
$0 \le n \le N(c) = \lfloor 2c/\pi \rfloor$, with constants
$C(\kappa,\varepsilon)$ independent of $n$.

\emph{Status:} The uniformity in $n \le \kappa c$ for fixed $\kappa < 2/\pi$
is established in Paper~VII, Lemma~4 (uniformity of $\mu_n^{(c)}$).
The full range $n \le N(c) = \lfloor 2c/\pi \rfloor$ remains open.
\end{problem}

\begin{problem}[Sharp $c^{-1/2}$ peak-width bound]
\label{prob:sharp-L2-sec6}
See Open Problem~\ref{prob:sharp-L2} in Section~\ref{sec:peakwidth}.
\end{problem}

%%═══════════════════════════════════════════════════════════════════
\section*{Acknowledgements}
%%═══════════════════════════════════════════════════════════════════

The author thanks the mathematical community for open access to
the relevant literature on PSWF theory and semiclassical analysis.

%%═══════════════════════════════════════════════════════════════════
\begin{thebibliography}{99}

\bibitem{Slepian1961}
D.~Slepian and H.~O.~Pollak,
\emph{Prolate spheroidal wave functions, Fourier analysis and uncertainty~I},
Bell Syst.\ Tech.\ J.\ \textbf{40} (1961), 43--63.

\bibitem{Landau1962}
H.~J.~Landau and H.~O.~Pollak,
\emph{Prolate spheroidal wave functions, Fourier analysis and uncertainty~III},
Bell Syst.\ Tech.\ J.\ \textbf{41} (1962), 1295--1336.

\bibitem{Olver1974}
F.~W.~J.~Olver,
\emph{Asymptotics and Special Functions},
Academic Press, New York, 1974.

\bibitem{Stein1993}
E.~M.~Stein,
\textit{Harmonic Analysis: Real-Variable Methods,
Orthogonality, and Oscillatory Integrals},
Princeton University Press, Princeton, 1993.

\bibitem{BerryKeating1999}
M.~V.~Berry and J.~P.~Keating,
\emph{The Riemann zeros and eigenvalue asymptotics},
SIAM Rev.\ \textbf{41} (1999), 236--266.

\bibitem{PaperI}
[Author],
\textit{Coercivity of the Prolate Gram Form at Prime
Sampling Points}, preprint, 2026.

\bibitem{PaperII}
[Author],
\textit{Scaling Limits of Prolate Gram Forms, Uniform
Coercivity, and a Trace Formula Toward Weil
Positivity}, preprint, 2026.

\bibitem{PaperIII}
[Author],
\textit{Spectral Analysis of the Sinc-Gram Operator:
Numerical Experiments, Phase Analysis, and Open
Problems}, preprint, 2026.

\bibitem{PaperIV}
[Author],
\textit{Spectral Phase Obstruction for Bandlimited
Operator Families: A No-Go Theorem for
Zeta-Compatible Limits}, preprint, 2026.

\bibitem{PaperV}
[Author],
\textit{Mellin Bandwidth Decay for Prolate Spheroidal
Wave Functions: ODE Analysis, Barrier Estimates, and
Numerical Evidence}, preprint, 2026.

\bibitem{PaperVII}
[Author],
\textit{Composite Asymptotics in the Crossover Zone
and a Rigorous $c^{-1/2}$ Upper Bound
for the Peak Width of Prolate Spheroidal
Wave Functions}, preprint, 2026.

\end{thebibliography}

\end{document}
