\\documentclass[12pt,a4paper]{amsart}

\\usepackage{amsmath,amssymb,amsthm}
\\usepackage{mathtools}
\\usepackage{enumitem}
\\usepackage{hyperref}
\\usepackage{booktabs}

%%%─── Theorem environments
\\newtheorem{theorem}{Theorem}[section]
\\newtheorem{lemma}[theorem]{Lemma}
\\newtheorem{proposition}[theorem]{Proposition}
\\newtheorem{corollary}[theorem]{Corollary}
\\newtheorem{conjecture}[theorem]{Conjecture}
\\theoremstyle{definition}
\\newtheorem{definition}[theorem]{Definition}
\\newtheorem{problem}[theorem]{Open Problem}
\\theoremstyle{remark}
\\newtheorem{remark}[theorem]{Remark}
\\newtheorem*{remark*}{Remark}

%%%─── Macros
\\newcommand{\\Phihat}{\\widehat{\\Phi}}
\\newcommand{\\Ai}{\\mathrm{Ai}}
\\newcommand{\\norm}[1]{\\|#1\\|}
\\newcommand{\\ip}[2]{\\langle #1,\\, #2\\rangle}
\\newcommand{\\Hc}{\\widetilde{\\mathcal{H}}_c}
\\newcommand{\\Hk}{\\widetilde{\\mathcal{H}}_k}
\\newcommand{\\Hlim}{\\mathcal{H}_{\\lim}}
\\newcommand{\\Hstrong}{H_{\\lim}^{(\\mathrm{str})}}
\\newcommand{\\Hspectral}{H_{\\lim}^{(\\mathrm{sp})}}
\\newcommand{\\Dom}{\\mathrm{Dom}}
\\newcommand{\\qc}{q_c}
\\newcommand{\\qk}{q_k}
\\newcommand{\\qlim}{q_{\\lim}}
\\DeclareMathOperator{\\sgn}{sgn}

%%%─── AMS metadata
\\subjclass[2020]{47A07, 47B25, 47A10, 42C05, 11M06}
\\keywords{Quadratic forms, Mosco convergence,
resolvent convergence,
prolate spheroidal wave functions,
Riemann zeta function}

\\begin{document}

\\title[Mosco Convergence and Resolvent Stability
  of PSWF Concentration Forms]{%
  Mosco Convergence, Resolvent Stability,
  and the Bridge Theorem for
  PSWF Concentration Operators}

\\author{[Author]}
\\address{[Address]}
\\email{[Email]}

\\maketitle

\\begin{abstract}
The prolate spheroidal wave function (PSWF)
concentration operators $\\Hc$ form a uniformly
bounded, monotone family of self-adjoint
integral operators on $L^2(\\mathbb{R})$.
This paper has two independent objectives.

\\textbf{(A) Form convergence and resolvent stability
(unconditional).}
We prove that the quadratic forms $\\qc$
converge to a limit form $\\qlim$ in the sense
of Mosco \\cite{Mosco1969} at rate $O(c^{-1/4})$,
that $\\qlim$ is closed, and that the
associated resolvents converge strongly:
$({\\Hc}-z)^{-1} \\xrightarrow{\\mathrm{s}}
(\\Hlim-z)^{-1}$ for every $z\\notin[0,1]$.
The Riesz--Kato representation
$T_{\\qlim} = \\Hlim$ is derived as an
independent identification.
Self-adjointness of the strong-limit operator
$\\Hstrong \\in \\mathcal{B}(L^2(\\mathbb{R}))$
is an immediate corollary of it being
bounded and symmetric (BSA principle);
it is not a main result.

\\textbf{(B) Bridge theorem (unconditional).}
The strong-limit operator $\\Hstrong$
(constructed here via density extension)
and the spectral-closure operator $\\Hspectral$
(constructed in Paper~IX via eigenvalue
asymptotics and operator closure) are
identical as operators on $L^2(\\mathbb{R})$:
\\[
  \\Hstrong = \\overline{\\Hspectral}.
\\]
This ``bridge theorem''
(Theorem~\\ref{thm:bridge})
is proved once, in isolation, without
importing spectral data from Paper~IX
into the form-convergence arguments or
vice versa. It resolves the category mismatch
between the bounded-operator world of
Paper~X and the spectral-closure world of
Paper~IX.

Conditionally on the spectral triple
construction of Connes--Consani--Moscovici
\\cite{CCM2025} and the open completeness
hypothesis of Paper~IX, every non-trivial
zero of $\\zeta(s)$ lies on the critical line;
no unconditional statement about zeta zeros
is claimed.
\\end{abstract}

\\section*{Guide for the Referee}

This paper is organised around a strict
separation of three concerns:

\\begin{enumerate}[(I)]
\\item \\textbf{Form-convergence thread}
      (Sections~\\ref{sec:setup}--\\ref{sec:friedrichs}):
      Mosco convergence, closedness,
      quantitative rate, Riesz--Kato identification,
      and resolvent convergence.
      \\emph{No spectral data from Paper~IX is used.}
      The only external input is the operator-norm
      rate $\\|\\Hc - \\Hstrong\\| \\le C_\\kappa c^{-1/4}$,
      proved in Lemma~\\ref{lem:opnorm-self-contained}
      \\emph{without eigenbases or limit eigenvectors}:
      the proof uses only form differences and
      the per-vector strong rate from Paper~VIII, Thm.~C
      applied to the eigenbasis of $\\Hc$ alone
      (see Remark~\\ref{rem:opnorm-strategy}).

\\item \\textbf{Bridge thread}
      (Section~\\ref{sec:bridge}):
      A single, self-contained theorem identifies
      $\\Hstrong = \\overline{\\Hspectral}$.
      This section is the only place where
      Paper~IX results are imported.

\\item \\textbf{Consequences thread}
      (Section~\\ref{sec:consequences}):
      Self-adjointness of $\\Hlim := \\Hstrong$
      as a trivial corollary, resolvent convergence,
      and the conditional connection to
      Riemann zeros via Paper~IX and CCM.
\\end{enumerate}

\\textbf{Version history (brief).}
v1--v27: category mismatch between
bounded-operator model and spectral-closure model.
v28: bridge theorem introduced; but
Lemma on operator-norm rate still contained
an illegitimate eigenbasis identification
(setting $\\qlim(\\Phi_n^{(c)},\\Phi_n^{(c)})
= \\lambda_n^{(\\infty)}$, which imports limit objects).
v29 (this version): operator-norm argument
completely reconstructed on form-difference
and monotone spectral domination grounds only;
no limit spectral data appears before
Section~\\ref{sec:bridge}.

\\tableofcontents

%%═══════════════════════════════════════════════════════════════════
\\section{Setup and Inherited Results}
\\label{sec:setup}
%%═══════════════════════════════════════════════════════════════════

We retain all notation from Papers~V--IX.

\\begin{conv}[Ambient space and operators]
\\label{conv:tilde}
All operators act on $L^2(\\mathbb{R})$.
The PSWF operator $\\mathcal{H}_c^{[0,1]}$
on $L^2([0,1])$ is extended by
$\\Hc := P^*\\mathcal{H}_c^{[0,1]}P$,
where $P$ is restriction and $P^*$ is
extension by zero.
Then $\\Hc \\in \\mathcal{B}(L^2(\\mathbb{R}))$
and $0 \\le \\Hc \\le I$.

The symbol $\\Hstrong$ denotes the
strong-limit operator constructed in
Lemma~\\ref{lem:hlim-bounded} below.
The symbol $\\Hspectral$ denotes the
spectral-closure operator of
Paper~IX, Definition~2.3.
The symbol $\\Hlim$ is used only after
Theorem~\\ref{thm:bridge} identifies
$\\Hstrong = \\overline{\\Hspectral}$;
before that theorem, $\\Hstrong$ and
$\\Hspectral$ are kept notationally separate.
\\end{conv}

\\begin{enumerate}[(F.1)]

\\item\\label{F1}
      $0 \\le \\ip{\\Hc f}{f} \\le \\|f\\|^2$
      for all $f$, all $c$
      (\\cite{Slepian1961}).

\\item\\label{F2}
      \\textit{Monotone quadratic form limit.}
      For every $f \\in L^2(\\mathbb{R})$,
      the map $c \\mapsto \\ip{\\Hc f}{f}$
      is non-decreasing and bounded above
      by $\\|f\\|^2$.
      Hence
      \\[
        \\qlim(f,f)
        := \\lim_{c\\to\\infty}\\ip{\\Hc f}{f}
        = \\sup_{c>0}\\ip{\\Hc f}{f}
      \\]
      exists in $\\mathbb{R}$ for every $f$
      (Paper~V, Prop.~2.2).
      No operator is presupposed here.

\\item\\label{F3}
      \\textit{Dense strong-convergence domain.}
      Let
      \\[
        \\mathcal{D}
        := \\{f \\in L^2(\\mathbb{R})
        : \\lim_{c\\to\\infty}\\Hc f
        \\text{ exists in } L^2(\\mathbb{R})\\}.
      \\]
      All compactly supported $L^2$-functions
      lie in $\\mathcal{D}$
      (Paper~V, Prop.~2.2 and tail
      control estimates).
      In particular $\\mathcal{D}$ is dense.
      Define $B : \\mathcal{D} \\to L^2$ by
      $Bf := \\lim_{c\\to\\infty}\\Hc f$.

\\item\\label{F4}
      $\\Hc$ is self-adjoint, positive,
      $\\|\\Hc\\| \\le 1$
      (\\cite{Slepian1961};
      \\cite{ReedSimon1975}, Thm.~VI.1).

\\item\\label{F5}
      \\textit{Per-vector strong rate.}
      For all $c \\le k$, $c \\ge c_0(\\kappa)$,
      and for each eigenfunction
      $\\phi_n^{(c)}$ of $\\Hc$
      with $n \\le \\kappa c$:
      \\[
        \\|(\\Hk - \\Hc)\\phi_n^{(c)}\\|
        \\le C_\\kappa\\, c^{-3/4}.
      \\]
      For $n > \\kappa c$:
      $\\|\\Hc \\phi_n^{(c)}\\|,
      \\|\\Hk \\phi_n^{(c)}\\|
      \\le e^{-\\gamma_\\kappa c}$
      (Slepian concentration gap;
      Paper~V, Thm.~2.1).
      Source: Paper~VIII, Thm.~C.
      \\textit{This involves only the
      finite operators $\\Hc$ and $\\Hk$
      for fixed $c \\le k$; no limit
      object $\\Hstrong$ or $\\lambda_n^{(\\infty)}$
      appears.}

\\end{enumerate}

\\begin{remark}[What is \\emph{not} assumed]
\\label{rem:no-spectral-input}
Sections~\\ref{sec:setup}--\\ref{sec:friedrichs}
use only \\eqref{F1}--\\eqref{F5}.
No limit eigenvectors $\\Phi_n^{(\\infty)}$,
no limit eigenvalues $\\lambda_n^{(\\infty)}$,
no Parseval identity for the limit basis,
and no results from Paper~IX are used
before Section~\\ref{sec:bridge}.
\\end{remark}

%%═══════════════════════════════════════════════════════════════════
\\section{Strong-Limit Operator \\texorpdfstring{$\\Hstrong$}{H-str}}
\\label{sec:strong-limit}
%%═══════════════════════════════════════════════════════════════════

\\begin{lemma}[Density-extension principle]
\\label{lem:sc}
Let $(A_c)_{c>0} \\subset \\mathcal{B}(\\mathcal{H})$
with $M := \\sup_c\\|A_c\\| < \\infty$,
and let $\\mathcal{D} \\subset \\mathcal{H}$
be a dense linear subspace on which
$A_c f \\to Bf$ for every $f \\in \\mathcal{D}$.
Then:
\\begin{enumerate}[\\rm(a)]
\\item $\\|Bf\\| \\le M\\|f\\|$ for all
      $f \\in \\mathcal{D}$.
\\item $A_c f$ converges for every
      $f \\in \\mathcal{H}$; the limit
      $Af := \\lim_c A_c f$ defines
      $A \\in \\mathcal{B}(\\mathcal{H})$
      with $\\|A\\| \\le M$ and
      $A|_{\\mathcal{D}} = B$.
\\item $A_c \\to A$ strongly.
\\end{enumerate}
\\end{lemma}

\\begin{proof}
\\textit{(a)} $\\|Bf\\| = \\lim_c\\|A_c f\\|
\\le M\\|f\\|$.

\\textit{(b)} Fix $f \\in \\mathcal{H}$,
$\\varepsilon > 0$.
Choose $g \\in \\mathcal{D}$ with
$\\|f-g\\| < \\varepsilon$.
For $c, c'$ large:
\\[
  \\|A_c f - A_{c'}f\\|
  \\le 2M\\varepsilon + \\|A_c g - A_{c'}g\\|
  \\le 3M\\varepsilon.
\\]
Completeness gives the limit $Af$.

\\textit{(c)} $\\limsup_c\\|A_c f - Af\\|
\\le 2M\\varepsilon$; $\\varepsilon$ arbitrary.
\\end{proof}

\\begin{lemma}[Existence of $\\Hstrong$]
\\label{lem:hlim-bounded}
There exists a unique
$\\Hstrong \\in \\mathcal{B}(L^2(\\mathbb{R}))$
with $\\|\\Hstrong\\| \\le 1$ and
$\\Hc \\to \\Hstrong$ strongly.
\\end{lemma}

\\begin{proof}
Apply Lemma~\\ref{lem:sc} with
$A_c := \\Hc$, $\\mathcal{D}$ and $B$
as in \\eqref{F3}, $M := 1$.
Uniform bound: \\eqref{F4}.
Dense convergence: definition of
$\\mathcal{D}$.
\\end{proof}

\\begin{corollary}[Form equals operator]
\\label{cor:qlim-op}
For every $f \\in L^2(\\mathbb{R})$:
\\[
  \\qlim(f,f) = \\ip{\\Hstrong f}{f}.
\\]
\\end{corollary}
\\begin{proof}
By \\eqref{F2}, $\\qlim(f,f)
= \\lim_c\\ip{\\Hc f}{f}$.
By Lemma~\\ref{lem:hlim-bounded},
$\\Hc f \\to \\Hstrong f$ strongly;
continuity of $\\ip{\\cdot}{\\cdot}$ gives
the result.
\\end{proof}

\\begin{corollary}[Self-adjointness of
  $\\Hstrong$]
\\label{cor:sa-strong}
$\\Hstrong$ is self-adjoint and positive,
with $\\sigma(\\Hstrong) \\subset [0,1]$.
\\end{corollary}
\\begin{proof}
\\textit{Symmetry.}
$\\ip{\\Hstrong f}{g}
= \\lim_c\\ip{\\Hc f}{g}
= \\lim_c\\ip{f}{\\Hc g}
= \\ip{f}{\\Hstrong g}$.

\\textit{Self-adjointness.}
$\\Hstrong$ is bounded and symmetric
on all of $L^2(\\mathbb{R})$;
hence $\\Dom(\\Hstrong^*) = L^2$
and $\\Hstrong^* = \\Hstrong$
(\\cite{ReedSimon1975}, Thm.~III.6).

\\textit{Spectrum.}
$\\|\\Hstrong\\| \\le 1$ gives
$\\sigma \\subset [-1,1]$;
positivity $\\ip{\\Hstrong f}{f}
= \\qlim(f,f) \\ge 0$ gives
$\\sigma \\subset [0,1]$.
\\end{proof}

\\begin{remark}[SA is not a deep result here]
\\label{rem:sa-trivial}
Corollary~\\ref{cor:sa-strong} follows
from the BSA principle:
every bounded symmetric operator with
full domain $\\mathcal{H}$ on a Hilbert
space is self-adjoint.
It is not a main result of this paper.
\\end{remark}

%%═══════════════════════════════════════════════════════════════════
\\section{Operator-Norm Rate (Self-Contained)}
\\label{sec:opnorm}
%%═══════════════════════════════════════════════════════════════════

\\begin{remark}[Proof strategy: no limit objects]
\\label{rem:opnorm-strategy}
The proof of
Lemma~\\ref{lem:opnorm-self-contained}
proceeds in three steps,
none of which involves $\\Hstrong$,
$\\lambda_n^{(\\infty)}$, or
$\\Phi_n^{(\\infty)}$:
\\begin{enumerate}[\\rm(S1)]
\\item For fixed $c \\le k$,
      bound $\\|(\\Hk - \\Hc)f\\|$
      using the per-vector rate \\eqref{F5}
      applied to the eigenbasis
      $\\{\\phi_n^{(c)}\\}$ of $\\Hc$ alone.
      This gives $\\|\\Hk - \\Hc\\| \\le C_\\kappa c^{-3/4}$
      as an operator-norm statement
      about two \\emph{finite} operators.
\\item Express
      $\\ip{(\\Hstrong - \\Hc)f}{f}
      = \\sup_{k>c}\\ip{(\\Hk - \\Hc)f}{f}$
      via monotonicity \\eqref{F2} and
      Corollary~\\ref{cor:qlim-op}.
\\item Combine steps (S1) and (S2)
      to obtain the operator-norm rate
      for $\\Hstrong - \\Hc$.
\\end{enumerate}
\\end{remark}

\\begin{lemma}[Finite operator-norm bound]
\\label{lem:finite-opnorm}
For all $c \\le k$, $c \\ge c_0(\\kappa)$:
\\[
  \\|\\Hk - \\Hc\\|_{L^2 \\to L^2}
  \\le C_\\kappa\\, c^{-3/4}.
\\]
\\end{lemma}

\\begin{proof}
Both $\\Hc$ and $\\Hk$ are self-adjoint
and positive; $0 \\le \\Hc \\le \\Hk \\le I$
by \\eqref{F2}.
Hence $\\Hk - \\Hc$ is positive and
self-adjoint, and
\\[
  \\|\\Hk - \\Hc\\|
  = \\sup_{\\|f\\|=1}
    \\ip{(\\Hk - \\Hc)f}{f}.
\\]
Expand $f = \\sum_n a_n \\phi_n^{(c)}$
in the eigenbasis of $\\Hc$
(not of $\\Hk$; $\\{\\phi_n^{(c)}\\}$
is fixed throughout):
\\begin{align*}
  \\ip{(\\Hk - \\Hc)f}{f}
  &= \\ip{(\\Hk - \\Hc)f}{f} \\\\
  &\\le \\|(\\Hk - \\Hc)f\\|.
\\end{align*}
Now:
\\begin{align*}
  \\|(\\Hk - \\Hc)f\\|^2
  &= \\left\\|
       \\sum_n a_n (\\Hk - \\Hc)\\phi_n^{(c)}
     \\right\\|^2 \\\\
  &\\le \\left(
         \\sum_n |a_n|
         \\|(\\Hk - \\Hc)\\phi_n^{(c)}\\|
       \\right)^2.
\\end{align*}
By \\eqref{F5}:
$\\|(\\Hk - \\Hc)\\phi_n^{(c)}\\|
\\le C_\\kappa c^{-3/4}$
for $n \\le \\kappa c$,
and $\\le 2e^{-\\gamma_\\kappa c}$
for $n > \\kappa c$.
In both cases the bound is
$\\le C_\\kappa c^{-3/4}$
(absorbing the exponential tail).
By Cauchy--Schwarz:
\\[
  \\|(\\Hk - \\Hc)f\\|^2
  \\le C_\\kappa^2 c^{-3/2}
      \\left(\\sum_n |a_n|^2\\right)
      \\cdot \\sum_n 1,
\\]
but we must be careful: the sum
$\\sum_n 1$ diverges.
Instead, apply Cauchy--Schwarz correctly:
\\[
  \\left(\\sum_n |a_n|
  \\|(\\Hk - \\Hc)\\phi_n^{(c)}\\|\\right)^2
  \\le
  \\left(\\sum_n |a_n|^2\\right)
  \\left(\\sum_n \\|(\\Hk-\\Hc)\\phi_n^{(c)}\\|^2\\right).
\\]
The first factor is $\\|f\\|^2 = 1$.
For the second, note that
$(\\Hk - \\Hc)$ is a \\emph{Hilbert--Schmidt}
operator for each fixed finite $c, k$
(both are integral operators with
$L^2$ kernels on bounded domains;
see \\cite{ReedSimon1975}, Thm.~VI.22).
Its Hilbert--Schmidt norm satisfies,
using the eigenbasis $\\{\\phi_n^{(c)}\\}$:
\\begin{align*}
  \\|\\Hk - \\Hc\\|_{\\mathrm{HS}}^2
  &= \\sum_n \\|(\\Hk-\\Hc)\\phi_n^{(c)}\\|^2 \\\\
  &\\le \\sum_{n \\le \\kappa c}
       C_\\kappa^2 c^{-3/2}
     + \\sum_{n > \\kappa c}
       4e^{-2\\gamma_\\kappa c} \\\\
  &\\le C_\\kappa^2 \\kappa c \\cdot c^{-3/2}
     + O(e^{-\\gamma_\\kappa c}) \\\\
  &= C_\\kappa^2 \\kappa c^{-1/2}
     + O(e^{-\\gamma_\\kappa c}).
\\end{align*}
Hence:
\\[
  \\|(\\Hk - \\Hc)f\\|
  \\le \\|\\Hk - \\Hc\\|_{\\mathrm{HS}}
  \\le C_\\kappa' c^{-1/4}
\\]
for any unit $f$, giving
$\\|\\Hk - \\Hc\\| \\le C_\\kappa' c^{-1/4}$.
(The rate $c^{-3/4}$ stated in the lemma
follows if we do not apply Cauchy--Schwarz
but instead bound $\\sup_n$ directly:
$\\sup_{\\|f\\|=1}\\|(\\Hk-\\Hc)f\\|
\\le \\sup_n \\|(\\Hk-\\Hc)\\phi_n^{(c)}\\|
\\le C_\\kappa c^{-3/4}$
by \\eqref{F5} and the identity
$\\|(\\Hk-\\Hc)f\\|^2
= \\sum_n |a_n|^2
  \\|(\\Hk-\\Hc)\\phi_n^{(c)}\\|^2
+ \\text{cross terms}$.
The cross terms vanish only if
$\\{\\phi_n^{(c)}\\}$ diagonalizes $\\Hk$,
which it does not in general.
We therefore state the lemma with
the HS-based rate $c^{-1/4}$,
which is unconditionally valid.)
\\end{proof}

\\begin{lemma}[Operator-norm rate,
  self-contained proof]
\\label{lem:opnorm-self-contained}
For all $c \\ge c_0(\\kappa)$:
\\[
  \\|\\Hc - \\Hstrong\\|_{L^2 \\to L^2}
  \\le C_\\kappa\\, c^{-1/4}.
\\]
\\end{lemma}

\\begin{proof}
Since $\\Hstrong - \\Hc$ is positive
(by \\eqref{F2} and
Corollary~\\ref{cor:qlim-op}):
\\begin{align*}
  \\|\\Hstrong - \\Hc\\|
  &= \\sup_{\\|f\\|=1}
     \\ip{(\\Hstrong - \\Hc)f}{f} \\\\
  &= \\sup_{\\|f\\|=1}
     [\\qlim(f,f) - \\qc(f,f)]
     \\tag{Cor.~\\ref{cor:qlim-op}}\\\\
  &= \\sup_{\\|f\\|=1}
     \\sup_{k > c}
     \\ip{(\\Hk - \\Hc)f}{f}
     \\tag{\\eqref{F2}: $\\qlim = \\sup_k\\qk$}\\\\
  &\\le \\sup_{k > c}
     \\|\\Hk - \\Hc\\|
     \\tag{Cauchy--Schwarz}\\\\
  &\\le C_\\kappa c^{-1/4}.
     \\tag{Lem.~\\ref{lem:finite-opnorm}}
\\end{align*}
\\end{proof}

\\begin{remark}[What this proof uses]
\\label{rem:opnorm-uses}
The argument uses only:
\\begin{itemize}
\\item Corollary~\\ref{cor:qlim-op}
      (form = operator, no spectral data).
\\item Monotonicity \\eqref{F2}.
\\item Lemma~\\ref{lem:finite-opnorm},
      which uses the eigenbasis
      $\\{\\phi_n^{(c)}\\}$ of $\\Hc$
      for \\emph{fixed} $c$ only,
      and per-vector rates \\eqref{F5}
      from Paper~VIII, Thm.~C
      (involving only $\\Hc$ and $\\Hk$,
      two concrete operators for fixed
      $c \\le k$).
\\end{itemize}
At no point are $\\lambda_n^{(\\infty)}$,
$\\Phi_n^{(\\infty)}$, or $\\Hstrong$
introduced inside the bound.
\\end{remark}

%%═══════════════════════════════════════════════════════════════════
\\section{Limit Form}
\\label{sec:forms}
%%═══════════════════════════════════════════════════════════════════

\\begin{definition}[Limit form]
\\label{def:qlim}
For $f \\in L^2(\\mathbb{R})$,
$\\qlim(f,f) := \\lim_{c\\to\\infty}\\ip{\\Hc f}{f}$
(exists by \\eqref{F2}).
Extend by polarisation to a Hermitian form
$\\qlim : L^2 \\times L^2 \\to \\mathbb{C}$
with $\\Dom(\\qlim) = L^2(\\mathbb{R})$.
\\end{definition}

\\begin{theorem}[Existence and boundedness]
\\label{thm:qlim-exists}
$\\qlim \\ge 0$ is a bounded Hermitian form
on $L^2(\\mathbb{R})$ with
$\\qlim(f,f) \\le \\|f\\|^2$.
\\end{theorem}
\\begin{proof}
Non-negativity and boundedness: \\eqref{F1}.
\\end{proof}

%%═══════════════════════════════════════════════════════════════════
\\section{Mosco Convergence}
\\label{sec:mosco}
%%═══════════════════════════════════════════════════════════════════

\\begin{definition}[Mosco convergence
  \\cite{Mosco1969}]
\\label{def:mosco}
$(\\qc)$ converge to $\\qlim$ in the
sense of Mosco if:
\\begin{enumerate}[\\rm(M1)]
\\item For every $c_n \\to \\infty$ and
      $f_n \\rightharpoonup f$:
      $\\qlim(f,f)
      \\le \\liminf_{n} q_{c_n}(f_n,f_n)$.
\\item For every $f$ there exist
      $c_n \\to \\infty$, $f_n \\to f$
      strongly with
      $q_{c_n}(f_n,f_n) \\to \\qlim(f,f)$.
\\end{enumerate}
\\end{definition}

\\begin{lemma}[Compactness and
  weak continuity]
\\label{lem:hs}
For each fixed $c_0 > 0$:
\\begin{enumerate}[\\rm(a)]
\\item $\\Hc[c_0] \\in \\mathcal{B}_2(L^2)$
      (Hilbert--Schmidt, hence compact).
\\item $f_n \\rightharpoonup f$
      $\\Rightarrow$
      $\\ip{\\Hc[c_0] f_n}{f_n}
      \\to \\ip{\\Hc[c_0] f}{f}$.
\\end{enumerate}
\\end{lemma}
\\begin{proof}
\\textit{(a)} The integral kernel
$K_{c_0}(x,y)
= \\frac{\\sin c_0(x-y)}{\\pi(x-y)}$
has $\\|K_{c_0}\\|_{L^\\infty} \\le c_0/\\pi$,
giving $\\|\\Hc[c_0]\\|_{\\mathrm{HS}} < \\infty$
(\\cite{ReedSimon1975}, Thm.~VI.22).

\\textit{(b)} Write
$\\ip{\\Hc[c_0] f_n}{f_n}
- \\ip{\\Hc[c_0] f}{f}
= I + II$
with
$I := \\ip{\\Hc[c_0](f_n - f)}{f_n}$
and
$II := \\ip{\\Hc[c_0] f}{f_n - f}$.
$II \\to 0$ by weak convergence.
For $I$: compactness gives
$\\Hc[c_0](f_n - f) \\to 0$ strongly;
$\\sup_n\\|f_n\\| < \\infty$
(Banach--Steinhaus),
so $|I| \\to 0$.
\\end{proof}

\\begin{theorem}[Mosco convergence]
\\label{thm:mosco}
$\\qc \\to \\qlim$ in the Mosco sense.
\\end{theorem}
\\begin{proof}
\\textit{(M2).}
Take $f_n := f$, $c_n \\to \\infty$:
$q_{c_n}(f,f) = \\ip{\\Hc[c_n]f}{f}
\\to \\ip{\\Hstrong f}{f} = \\qlim(f,f)$
(Lemma~\\ref{lem:hlim-bounded},
Corollary~\\ref{cor:qlim-op}).

\\textit{(M1).}
Let $c_n \\to \\infty$,
$f_n \\rightharpoonup f$.
Fix $c_0 > 0$.
For $n$ with $c_n \\ge c_0$,
monotonicity \\eqref{F2} gives
$q_{c_n}(f_n,f_n) \\ge q_{c_0}(f_n,f_n)$.
Hence:
\\[
  \\liminf_n q_{c_n}(f_n,f_n)
  \\ge \\liminf_n q_{c_0}(f_n,f_n)
  = q_{c_0}(f,f)
\\]
by Lemma~\\ref{lem:hs}(b)
(fixed $c_0$, weak convergence).
This holds for every fixed $c_0$.
Since $q_{c_0}(f,f)$ is independent
of $n$, taking $\\sup$ over $c_0$:
\\[
  \\liminf_n q_{c_n}(f_n,f_n)
  \\ge \\sup_{c_0>0} q_{c_0}(f,f)
  = \\qlim(f,f).
\\]
(If $A \\ge b$ for every $b \\in S$
then $A \\ge \\sup S$;
no interchange of limits is involved.)
\\end{proof}

%%═══════════════════════════════════════════════════════════════════
\\section{Closedness and Quantitative Rate}
\\label{sec:form-closed}
%%═══════════════════════════════════════════════════════════════════

\\begin{theorem}[Closedness of $\\qlim$]
\\label{thm:form-closed}
$\\qlim$ is a closed form.
\\end{theorem}
\\begin{proof}
Since $\\qlim(f,f) = \\ip{\\Hstrong f}{f}$
and $\\Hstrong \\in \\mathcal{B}(L^2)$,
the form norm
$\\|f\\|_{\\qlim}^2 := \\|f\\|^2 + \\qlim(f,f)$
is equivalent to $\\|f\\|^2$.
Any $\\|\\cdot\\|_{\\qlim}$-Cauchy sequence
is $L^2$-Cauchy, converges in $L^2$,
and continuity of $\\qlim$ gives closedness.
\\end{proof}

\\begin{theorem}[Quantitative form
  convergence]
\\label{thm:form-convergence}
For all $f \\in L^2(\\mathbb{R})$,
$c \\ge c_0(\\kappa)$:
\\[
  0 \\le \\qlim(f,f) - \\qc(f,f)
  \\le C_\\kappa c^{-1/4}\\|f\\|^2.
\\]
\\end{theorem}
\\begin{proof}
Lower bound: \\eqref{F2}.
Upper bound:
$\\qlim(f,f) - \\qc(f,f)
= \\ip{(\\Hstrong - \\Hc)f}{f}
\\le \\|\\Hc - \\Hstrong\\|\\cdot\\|f\\|^2
\\le C_\\kappa c^{-1/4}\\|f\\|^2$
(Corollary~\\ref{cor:qlim-op};
Lemma~\\ref{lem:opnorm-self-contained}).
\\end{proof}

%%═══════════════════════════════════════════════════════════════════
\\section{Riesz--Kato Form Identification}
\\label{sec:friedrichs}
%%═══════════════════════════════════════════════════════════════════

\\begin{remark}[Terminology]
\\label{rem:friedrichs-terminology}
Since $\\Dom(\\qlim) = L^2(\\mathbb{R})$
and $\\qlim$ is bounded, this is not
an extension problem.
The identification below follows directly
from the Riesz representation theorem
for bounded Hermitian forms.
\\end{remark}

\\begin{theorem}[Riesz--Kato identification]
\\label{thm:friedrichs}
$T_{\\qlim} = \\Hstrong$,
where $T_{\\qlim}$ is the unique bounded
self-adjoint operator with
$\\qlim(f,g) = \\ip{T_{\\qlim} f}{g}$
for all $f, g \\in L^2$.
\\end{theorem}
\\begin{proof}
The Riesz representation theorem
gives a unique
$T_{\\qlim} \\in \\mathcal{B}(L^2)$
representing $\\qlim$.
By Corollary~\\ref{cor:qlim-op}:
$\\ip{\\Hstrong f}{g}
= \\qlim(f,g) = \\ip{T_{\\qlim}f}{g}$
for all $f, g$, so
$T_{\\qlim} = \\Hstrong$.
\\end{proof}

%%═══════════════════════════════════════════════════════════════════
\\section{The Bridge Theorem}
\\label{sec:bridge}
%%═══════════════════════════════════════════════════════════════════

This section is the \\emph{only} section
of this paper that imports results from
Paper~IX.

\\begin{enumerate}[(IX.a)]

\\item\\label{IXa}
      \\textit{Spectral-closure operator.}
      Paper~IX, Definition~2.3 defines
      \\[
        \\Hspectral f
        := \\sum_{n=0}^\\infty
           \\lambda_n^{(\\infty)}
           \\ip{f}{\\Phi_n^{(\\infty)}}
           \\Phi_n^{(\\infty)}
      \\]
      on the domain
      $\\Dom(\\Hspectral) = \\{f \\in L^2 :
      \\sum_n (\\lambda_n^{(\\infty)})^2
      |\\ip{f}{\\Phi_n^{(\\infty)}}|^2
      < \\infty\\}$.
      This is symmetric, densely defined,
      and closable
      (Paper~IX, Thm.~2.5, Thm.~3.2).

\\item\\label{IXb}
      \\textit{Closure is bounded.}
      $\\lambda_n^{(\\infty)} \\in [0,1]$
      for all $n$, so
      $\\|\\Hspectral f\\|^2
      \\le \\|f\\|^2$
      on $\\Dom(\\Hspectral)$.
      The closure $\\overline{\\Hspectral}
      \\in \\mathcal{B}(L^2)$.

\\item\\label{IXc}
      \\textit{Agreement on $\\mathcal{D}$.}
      For $f \\in \\mathcal{D}$:
      $\\Hc f \\to \\Hstrong f$
      (definition of $\\Hstrong$)
      and
      $\\Hc f \\to \\overline{\\Hspectral} f$
      (Paper~IX, Remark~2.6).
      Uniqueness of $L^2$-limits:
      $\\Hstrong f
      = \\overline{\\Hspectral} f$
      on $\\mathcal{D}$.

\\end{enumerate}

\\begin{theorem}[Bridge theorem]
\\label{thm:bridge}
$\\Hstrong = \\overline{\\Hspectral}$
as bounded operators on $L^2(\\mathbb{R})$.
\\end{theorem}

\\begin{proof}
Both operators are bounded
(\\eqref{IXb} and Lemma~\\ref{lem:hlim-bounded}).
They agree on the dense set $\\mathcal{D}$
(\\eqref{IXc}).
Hence they agree on all of $L^2$.
\\end{proof}

\\begin{remark}[What the bridge achieves]
\\label{rem:bridge}
After Theorem~\\ref{thm:bridge}, we write
$\\Hlim := \\Hstrong = \\overline{\\Hspectral}$.
Paper~X builds $\\Hlim$ as a bounded operator;
Paper~IX builds it as a spectral closure.
Both are the same object.
\\end{remark}

\\begin{remark}[$\\Hlim$ is not a
  multiplication operator]
\\label{rem:hlim-not-mult}
By Theorem~\\ref{thm:bridge},
$\\sigma(\\Hlim) \\subset
\\overline{\\{\\lambda_n^{(\\infty)} : n \\ge 0\\}}$.
The Landau--Pollak transition
(\\cite{Slepian1961}) gives values
accumulating densely in $(0,1)$,
so $\\sigma(\\Hlim) \\not\\subset \\{0,1\\}$,
distinguishing $\\Hlim$ from
$M_{\\chi_{[-1,1]}}$.
\\end{remark}

%%═══════════════════════════════════════════════════════════════════
\\section{Spectral Consequences}
\\label{sec:consequences}
%%═══════════════════════════════════════════════════════════════════

\\begin{corollary}[Spectral inclusion]
\\label{cor:spec}
$\\sigma(\\Hlim) \\subset [0,1]$.
\\end{corollary}

\\begin{corollary}[Strong resolvent
  convergence]
\\label{cor:resolvent}
For every $z \\in \\mathbb{C} \\setminus [0,1]$:
$(\\Hc - z)^{-1}
\\xrightarrow{\\mathrm{s}}
(\\Hlim - z)^{-1}$.
\\end{corollary}
\\begin{proof}
Mosco convergence
(Theorem~\\ref{thm:mosco}),
Corollary~\\ref{cor:spec},
and \\cite{KuwaeShioya2003}, Thm.~2.4.
\\end{proof}

\\begin{remark}[Unconditional results]
\\label{rem:unconditional}
Without any hypothesis:
$\\Hlim \\in \\mathcal{B}(L^2)$,
$\\Hlim = \\Hlim^* \\ge 0$,
$\\sigma(\\Hlim) \\subset [0,1]$,
$T_{\\qlim} = \\Hlim$,
$(\\Hc-z)^{-1} \\xrightarrow{s} (\\Hlim-z)^{-1}$,
$\\qc \\xrightarrow{\\mathrm{Mosco}} \\qlim$
at rate $O(c^{-1/4})$,
and $\\Hstrong = \\overline{\\Hspectral}$.
\\end{remark}

\\begin{corollary}[Conditional RH]
\\label{cor:rh}
Conditionally on
Hypothesis~CCM (\\cite{CCM2025}, Thm.~1.3)
and Paper~IX, Hypothesis~4.1,
every non-trivial zero $\\rho$ of $\\zeta(s)$
satisfies $\\mathrm{Re}(\\rho) = \\tfrac{1}{2}$.

\\textit{No unconditional claim about
zeta zeros is made in this paper.}
\\end{corollary}
\\begin{proof}
Under Paper~IX, Hyp.~4.1:
$\\Hlim$ is essentially self-adjoint
(Paper~IX, Cor.~5.2),
so $\\sigma_{\\mathrm{app}}(\\Hlim)
= \\sigma(\\Hlim)$.
Under CCM:
$\\rho \\in \\sigma_{\\mathrm{app}}(\\Hlim)$
(Paper~VIII, Thm.~5.2),
hence $\\rho \\in \\sigma(\\Hlim)
\\subset \\mathbb{R}$,
and the CCM embedding forces
$\\mathrm{Re}(\\rho) = 1/2$.
\\end{proof}

%%═══════════════════════════════════════════════════════════════════
\\section{Open Problems}
\\label{sec:problems}
%%═══════════════════════════════════════════════════════════════════

\\begin{problem}[Completeness of
  $\\{\\Phi_n^{(\\infty)}\\}$]
\\label{prob:completeness}
Prove Paper~IX, Hypothesis~4.1.
By the bridge theorem, this is equivalent
to: $\\overline{\\Hspectral}$ has no
deficiency vectors in $L^2$.
\\end{problem}

\\begin{problem}[Sharp rate]
\\label{prob:rate}
Is $c^{-1/4}$ sharp in
Theorem~\\ref{thm:form-convergence}?
The HS-norm argument in
Lemma~\\ref{lem:finite-opnorm}
yields $c^{-1/4}$.
A direct sup-over-eigenvectors argument
would yield $c^{-3/4}$ (if
the cross terms in the eigenbasis of $\\Hc$
could be controlled), but this remains open.
\\end{problem}

\\begin{problem}[Quantitative resolvent]
\\label{prob:resolvent}
Find an explicit rate $F(c,z) \\to 0$
with
$\\|(\\Hc-z)^{-1}-(\\Hlim-z)^{-1}\\|
\\le F(c,z)$.
\\end{problem}

\\begin{problem}[Spectrum of $\\Hlim$]
\\label{prob:spectrum}
Is $\\sigma(\\Hlim) = [0,1]$?
\\end{problem}

%%%─── Bibliography
\\begin{thebibliography}{99}

\\bibitem{PaperV}
[Author],
\\textit{Effective bandwidth and
$L^2$-concentration of PSWFs},
Paper~V, preprint 2025.

\\bibitem{PaperVI}
[Author],
\\textit{Pointwise Mellin decay,
log-free $L^2$ tails},
Paper~VI, preprint 2025.

\\bibitem{PaperVII}
[Author],
\\textit{Composite asymptotics and
$c^{-1/2}$ upper bound},
Paper~VII, preprint 2025.

\\bibitem{PaperVIII}
[Author],
\\textit{Non-cancellation, sharp
$c^{-1/2}$ lower bound},
Paper~VIII, preprint 2026.

\\bibitem{PaperIX}
[Author],
\\textit{Domain, deficiency indices,
self-adjointness},
Paper~IX, preprint 2026.

\\bibitem{Kato1966}
T.~Kato,
\\textit{Perturbation Theory for
Linear Operators},
Springer, 1966.

\\bibitem{ReedSimon1975}
M.~Reed and B.~Simon,
\\textit{Methods of Modern
Mathematical Physics, Vol.~II},
Academic Press, 1975.

\\bibitem{Mosco1969}
U.~Mosco,
\\textit{Convergence of convex
sets and of solutions of
variational inequalities},
Adv.\\ Math.\\ \\textbf{3} (1969),
510--585.

\\bibitem{KuwaeShioya2003}
K.~Kuwae and T.~Shioya,
\\textit{Convergence of spectral
structures},
Comm.\\ Anal.\\ Geom.\\ \\textbf{11}
(2003), 599--673.

\\bibitem{Slepian1961}
D.~Slepian and H.~O.~Pollak,
\\textit{Prolate spheroidal
wave functions, Fourier
analysis and uncertainty---I},
Bell Syst.\\ Tech.\\ J.\\
\\textbf{40} (1961), 43--63.

\\bibitem{CCM2025}
A.~Connes, C.~Consani,
and H.~Moscovici,
\\textit{Zeta Spectral Triples},
arXiv:2511.22755, 2025.

\\bibitem{Rudin1991}
W.~Rudin,
\\textit{Functional Analysis},
2nd ed., McGraw-Hill, 1991.

\\bibitem{Conway1990}
J.~B.~Conway,
\\textit{A Course in Functional
Analysis},
2nd ed., Springer, 1990.

\\end{thebibliography}

\\end{document}
