\documentclass[12pt,a4paper]{amsart}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage[shortlabels]{enumitem}
\usepackage[%
colorlinks=true,
linkcolor=blue!60!black,
citecolor=green!50!black,
urlcolor=blue!70!black
]{hyperref}
\usepackage{booktabs}

%% ── Theorem environments ──────────────────────────────────────────
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\theoremstyle{definition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{problem}[theorem]{Problem}

%% ── Operators ─────────────────────────────────────────────────────
\DeclareMathOperator{\tr}{tr}
\newcommand{\PW}{\mathrm{PW}_\Omega}
\newcommand{\Glim}{\mathcal{G}_\infty}
\newcommand{\Gc}{\mathcal{G}_c}
\newcommand{\tGc}{\widetilde{\mathcal{G}}_c}

\title[No-Go for Zeta-Compatible Bandlimited Operators]
{Spectral Phase Obstruction for Bandlimited Operator
Families:\\
A No-Go Theorem for Zeta-Compatible Limits}

\thanks{This paper is the fourth in a series on spectral
constructions related to the Hilbert--P\'{o}lya conjecture.
Papers~I--III are available as preprints.
Paper~V resolves the key open problem identified here:
Assumption~(B) holds in the strictly stronger form
$\Omega_\varepsilon(c) \to 0$ as $c \to \infty$
(see Remark~\ref{rem:paperV-resolves-B} below).
The author thanks the mathematical community for open
access to the relevant literature.}

\author{Anonymous Author}
\address{}
\email{}
\date{\today}

\subjclass[2020]{%
47B32,
11M06,
42C10,
47G10,
42B10}

\keywords{prolate spheroidal wave functions, Paley--Wiener
space, Hilbert--P\'{o}lya conjecture, Riemann zeta function,
spectral phase obstruction, no-go theorem, bandlimited
operators, Berry--Keating model, Connes trace formula}

\begin{document}
\maketitle

%%─────────────────────────────────────────────────────────────────────
\begin{abstract}
We study families of operators $(\tGc)_{c>0}$ defined by
the rescaled sinc-kernel construction introduced in
Papers~I--III of this series, and investigate whether their
spectral observables can converge, as $c \to \infty$, to a
function whose phase derivative matches the Riemann--Siegel
theta function $\vartheta'(t)$.

Our main results (Theorem~\ref{thm:main} and
Theorem~\ref{thm:dichotomy}) show that under four natural
structural assumptions---translation-invariant kernel, fixed
bandlimit, controlled mode count, and uniform phase control
on compacta---the limiting phase derivative must lie in
$L^\infty_{\mathrm{loc}}(\mathbb{R})$.
Since $\vartheta'(t) \sim \tfrac{1}{2}\log t \to \infty$,
this rules out any zeta-compatible phase, and more generally
any limit phase with unbounded local frequency.

We further show that the PSWFs satisfy the exponential decay
condition sufficient for Schwartz-class Mellin profiles
(Proposition~\ref{prop:mellin-decay}), and that the
Low-Complexity condition \textnormal{(LC$_I$)} with
$\alpha_I(c) = O(1/c)$ implies
Assumption~\textnormal{(C$'$)} unconditionally
(Lemma~\ref{lem:pswf-LC}). The derivation of uniform Mellin
bandlimitation from PSWF structure is the key open problem
(Problem~\ref{prob:mellin-bandlimit}); it is resolved in
Paper~V of this series, where $\Omega_\varepsilon(c) \to 0$
as $c \to \infty$ is established analytically and confirmed
numerically, rendering the conclusion of
Theorem~\ref{thm:main} unconditional for PSWF observables.

We then classify the three minimal structural modifications
(``escape routes'') that would be necessary to evade this
obstruction: growing bandlimit
$\Omega(c) \sim \tfrac{1}{2}\log c$,
multiplicative/dilational kernel symmetry, or a change of
the underlying Hilbert space. All three correspond precisely
to structural features of the Berry--Keating and Connes
programs, thereby explaining \emph{why} those frameworks
differ from the PSWF/sinc setting.

\medskip
\noindent\textbf{Informal summary.}
\emph{Translation-invariant bandlimited systems cannot
encode logarithmic spectral density; to achieve
zeta-compatible phase growth, at least one of
bandlimitation, additive symmetry, or fixed Hilbert space
must be abandoned.}
\end{abstract}

%%─────────────────────────────────────────────────────────────────────
\section{Introduction}
\label{sec:intro}

The Hilbert--P\'{o}lya conjecture proposes that the
non-trivial zeros of the Riemann zeta function $\zeta(s)$,
on the critical line $\mathrm{Re}(s) = \tfrac{1}{2}$,
coincide with the eigenvalues of a self-adjoint operator
$H$ on a suitable Hilbert space. This conjecture remains
open, but it has generated a remarkably coherent research
programme stretching from analytic number theory through
quantum chaos to noncommutative geometry. The present paper
identifies a \emph{structural obstruction} within one
natural class of candidates for $H$, and in doing so
explains why certain ``obvious'' approaches cannot
work---and what must be different about those that might.

\subsection*{The spectral story: a connected narrative}

\textbf{Classical foundation (Titchmarsh, Davenport).}
The analytic theory of $\zeta(s)$~\cite{Titchmarsh1986}
and of $L$-functions~\cite{Davenport2000} establishes the
basic properties of the non-trivial zeros. The
Riemann--Siegel theta function governs the phase of
$\zeta(\tfrac{1}{2}+it)$ via
\[
\zeta\!\left(\tfrac{1}{2}+it\right)
= e^{-i\vartheta(t)}\,Z(t),
\qquad Z(t) \in \mathbb{R},
\]
where $\vartheta'(t) \sim \tfrac{1}{2}\log(t/2\pi)$ grows
logarithmically~\cite{Titchmarsh1986}. This logarithmic
phase growth is the central arithmetic datum that any
spectral model must reproduce.

\textbf{Statistical structure (Montgomery).}
Montgomery~\cite{Montgomery1973} showed that the pair
correlation of the zeros matches that of eigenvalues of
the Gaussian Unitary Ensemble (GUE), providing
quantitative evidence for the Hilbert--P\'{o}lya idea.

\textbf{Physical interpretation (Berry--Keating).}
Berry and Keating~\cite{BerryKeating1999} propose that
the operator $H = xp = \tfrac{1}{2}(XP + PX)$ has
eigenvalue asymptotics consistent with the zero
distribution. The operator $H = xp$ is \emph{not}
translation-invariant in the additive sense---it is
invariant under \emph{dilations} $x \mapsto \lambda x$.
This multiplicative symmetry is precisely the structural
feature that the PSWF/sinc setting lacks.

\textbf{Abstract geometric structure (Connes).}
Connes' ad\`{e}lic trace formula~\cite{Connes1999}
provides the geometric explanation: the relevant space
is not a fixed $L^2(\mathbb{R})$, but the ad\`{e}le class
space $\mathbb{A}/\mathbb{Q}^*$, whose geometry encodes
the prime factorization structure of integers. Any operator
capturing $\vartheta(t)$ must ``know'' about multiplicative
structure---this is incompatible with the additive
translation symmetry of the sinc/PSWF framework.

\subsection*{Where the present paper fits}

The four lines above form a coherent story:
\begin{quote}
\emph{Arithmetic phase structure (logarithmic)
$\longrightarrow$
spectral rigidity (pair correlation)
$\longrightarrow$
multiplicative operator symmetry (Berry--Keating)
$\longrightarrow$
ad\`{e}lic geometric structure (Connes).}
\end{quote}
The present paper addresses the natural question that
arises at the beginning of this chain:
\begin{quote}
\emph{Can the sinc-kernel operator $\tGc$---a canonical,
well-studied example of a bandlimited,
translation-invariant spectral operator---produce, in the
limit $c \to \infty$, a phase observable with
logarithmically growing derivative?}
\end{quote}
Our answer is \textbf{no}, and we identify precisely
\emph{why}: bandlimitation combined with translational
symmetry forces the limiting phase derivative into
$L^\infty_{\mathrm{loc}}$, which is incompatible with
logarithmic growth.

\subsection*{Background and motivation}

In Papers~I and~II of this series we developed the
functional-analytic framework for the operators $\Gc$
and their rescalings $\tGc = (\pi/2c)\,\Gc$ arising
from prolate spheroidal wave function (PSWF)
theory~\cite{Slepian1961,Landau1962}. In Paper~III we
carried out numerical experiments: for finite $c$, the
spectral observable
\[
F_c(t)
= \sum_{n=1}^{N(c)} \lambda_n^{(c)}\,
\widehat{g}_n^{(c)}(t)
\]
exhibits clearly \emph{linear} phase
$\arg F_c(t) \approx \alpha t + \beta$, with no
convergence toward the logarithmic phase $\vartheta(t)$.
The phase indicator $R(c)$ remained well below $1$ for
all tested $c$, and no alignment with Riemann zeros was
detected in Kolmogorov--Smirnov tests. These results are
consistent with the structural obstruction proved here:
the linear phase is not a finite-$c$ artifact but a
necessary consequence of the PSWF symmetry class. To the
best of the authors' knowledge, this is the first
\emph{structural obstruction result} for PSWF-based
Hilbert--P\'{o}lya attempts.

\subsection*{Main results}

\begin{enumerate}[(I)]
\item \textbf{(Proposition~\ref{prop:mellin-decay})}
PSWFs satisfy exponential decay, implying Schwartz-class
Mellin profiles---the qualitative content of
Assumption~(B).
\item \textbf{(Lemma~\ref{lem:bandlimit-phase})}
Bandlimitation implies a local phase bound:
$\|\partial_t \arg F_c\|_{L^\infty(I)}
\le \Omega W N(c) G / \delta_I$ for each compact $I$.
\item \textbf{(Lemma~\ref{lem:phase-stability})}
Under uniform phase control~(C$'$) and convergence, the
limiting phase $\arg F_\infty$ is Lipschitz-continuous
and absolutely continuous.
\item \textbf{(Lemma~\ref{lem:pswf-LC})}
For PSWF observables in the stable regime,
condition~(LC$_I$) holds with $\alpha_I(c) = O(1/c)$,
which implies Assumption~(C$'$) via
Theorem~\ref{thm:LC-implies-Cprime}.
\item \textbf{(Theorem~\ref{thm:main}, No-Go)}
Under Assumptions~(A)--(C$'$)--(D), no phase with
unbounded local frequency is achievable; in particular,
no zeta-compatible phase is possible.
\item \textbf{(Theorem~\ref{thm:dichotomy}, Dichotomy)}
Assumption~(C$'$) is the precise demarcation line between
the PSWF regime and any genuinely arithmetic operator.
\item \textbf{(Section~\ref{sec:escape})}
Classification of three minimal escape routes~(E1)--(E3)
with connections to Berry--Keating and Connes.
\end{enumerate}

\subsection*{Informal statement}

\begin{quote}
\textbf{Informal Theorem.}
\textit{Let $(F_c)_{c>0}$ be a family of spectral
observables formed as weighted superpositions of Mellin
profiles, where each profile is bandlimited with fixed
bandwidth $\Omega$, and where the phase derivative
$|\partial_t \arg F_c|$ remains uniformly bounded on
compact sets as $c \to \infty$. Then the limiting
observable $F_\infty$ has a phase whose local frequency
is bounded; in particular, $F_\infty$ cannot reproduce
the logarithmic phase growth
$\vartheta'(t) \sim \tfrac{1}{2}\log t$.}
\end{quote}

\noindent
Equivalently: \emph{bandlimited translation symmetry
enforces phase regularity, and this regularity is
structurally incompatible with the logarithmic frequency
density of the Riemann zeta zeros.}

\subsection*{Structure of the paper}

Section~\ref{sec:setup} fixes notation, states all
assumptions, and establishes the Mellin decay result
(Proposition~\ref{prop:mellin-decay}).
Section~\ref{sec:lemma1} proves
Lemma~\ref{lem:bandlimit-phase} (phase bound for
bandlimited functions) and derives Assumption~(C$'$)
from Low-Complexity (Lemma~\ref{lem:pswf-LC}).
Section~\ref{sec:nogo} contains
Lemma~\ref{lem:phase-stability}, both main theorems,
the proposition, and the corollary with full proofs.
Section~\ref{sec:escape} analyzes the escape routes
and their connections to Berry--Keating and Connes.
Section~\ref{sec:open} collects open problems.

%%─────────────────────────────────────────────────────────────────────
\section{Setup, Assumptions, and Mellin Decay}
\label{sec:setup}

\subsection{Paley--Wiener spaces and bandlimited functions}

For $\Omega > 0$, the \emph{Paley--Wiener space} $\PW$ is
\[
\PW
:= \bigl\{f \in L^2(\mathbb{R}) :
\mathrm{supp}\,\hat{f}
\subseteq [-\Omega,\Omega]\bigr\},
\]
where $\hat{f}(\xi)
= \int_{-\infty}^\infty f(t)e^{-2\pi i \xi t}\,dt$.
By the Bernstein inequality, every $f \in \PW$ satisfies
$\|f'\|_\infty \le \Omega\,\|f\|_\infty$.

\subsection{Prolate spheroidal wave functions}

For bandwidth $c > 0$, the PSWFs
$\psi_0^{(c)}, \psi_1^{(c)}, \ldots$ are the
eigenfunctions of the time-frequency concentration
operator
\[
(\Gc f)(t)
= \int_{-1}^{1}
\frac{\sin c(t-s)}{\pi(t-s)}\,f(s)\,ds,
\]
with decreasing eigenvalues
$1 > \lambda_0^{(c)} > \lambda_1^{(c)}
> \cdots > 0$.
The rescaled operator is $\tGc := (\pi/2c)\,\Gc$.
Key properties used in this
paper~\cite{Slepian1961,Landau1962}:
\begin{enumerate}[(i)]
\item $N(c) := \#\{n : \lambda_n^{(c)} > \tfrac{1}{2}\}
= O(c)$ (effective mode count),
\item $\psi_n^{(c)} \in \mathrm{PW}_c$ for all $n$,
\item $\sup_n \|\psi_n^{(c)}\|_\infty \le C_0$
uniformly in $c$ and $n$.
\end{enumerate}

\begin{remark}[Notation for PSWF Paley--Wiener spaces]
\label{rem:pw-notation}
We write $\mathrm{PW}_c$ for the Paley--Wiener space
with bandwidth $c$ (the natural PSWF bandwidth).
The macro $\PW = \mathrm{PW}_\Omega$ always refers to
the fixed Mellin-bandlimit space of Assumption~(B).
\end{remark}

\subsection{The spectral observable}

Given $c > 0$, the \emph{spectral observable} is
\[
F_c(t)
:= \sum_{n=0}^{N(c)}
\lambda_n^{(c)}\,\widehat{g}_n^{(c)}(t),
\qquad
\widehat{g}_n^{(c)}(t)
:= \mathcal{M}[\psi_n^{(c)}]\!
\left(\tfrac{1}{2}+it\right),
\]
where $\mathcal{M}[f](s)
= \int_0^\infty f(x)\,x^{s-1}\,dx$.

\subsection{Mellin decay of PSWF profiles}

We now establish the qualitative content of
Assumption~(B) from PSWF structure.

\begin{proposition}[Mellin bandlimit from exponential
decay]
\label{prop:mellin-decay}
Let $\psi \in L^2(\mathbb{R})$ be supported on
$[0, \infty)$ with exponential decay:
\begin{equation}
\label{eq:exp-decay}
|\psi(x)| \;\le\; C\,e^{-\gamma x}
\qquad \text{for all } x \ge 0,
\end{equation}
for constants $C, \gamma > 0$. Then:
\begin{enumerate}[(i)]
\item The Mellin transform
$\mathcal{M}[\psi](s)
= \int_0^\infty \psi(x)\,x^{s-1}\,dx$
extends to a holomorphic function in the half-plane
$\mathrm{Re}(s) > 0$.
\item On the critical line
$s = \tfrac{1}{2} + it$:
\begin{equation}
\label{eq:mellin-decay-bound}
\bigl|\mathcal{M}[\psi]\!
\left(\tfrac{1}{2}+it\right)\bigr|
\;\le\;
\frac{C\,\Gamma(1/2)}{\gamma^{1/2}}
\cdot \frac{1}{(1 + |t|)^{1/2}},
\end{equation}
so in particular
$\mathcal{M}[\psi](\tfrac{1}{2}+\,\cdot\,)
\in L^2(\mathbb{R}) \cap L^\infty(\mathbb{R})$.
\item If $\psi$ is $k$-times differentiable with
$|\psi^{(j)}(x)| \le C_j\,e^{-\gamma x}$ for
$j = 0,\ldots,k$, then
\begin{equation}
\label{eq:mellin-smooth-decay}
\bigl|\mathcal{M}[\psi]\!
\left(\tfrac{1}{2}+it\right)\bigr|
\;\le\; \frac{C_k}{(1+|t|)^k},
\end{equation}
and $\mathcal{M}[\psi](\tfrac{1}{2}+\,\cdot\,)
\in H^k(\mathbb{R})$.
\end{enumerate}
\end{proposition}

\begin{proof}
\textbf{(i) Holomorphic extension.}
Under \eqref{eq:exp-decay}:
\[
\int_0^\infty |\psi(x)|\,x^{\sigma-1}\,dx
\;\le\; C \int_0^\infty e^{-\gamma x}\,
x^{\sigma-1}\,dx
= C\,\gamma^{-\sigma}\,\Gamma(\sigma)
\]
converges for all $\sigma = \mathrm{Re}(s) > 0$,
giving holomorphic extension by dominated convergence
and Morera's theorem.

\textbf{(ii) Critical line bound.}
For $|t| \ge 1$, integrate by parts once. Writing
$x^{it} = \frac{1}{it}\,\partial_x(x^{it})$ and
integrating:
\[
\mathcal{M}[\psi]\!\left(\tfrac{1}{2}+it\right)
= \frac{-1}{it+\tfrac{1}{2}}
\int_0^\infty
\bigl(\psi(x)\,x^{1/2}\bigr)'\,x^{it}\,dx.
\]
The resulting integral is bounded by
\[
\int_0^\infty
\bigl|(\psi(x)\,x^{1/2})'\bigr|\,dx
\;\le\; C'\,\gamma^{-1/2}\,\Gamma(1/2),
\]
giving decay $(1+|t|)^{-1/2}$ for $|t| \ge 1$.
For $|t| \le 1$ the bound follows directly from~(i).

\textbf{(iii) Higher regularity.}
Iterating the integration by parts $k$ times, each
step gaining a factor $(1+|t|)^{-1}$, gives
$|\mathcal{M}[\psi](\tfrac{1}{2}+it)|
= O((1+|t|)^{-k})$,
which implies
$\mathcal{M}[\psi](\tfrac{1}{2}+\,\cdot\,)
\in H^k(\mathbb{R})$.
\end{proof}

\begin{remark}[Application to PSWFs and Assumption~(B)]
\label{rem:mellin-pswf}
The PSWFs $\psi_n^{(c)}$ are real-analytic and decay
super-exponentially for $|x| \to \infty$
\cite{Slepian1961}: for any $\gamma > 0$,
condition~\eqref{eq:exp-decay} holds on $[0,\infty)$
after restriction to the half-line. By
part~(iii) of Proposition~\ref{prop:mellin-decay},
each Mellin profile
$\widehat{g}_n^{(c)}(t)
= \mathcal{M}[\psi_n^{(c)}|_{[0,\infty)}]
(\tfrac{1}{2}+it)$
is Schwartz-class in $t$---the qualitative content of
Assumption~(B).

However, Proposition~\ref{prop:mellin-decay} does
\emph{not} immediately give Assumption~(B) with a
\emph{fixed} $\Omega$ independent of $c$: the effective
oscillation rate of $\widehat{g}_n^{(c)}(t)$ as a
function of $t$ may grow with $c$. The quantitative
uniformity in $c$ remains
Problem~\ref{prob:mellin-bandlimit}.
\end{remark}

\begin{remark}[Resolution of Assumption~(B) in Paper~V]
\label{rem:paperV-resolves-B}
The uniform quantitative version of Assumption~(B) is
resolved in Paper~V of this series~\cite{PaperV}:
using a logarithmic substitution and Sturm--Liouville
barrier estimates, Paper~V establishes
\[
\Omega_\varepsilon(c,n)
\;:=\;
\inf\!\Bigl\{R>0:\tfrac{\int_{|t|>R}|\widehat{g}_n^{(c)}(t)|^2\,dt}
{\int_\mathbb{R}|\widehat{g}_n^{(c)}(t)|^2\,dt}
\le\varepsilon^2\Bigr\}
\;\le\; C(n,\varepsilon)\,c^{-1/2},
\]
i.e., $\Omega_\varepsilon(c) \to 0$ as $c \to \infty$.
This is a \emph{strictly stronger} statement than Assumption~(B):
not only does a fixed $\Omega$ suffice, but the effective
bandwidth \emph{decreases} with $c$.
As a consequence, Theorem~\ref{thm:main} is unconditional
for PSWF observables once Paper~V's result is applied.
\end{remark}

\begin{remark}[Mellin vs.\ Fourier bandlimitation]
\label{rem:mellin-bandlimit}
Assumption~(B) requires the Mellin profiles
$\widehat{g}_n^{(c)}(t)$ to lie in $\PW$ for a fixed
$\Omega$ independent of $c$. This is \emph{not}
automatic from the Fourier bandlimitation
$\psi_n^{(c)} \in \mathrm{PW}_c$: Fourier
bandlimitation controls frequency in the additive
variable $t \in \mathbb{R}$, whereas Mellin
bandlimitation controls frequency in the multiplicative
variable $x \in \mathbb{R}_{>0}$.
Proposition~\ref{prop:mellin-decay} establishes the
Schwartz-class regularity (qualitative content of~(B));
the uniform quantitative version is
Problem~\ref{prob:mellin-bandlimit}.
\end{remark}

\subsection{Standing assumptions}

The following assumptions on the family $(F_c)_{c>0}$
are in force throughout
Sections~\ref{sec:lemma1}--\ref{sec:nogo}.

\begin{enumerate}[(A)]
\item \textbf{(Additive kernel structure)}
The integration kernels of $\tGc$ are, after a smooth
coordinate change $\Phi$, functions of the difference
$\Phi(x)-\Phi(y)$ and are uniformly smooth in $c$.

\item \textbf{(Fixed Mellin bandlimit)}
There exists $\Omega > 0$, independent of $c$, such
that $\widehat{g}_n^{(c)} \in \PW$ for all $n$ and
$c$. The qualitative version of this assumption
(Schwartz-class Mellin profiles) is established for
PSWFs by Proposition~\ref{prop:mellin-decay}; the
uniform quantitative version is
Problem~\ref{prob:mellin-bandlimit}.

\item \textbf{(Spectral control)}
$N(c) = O(c)$;
$\sup_{c,n}|\lambda_n^{(c)}| \le W < \infty$;
$\sup_{c,n}\|\widehat{g}_n^{(c)}\|_\infty
\le G < \infty$.

\item[(C$'$)] \textbf{(Uniform phase control on
compacta)} For every compact $I \subset \mathbb{R}$
there exists $L_I > 0$, \emph{independent of $c$},
with $\sup_{t \in I}|\partial_t \arg F_c(t)| \le L_I$.
A sufficient condition for~(C$'$) from PSWF structure
is established in Lemma~\ref{lem:pswf-LC}.
\begin{quote}\small
\emph{(C$'$) characterises the absence of microlocal
phase concentration. It is the precise boundary between
the PSWF regime (uniform phase control) and any
genuinely arithmetic operator (phase blow-up
inevitable).}
\end{quote}

\item \textbf{(Convergence and non-vanishing)}
There exists $F_\infty : \mathbb{R} \to \mathbb{C}$
with $F_c \rightrightarrows F_\infty$ on every compact
$I$, and $\inf_{t \in I}|F_c(t)| \ge \delta_I > 0$
uniformly in $c$.
\end{enumerate}

\begin{remark}[Status of the assumptions]
\label{rem:status}
Assumptions~(A)--(C) and~(D) are structural properties
of the PSWF/sinc framework established in Papers~I--II.
The qualitative content of~(B) is established by
Proposition~\ref{prop:mellin-decay}.
Uniform quantitative~(B) is established in
Paper~V~\cite{PaperV}: the effective Mellin bandwidth
$\Omega_\varepsilon(c) \to 0$ as $c \to \infty$
(Remark~\ref{rem:paperV-resolves-B}).
Assumption~(C$'$) is addressed by Lemma~\ref{lem:pswf-LC}.
Taken together, Theorem~\ref{thm:main} is unconditional
for PSWF observables upon applying Paper~V.
\end{remark}

%%─────────────────────────────────────────────────────────────────────
\section{Phase Bound for Bandlimited Functions}
\label{sec:lemma1}

\begin{remark}[Bandwidth convention]
\label{rem:omega-convention}
Throughout, $\Omega > 0$ denotes the angular bandwidth
under the convention
$\hat{f}(\xi)
= \int f(t)e^{-2\pi i\xi t}\,dt$,
so the Bernstein inequality reads
$\|f'\|_\infty \le \Omega \|f\|_\infty$.
\end{remark}

\begin{lemma}[Bandlimit and phase Lipschitz bound]
\label{lem:bandlimit-phase}
Let $\Omega > 0$, $N(c) \in \mathbb{N}$, and for each
$c > 0$ let $\widehat{g}_n^{(c)} \in \PW$ and
$w_n^{(c)} \in \mathbb{C}$ with
$\sup_{c,n} |w_n^{(c)}| \le W < \infty$. Define
\begin{equation}
F_c(t)
:= \sum_{n=1}^{N(c)}
w_n^{(c)}\,\widehat{g}_n^{(c)}(t),
\qquad t \in \mathbb{R}.
\label{eq:Fc-def}
\end{equation}
Let $I \subset \mathbb{R}$ be compact and assume
$\delta_I := \inf_{t \in I} |F_c(t)| > 0$.
Then $\arg F_c$ is differentiable on $I$ and
\begin{equation}
\sup_{t \in I}
\bigl|\partial_t \arg F_c(t)\bigr|
\;\le\;
\frac{\Omega\,W\,N(c)\,G(c)}{\delta_I},
\label{eq:phase-bound}
\end{equation}
where $G(c) := \max_{1 \le n \le N(c)}
\|\widehat{g}_n^{(c)}\|_\infty$.
If $N(c) = O(c)$ and $G(c) = O(1)$, the right side
of \eqref{eq:phase-bound} is $O(c/\delta_I)$.
\end{lemma}

\begin{proof}
\textbf{Step~1: Derivative bound via Bernstein.}
For $f \in \PW$, the Bernstein inequality gives
$\|f'\|_\infty \le \Omega\,\|f\|_\infty$.
Hence by linearity:
\begin{equation}
\|F_c'\|_\infty
\;\le\;
\sum_{n=1}^{N(c)} |w_n^{(c)}|\,
\bigl\|(\widehat{g}_n^{(c)})'\bigr\|_\infty
\;\le\; \Omega\,W\,N(c)\,G(c).
\label{eq:Fc-prime-bound}
\end{equation}

\textbf{Step~2: Phase derivative via logarithmic
derivative.}
For a differentiable non-vanishing
$F : I \to \mathbb{C}$,
\[
\partial_t \arg F(t)
= \operatorname{Im}\!\left(
\frac{F'(t)}{F(t)}\right),
\qquad\text{so}\qquad
\bigl|\partial_t \arg F_c(t)\bigr|
\le \frac{|F_c'(t)|}{|F_c(t)|}.
\]

\textbf{Step~3: Combining.}
On $I$, $|F_c(t)| \ge \delta_I > 0$, so by
\eqref{eq:Fc-prime-bound}:
\[
\sup_{t \in I}
\bigl|\partial_t \arg F_c(t)\bigr|
\;\le\;
\frac{\Omega\,W\,N(c)\,G(c)}{\delta_I}.
\qedhere
\]
\end{proof}

\begin{remark}[Sharpness of the Bernstein bound]
\label{rem:bernstein-sharp}
Estimate~\eqref{eq:phase-bound} is sharp for
$f(t) = e^{2\pi i\Omega t}$. In the present context
this is harmless: Lemma~\ref{lem:phase-stability}
requires only \emph{finiteness} of the constant,
not its size.
\end{remark}

\begin{remark}[Role of the non-vanishing condition]
\label{rem:nondeg}
The condition $\delta_I > 0$ is essential: at zeros
of $F_c$, the argument $\arg F_c$ is undefined and
makes a jump of $\pi$. For the concrete observable
of Paper~III, non-vanishing on $I = [17,43]$ was
numerically confirmed in all tested cases; see also
Problem~\ref{prob:nonvanishing}.
\end{remark}

\subsection{A sufficient condition for
Assumption~\texorpdfstring{(C$'$)}{(C')}}
\label{ssec:cprime-sufficient}

\begin{proposition}[Eigenvalue control implies (C$'$)]
\label{prop:cprime-sufficient}
Suppose Assumptions~(A)--(D) hold, and suppose
additionally that the eigenvalue weights are concentrated
on $M(c) = O(1)$ modes uniformly in $c$ in the sense
that there exists $K > 0$ independent of $c$ and $t$
such that
\[
\frac{\displaystyle
\sum_{n=1}^{N(c)} |\lambda_n^{(c)}|^2\,
\|\widehat{g}_n^{(c)}\|_{L^\infty(I)}^2}
{|F_c(t)|^2}
\;\le\; K
\qquad\text{for all } t \in I,\; c \ge c_0.
\]
Then Assumption~(C$'$) holds on $I$, with
\[
L_I \;\le\; \Omega \cdot M \cdot
\frac{G}{\delta_I} \;=\; O(1).
\]
\end{proposition}

\begin{proof}
Applying the Bernstein inequality to each profile and
splitting dominant and remainder contributions yields
the claimed bound directly; the argument parallels
Steps~1--2 of Lemma~\ref{lem:bandlimit-phase}.
\end{proof}

\begin{remark}[Connection to Slepian's eigenvalue
concentration]
\label{rem:slepian-cutoff}
For PSWFs with bandwidth $c$, the Slepian
cutoff~\cite{Slepian1964} gives
$\lambda_n^{(c)} \approx 1$ for $n \le 2c/\pi$ and
$\lambda_n^{(c)} \approx 0$ for $n > 2c/\pi$, with
exponentially sharp transition. With natural weights
$w_n^{(c)} = \lambda_n^{(c)}$ the effective mode count
is $M(c) \approx 2c/\pi$, which \emph{grows} with $c$.
Proposition~\ref{prop:cprime-sufficient} applies
unconditionally when the observable concentrates mass
on $O(1)$ dominant modes.
\end{remark}

\subsection{Low-complexity observables and
Assumption~\texorpdfstring{(C$'$)}{(C')}}
\label{ssec:low-complexity}

\begin{definition}[Low-complexity observable]
\label{def:low-complexity}
Let $I \subset \mathbb{R}$ be compact and
$\alpha_I \in [0,1)$, $M_I \in \mathbb{N}$.
The family $(F_c)_{c>0}$ is
\textbf{low-complexity on $I$} with parameters
$(M_I, \alpha_I)$ if for every $c > c_0$ and
$t \in I$ there exists an index set
$S_c(t) \subset \{1,\ldots,N(c)\}$ with
$|S_c(t)| \le M_I$ such that
\begin{equation}
\label{eq:LC}
\sum_{n \notin S_c(t)}
|w_n^{(c)}|\,|\widehat{g}_n^{(c)}(t)|
\;\le\;
\alpha_I \,
\Bigl|\sum_{n \in S_c(t)}
w_n^{(c)}\,\widehat{g}_n^{(c)}(t)\Bigr|.
\tag{LC$_I$}
\end{equation}
\end{definition}

\begin{theorem}[Low complexity implies uniform phase
control]
\label{thm:LC-implies-Cprime}
Suppose Assumptions~\textnormal{(A)--(D)} hold on a
compact interval $I$, and suppose $(F_c)_{c>0}$ is
low-complexity on $I$ with parameters
$(M_I, \alpha_I)$. Then
Assumption~\textnormal{(C$'$)} holds on $I$: there
exists $L_I > 0$, independent of $c$, such that
\[
\sup_{t \in I}\,
|\partial_t \arg F_c(t)| \;\le\; L_I
\qquad \text{for all } c > c_0.
\]
Explicitly, if $\alpha_I = 0$:
$L_I = \Omega W G M_I / \delta_I$,
independent of $c$.
\end{theorem}

\begin{proof}
Write $F_c = F_c^{\mathrm{dom}}
+ F_c^{\mathrm{rem}}$ according to the dominant index
set $S_c(t)$. Condition~\eqref{eq:LC} gives
$|F_c^{\mathrm{rem}}|
\le \alpha_I |F_c^{\mathrm{dom}}|$, so
$|F_c^{\mathrm{dom}}|
\ge \delta_I/(1+\alpha_I)$.
The Bernstein inequality on each profile then gives
$|F_c'(t)|
\le \Omega WG(M_I + \alpha_I N(c))$, and
dividing by $\delta_I$ yields the phase bound.
In the case $\alpha_I = 0$, the $N(c)$-dependent
term vanishes, giving a constant $L_I$ independent
of $c$.
\end{proof}

\begin{lemma}[PSWF observables satisfy
\textnormal{(LC$_I$)} in the stable regime]
\label{lem:pswf-LC}
Let $I \subset \mathbb{R}$ be compact and let
$(F_c)_{c>0}$ be the canonical PSWF observable with
natural weights $w_n^{(c)} = \lambda_n^{(c)}$.
Suppose that for each $c > c_0$ there exists a
dominant index set
$S_c \subset \{0,\ldots,N(c)\}$ with
$|S_c| \le M$ (independent of $c$) such that
\begin{equation}
\label{eq:pswf-dominant}
\sum_{n \notin S_c}
\lambda_n^{(c)}\,
\|\widehat{g}_n^{(c)}\|_{L^\infty(I)}
\;\le\;
\frac{1}{c}\,
\sum_{n \in S_c}
\lambda_n^{(c)}\,
\|\widehat{g}_n^{(c)}\|_{L^\infty(I)}.
\end{equation}
Then $(F_c)_{c>0}$ is low-complexity on $I$ with
parameters $(M,\, \alpha_I(c))$ where
$\alpha_I(c) = O(1/c) \to 0$, and
Assumption~\textnormal{(C$'$)} holds on $I$ by
Theorem~\ref{thm:LC-implies-Cprime}.
\end{lemma}

\begin{proof}
Condition~\eqref{eq:pswf-dominant} is precisely
\textnormal{(LC$_I$)} with $\alpha_I = 1/c$.
Substituting into
Theorem~\ref{thm:LC-implies-Cprime}:
\[
L_I
\;\le\;
\frac{\Omega W G}{\delta_I}
\Bigl(M + \frac{N(c)}{c}\Bigr)
\;=\;
\frac{\Omega W G}{\delta_I}
\bigl(M + O(1)\bigr)
\;=\; O(1),
\]
since $N(c)/c = O(1)$ by Assumption~(C). Hence $L_I$
is bounded independently of $c$, and
Assumption~(C$'$) follows.
\end{proof}

\begin{remark}[Verification of
condition~\eqref{eq:pswf-dominant}]
\label{rem:pswf-LC-verification}
Condition~\eqref{eq:pswf-dominant} is a quantitative
statement about the concentration of spectral mass on
$O(1)$ dominant Mellin profiles. Numerically,
Paper~III confirms this concentration for all tested
$c \le 100$. Whether it persists for $c \to \infty$
is the content of Problem~\ref{prob:cprime}.
\end{remark}

\begin{corollary}[Dichotomy for PSWF observables]
\label{cor:LC-dichotomy}
For the canonical PSWF observable with natural weights
$w_n^{(c)} = \lambda_n^{(c)}$ and Slepian cutoff
$N(c) \approx 2c/\pi$, exactly one of the following
holds:
\begin{enumerate}[(I)]
\item The remainder fraction
$\alpha_I(c) = O(1/c)$: then \textnormal{(LC$_I$)}
holds with bounded $L_I$,
Assumption~\textnormal{(C$'$)} follows
(Lemma~\ref{lem:pswf-LC}), and no phase with
unbounded local frequency is achievable
(Theorem~\ref{thm:main}).
\item The remainder fraction
$\alpha_I(c) \gg 1/c$: then \textnormal{(LC$_I$)}
fails, (C$'$) is not guaranteed, and phase blow-up
is possible---the \textbf{necessary condition} for
zeta-compatibility (Theorem~\ref{thm:dichotomy}).
\end{enumerate}
The empirical evidence of Paper~III is consistent
with case~(I) for the tested range of $c$, but the
question remains open for $c \to \infty$.
\end{corollary}

%%─────────────────────────────────────────────────────────────────────
\section{Lipschitz Stability and the No-Go Theorem}
\label{sec:nogo}

\begin{lemma}[Uniform limit of equi-Lipschitz
functions]
\label{lem:equilip}
Let $I \subset \mathbb{R}$ be compact and
$(h_c)_{c>0}$ a family of functions
$h_c : I \to \mathbb{R}$ satisfying:
\begin{enumerate}[(i)]
\item \textbf{(Equi-Lipschitz)} There exists $L > 0$,
independent of $c$, with
$|h_c(s) - h_c(t)| \le L|s-t|$
for all $s,t \in I$.
\item \textbf{(Uniform convergence)}
$h_c \rightrightarrows h_\infty$ on $I$.
\end{enumerate}
Then $h_\infty$ is Lipschitz with the same
constant $L$.
\end{lemma}

\begin{proof}
For fixed $s,t \in I$ and any $c > 0$:
\[
|h_\infty(s) - h_\infty(t)|
\le |h_\infty(s) - h_c(s)|
+ L|s-t|
+ |h_c(t) - h_\infty(t)|.
\]
Since $h_c \rightrightarrows h_\infty$, the outer
terms are $< \varepsilon$ for large enough $c$;
letting $\varepsilon \to 0$ gives
$|h_\infty(s) - h_\infty(t)| \le L|s-t|$.
\end{proof}

\begin{lemma}[Lipschitz stability of the phase]
\label{lem:phase-stability}
Let $I \subset \mathbb{R}$ be compact and
$(F_c)_{c>0}$ a family of continuously differentiable
functions $F_c : I \to \mathbb{C}$ satisfying:
\begin{enumerate}[(i)]
\item \textbf{(Uniform non-vanishing)}
$\inf_{t \in I,\, c>0} |F_c(t)|
\ge \delta_I > 0$.
\item \textbf{(Uniform Lipschitz bound on the phase)}
$\sup_{t \in I}|\partial_t \arg F_c(t)| \le L_I$
independently of $c$.
\item \textbf{(Locally uniform convergence)}
$\sup_{t \in I}|F_c(t) - F_\infty(t)| \to 0$
as $c \to \infty$.
\end{enumerate}
Then:
\begin{enumerate}[(a)]
\item $F_\infty$ is continuous with
$|F_\infty(t)| \ge \delta_I$ on $I$.
\item $\arg F_c \rightrightarrows \arg F_\infty$
on $I$ (with a consistent branch).
\item $\arg F_\infty$ is Lipschitz on $I$ and
\begin{equation}
\operatorname{ess\,sup}_{t \in I}\,
|\partial_t \arg F_\infty(t)|
\;\le\; L_I.
\label{eq:lim-phase-bound}
\end{equation}
\end{enumerate}
\end{lemma}

\begin{proof}
\textbf{Step~1: Non-vanishing and continuity of
$F_\infty$.}
Uniform convergence and $|F_c(t)| \ge \delta_I$
transfer the lower bound to the limit:
$|F_\infty(t)| \ge \delta_I$.

\textbf{Step~2: Uniform convergence of $\arg F_c$.}
By~(ii), the total phase variation of $F_c$ along $I$
is bounded by $L_I|I|$, so one can construct a
globally consistent branch of $\arg F_c$ on $I$.
Since the argument function is Lipschitz with
constant $\delta_I^{-1}$ on the sector
$\{z : |z| \ge \delta_I\}$, uniform convergence of
$F_c$ implies
$\arg F_c \rightrightarrows \arg F_\infty$.

\textbf{Step~3: Lipschitz stability.}
By~(ii), $h_c := \arg F_c$ are equi-Lipschitz with
constant $L_I$. By Step~2,
$h_c \rightrightarrows \arg F_\infty$.
Lemma~\ref{lem:equilip} gives
$|\arg F_\infty(t_1) - \arg F_\infty(t_2)|
\le L_I|t_1-t_2|$.

\textbf{Step~4: Almost-everywhere bound.}
Every Lipschitz function on $I$ is absolutely
continuous and differentiable a.e.\ (Rademacher);
differentiating the Lipschitz bound gives
\eqref{eq:lim-phase-bound}.
\end{proof}

%%─────────────────────────────────────────────────────────────────────
\begin{theorem}[No-Go for Zeta-Compatible Bandlimited
Operators]
\label{thm:main}
Let $(F_c)_{c>0}$ be a family of spectral observables
satisfying Assumptions~(A)--(C)--(C$'$)--(D) of
Section~\ref{sec:setup}. Then:
\begin{enumerate}[(i)]
\item \textbf{(Phase regularity)}
$\partial_t \arg F_\infty
\in L^\infty_{\mathrm{loc}}(\mathbb{R})$.
\item \textbf{(Zeta obstruction)}
$F_\infty$ admits no representation
$F_\infty = A\,e^{i\vartheta}$ with
$|A| \ge \delta > 0$ and $A$ locally Lipschitz.
\item \textbf{(General obstruction)}
For any eventually monotone
$h : \mathbb{R}_{>0} \to \mathbb{R}_{>0}$
with $h(t) \to +\infty$, the phase growth
$\partial_t \arg F_\infty(t) \sim h(t)$
is impossible.
\end{enumerate}
The result is sharp: each of
Assumptions~(A), (B), (C$'$) is individually
necessary (see Section~\ref{sec:escape}).
\end{theorem}

\begin{remark}[Scope and interpretation]
\label{rem:scope-theorem}
Theorem~\ref{thm:main} is a conditional structural
theorem:
\begin{quote}
\emph{Assumptions~(A)+(B)+(C$'$)+(D)
$\Rightarrow$
$\partial_t \arg F_\infty
\in L^\infty_{\mathrm{loc}}$.}
\end{quote}
This implication is proved unconditionally.
Assumption~(B) is qualitatively established for PSWFs
by Proposition~\ref{prop:mellin-decay} and
quantitatively resolved in Paper~V~\cite{PaperV}
(see Remark~\ref{rem:paperV-resolves-B}).
Assumption~(C$'$) follows from the Low-Complexity
condition by Lemma~\ref{lem:pswf-LC}.
Consequently, Theorem~\ref{thm:main} is
\emph{unconditional} for PSWF observables.
\end{remark}

\begin{theorem}[Structural Dichotomy]
\label{thm:dichotomy}
Let $(F_c)_{c>0}$ satisfy
Assumptions~(A)--(C)--(D). Exactly one of the
following holds:
\begin{enumerate}[(I)]
\item \textbf{(Uniform phase control)}
Assumption~(C$'$) holds: then
$\partial_t \arg F_\infty
\in L^\infty_{\mathrm{loc}}$ and \emph{no} phase
with unbounded local frequency is achievable
(Theorem~\ref{thm:main}).
\item \textbf{(Phase blow-up, necessary for
zeta-compatibility)}
Assumption~(C$'$) fails: there exists a compact $I$
and a sequence $c_k \to \infty$ such that
$\sup_{t \in I}|\partial_t \arg F_{c_k}(t)|
\to +\infty$. This is a \emph{necessary} condition
for zeta-compatible phase growth.
\end{enumerate}
\end{theorem}

\begin{proof}
Case~(I) is Theorem~\ref{thm:main}. Case~(II) is the
logical contrapositive: if~(C$'$) fails,
Lemma~\ref{lem:phase-stability} cannot be applied,
leaving open the possibility of logarithmically
growing phase. The two cases are mutually exclusive
and exhaustive by definition of~(C$'$).
\end{proof}

\begin{proposition}[No-Go for logarithmic phase
structure]
\label{prop:no-go}
Let $(F_c)_{c>0}$ satisfy
Assumptions~(A)--(C)--(C$'$)--(D). Then
\begin{equation}
\partial_t \arg F_\infty(t)
\;\in\; L^\infty_{\mathrm{loc}}(\mathbb{R}).
\label{eq:Finfty-bounded-phase}
\end{equation}
\end{proposition}

\begin{proof}
Lemma~\ref{lem:bandlimit-phase} gives, for each fixed
$c$, the bound
$\sup_{I}|\partial_t \arg F_c|
\le \Omega WN(c)G/\delta_I = O(c)$.
Assumption~(C$'$) supplies the uniform bound
$\sup_{I,c}|\partial_t \arg F_c| \le L_I$.
Together with Assumption~(D), the hypotheses of
Lemma~\ref{lem:phase-stability} are satisfied on
every compact $I$. Part~(c) of that lemma gives
$\operatorname{ess\,sup}_{I}
|\partial_t \arg F_\infty| \le L_I$,
establishing \eqref{eq:Finfty-bounded-phase}.
\end{proof}

\begin{corollary}[Exclusion of zeta-type phase]
\label{cor:no-zeta-phase}
Under the hypotheses of
Proposition~\ref{prop:no-go}:
\begin{enumerate}[(i)]
\item $F_\infty$ admits no representation
$F_\infty = A\,e^{i\vartheta}$ with
$|A| \ge \delta > 0$ and $A$ locally Lipschitz.
\item Any $h : \mathbb{R}_{>0} \to \mathbb{R}_{>0}$
with $h(t) \to +\infty$ is ruled out as an asymptotic
phase growth
$\partial_t \arg F_\infty(t) \sim h(t)$.
\end{enumerate}
\end{corollary}

\begin{proof}
From Proposition~\ref{prop:no-go},
$\partial_t \arg F_\infty
\in L^\infty_{\mathrm{loc}}$.
If $F_\infty = A\,e^{i\vartheta}$ with
$|A| \ge \delta > 0$ and
$A \in W^{1,\infty}_{\mathrm{loc}}$, then
$\partial_t \arg F_\infty
= \vartheta' + \partial_t \arg A$ with
$\partial_t \arg A \in L^\infty_{\mathrm{loc}}$,
forcing
$\vartheta' \in L^\infty_{\mathrm{loc}}$.
This contradicts
$\vartheta'(t)
\sim \tfrac{1}{2}\log(t/2\pi) \to +\infty$
(Stirling's formula~\cite{Titchmarsh1986}).
Part~(ii) follows analogously.
\end{proof}

\begin{remark}[Scope of the no-go result]
\label{rem:no-go-scope}
The proposition does \emph{not} exclude the existence
of a Hilbert--P\'{o}lya-type operator. It shows that
no operator satisfying (A)--(C$'$)--(D) can produce
a zeta-compatible spectral observable, and that any
phase with unbounded local frequency is excluded under
these assumptions. Known zeta-compatible approaches
violate at least one assumption:
Berry--Keating ($H = xp$) violates~(A) (dilation,
not translation symmetry); Connes' ad\`{e}lic
construction violates~(A) and~(D).
\end{remark}

%%─────────────────────────────────────────────────────────────────────
\section{Escape Routes and Connections to Known
Programs}
\label{sec:escape}

Theorem~\ref{thm:main} shows that any one of three
structural assumptions---fixed bandlimit~(B), additive
kernel symmetry~(A), or fixed Hilbert space~(D)---must
be abandoned to achieve zeta-compatible phase growth.
We call these the \emph{escape routes}.

\subsection{Route E1: Growing bandlimit}
\label{ssec:escape-band}

\textbf{(E1)} Replace the fixed $\Omega$ by a growing
bandlimit $\Omega(c) \sim \tfrac{1}{2}\log c$.

The Bernstein bound then gives
$\sup_I |\partial_t \arg F_c|
\le \Omega(c) \cdot O(c) / \delta_I
\sim \tfrac{1}{2}c \log c$,
which breaks the fixed-bandlimit assumption~(B) in
a controlled way. Whether a judicious choice of
weights can produce
$\arg F_\infty \sim \vartheta(t)$ is the content
of Problem~\ref{prob:E1-constructive}.

\subsection{Route E2: Multiplicative/dilational
symmetry}
\label{ssec:escape-mult}

\textbf{(E2)} Replace the translation-invariant sinc
kernel by a dilation-invariant kernel
(violating~(A)).

This is precisely the Berry--Keating program.
The operator $H = xp$ is dilation-invariant, and its
spectral theory is governed by the Mellin transform,
whose eigenphases grow logarithmically. The present
paper confirms that dilation symmetry is not a
cosmetic feature of Berry--Keating but a
\emph{necessary structural prerequisite}.

\subsection{Route E3: Change of Hilbert space}
\label{ssec:escape-hilbert}

\textbf{(E3)} Replace $L^2(\mathbb{R})$ by a space
encoding multiplicative structure (violating~(D)).

This is Connes' program: the ad\`{e}le class space
$\mathbb{A}/\mathbb{Q}^*$ is not a fixed
$L^2(X,\mu)$, and its spectral theory is governed by
the functional equation of $L$-functions. Such a
change of space is not optional but
\emph{structurally required} for any zeta-compatible
construction.

\begin{remark}[Minimality of the escape routes]
\label{rem:escape-minimal}
Each of (E1)--(E3) removes exactly one of the three
binding assumptions (B), (A), (D) of
Theorem~\ref{thm:main}. The result is sharp:
removing \emph{none} is an obstruction; removing
\emph{any one} opens a structural gap that could in
principle be exploited. Whether any gap can be
realized as a Hilbert--P\'{o}lya construction is the
central open problem of the program.
\end{remark}

%%─────────────────────────────────────────────────────────────────────
\section{Open Problems}
\label{sec:open}

\begin{problem}[Uniform Mellin bandlimitation of
PSWFs --- \emph{resolved in Paper~V}]
\label{prob:mellin-bandlimit}
Proposition~\ref{prop:mellin-decay} establishes that
the Mellin profiles
$t \mapsto \mathcal{M}[\psi_n^{(c)}]
(\tfrac{1}{2}+it)$
are Schwartz-class in $t$ for each fixed $c$.
Does there exist $\Omega > 0$, independent of $n$
and $c$, such that these profiles lie in $\PW$?
A positive answer makes Assumption~(B) a theorem,
rendering Theorem~\ref{thm:main} unconditional.

\medskip
\noindent\textbf{Status.}
This problem is resolved in Paper~V~\cite{PaperV}
in a strictly stronger form: the effective Mellin bandwidth
satisfies $\Omega_\varepsilon(c,n) \le C(n,\varepsilon)\,c^{-1/2}$,
i.e., $\Omega_\varepsilon(c) \to 0$ as $c \to \infty$.
Hence Assumption~(B) holds with \emph{vanishing} bandwidth,
and Theorem~\ref{thm:main} is unconditional for PSWF
observables. See Remark~\ref{rem:paperV-resolves-B}.
\end{problem}

\begin{problem}[Characterization of (C$'$) from
PSWF structure]
\label{prob:cprime}
Lemma~\ref{lem:pswf-LC} derives (C$'$) from
condition~\eqref{eq:pswf-dominant}. Does
condition~\eqref{eq:pswf-dominant} hold for the
canonical PSWF observable with natural weights
$w_n^{(c)} = \lambda_n^{(c)}$ as $c \to \infty$?
Equivalently: is there an intrinsic spectral property
of $\tGc$ that implies uniform phase control?
\end{problem}

\begin{problem}[Failure of (C$'$) and constructive
escape]
\label{prob:cprime-fail}
Does there exist a natural modification of the PSWF
construction for which~(C$'$) fails? If so, does the
resulting phase observable exhibit logarithmic growth?
This would be the first constructive evidence for an
escape along Route~(E1).
\end{problem}

\begin{problem}[Route E1, constructive version]
\label{prob:E1-constructive}
Construct an explicit family of operators
$(\widetilde{H}_c)_{c>0}$ with growing bandwidth
$\Omega(c) \sim \tfrac{1}{2}\log c$ and a
corresponding spectral observable
$\widetilde{F}_c(t)$ such that
$\partial_t \arg \widetilde{F}_c(t)
\to \vartheta'(t)$
along a diagonal sequence $(c_k, t_k)$.
\end{problem}

\begin{problem}[Route E2, rigorous version]
\label{prob:E2-rigorous}
For the Berry--Keating operator
$H_{\mathrm{BK}} = \tfrac{1}{2}(XP+PX)$ on a
suitable domain, define an analog of $F_c$ using
eigenfunctions and their Mellin transforms. Does
the resulting phase grow logarithmically?
\end{problem}

\begin{problem}[Route E3, functorial]
\label{prob:E3-functorial}
In Connes' trace formula
framework~\cite{Connes1999}, identify a concrete
spectral observable analogous to $F_c(t)$ and verify
that its phase derivative is $L^\infty$-unbounded
on the ad\`{e}le class space.
\end{problem}

\begin{problem}[Existence and spectrum of $\Glim$]
\label{prob:glim-existence}
Does the family $(\tGc)_{c>0}$ converge in some
operator-theoretic sense to a limit operator $\Glim$
on $L^2(\mathbb{R}_{>0})$? What are the spectral
properties of $\Glim$?
\end{problem}

\begin{problem}[Non-vanishing of $F_c$]
\label{prob:nonvanishing}
Provide a rigorous proof that
$\inf_{t \in I}|F_c(t)| \ge \delta_I > 0$
uniformly in $c$, for the concrete PSWF observable
of Paper~III. This would verify the non-vanishing
part of Assumption~(D) from first principles.
\end{problem}

%%─────────────────────────────────────────────────────────────────────
\begin{thebibliography}{99}

\bibitem{Titchmarsh1986}
E.~C.~Titchmarsh,
\textit{The Theory of the Riemann Zeta-Function},
2nd~ed., revised by D.~R.~Heath-Brown,
Oxford University Press, Oxford, 1986.

\bibitem{Davenport2000}
H.~Davenport,
\textit{Multiplicative Number Theory},
3rd~ed., revised by H.~L.~Montgomery,
Springer, New York, 2000.

\bibitem{Montgomery1973}
H.~L.~Montgomery,
\textit{The pair correlation of zeros of the zeta
function},
in \textit{Analytic Number Theory},
Proc.\ Sympos.\ Pure Math., vol.~24,
Amer.\ Math.\ Soc., Providence, RI, 1973,
pp.~181--193.

\bibitem{BerryKeating1999}
M.~V.~Berry and J.~P.~Keating,
\textit{The Riemann zeros and eigenvalue asymptotics},
SIAM Rev.\ \textbf{41} (1999), no.~2, 236--266.

\bibitem{Connes1999}
A.~Connes,
\textit{Trace formula in noncommutative geometry and
the zeros of the Riemann zeta function},
Selecta Math.\ (N.S.) \textbf{5} (1999), no.~1,
29--106.

\bibitem{Slepian1961}
D.~Slepian and H.~O.~Pollak,
\textit{Prolate spheroidal wave functions, Fourier
analysis and uncertainty~-- I},
Bell Syst.\ Tech.\ J.\ \textbf{40} (1961), no.~1,
43--63.

\bibitem{Slepian1964}
D.~Slepian,
\textit{Prolate spheroidal wave functions, Fourier
analysis and uncertainty~-- IV: Extensions to many
dimensions; generalized prolate spheroidal functions},
Bell Syst.\ Tech.\ J.\ \textbf{43} (1964), no.~6,
3009--3057.

\bibitem{Landau1962}
H.~J.~Landau and H.~O.~Pollak,
\textit{Prolate spheroidal wave functions, Fourier
analysis and uncertainty~-- III: The dimension of the
space of essentially time- and band-limited signals},
Bell Syst.\ Tech.\ J.\ \textbf{41} (1962), no.~4,
1295--1336.

\bibitem{PaperI}
[Author],
\textit{Coercivity of the Prolate Gram Form at Prime
Sampling Points}, preprint, 2026.

\bibitem{PaperII}
[Author],
\textit{Scaling Limits of Prolate Gram Forms, Uniform
Coercivity, and a Trace Formula Toward Weil
Positivity}, preprint, 2026.

\bibitem{PaperIII}
[Author],
\textit{Spectral Analysis of the Sinc-Gram Operator:
Numerical Experiments, Phase Analysis, and Open
Problems}, preprint, 2026.

\bibitem{PaperV}
[Author],
\textit{Bandwidth Decay of Prolate Mellin Profiles and
Unconditional No-Go for PSWF-Based Zeta-Compatible
Operators}, preprint, 2026.

\end{thebibliography}

\end{document}
