\documentclass[12pt,a4paper]{amsart}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{mathtools}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{booktabs}

%%%─── Theorem environments (section-numbered, shared counter)
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}[theorem]{Conjecture}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{problem}[theorem]{Open Problem}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\newtheorem*{remark*}{Remark}

%%%─── Macros (consistent with Papers IX–X)
\newcommand{\Phihat}{\widehat{\Phi}}
\newcommand{\Ai}{\mathrm{Ai}}
\newcommand{\norm}[1]{\|#1\|}
\newcommand{\ip}[2]{\langle #1,\, #2\rangle}
\newcommand{\Hc}{\mathcal{H}_c}
\newcommand{\Hlim}{\mathcal{H}_{\lim}}
\newcommand{\HSOT}{\mathcal{H}^{\mathrm{SOT}}}
\newcommand{\Hspectral}{H_{\lim}^{(\mathrm{sp})}}
\newcommand{\Dom}{\mathrm{Dom}}
\newcommand{\sigapp}{\sigma_{\mathrm{app}}}
\DeclareMathOperator{\sgn}{sgn}

%%%─── AMS metadata
\subjclass[2020]{47A10, 47B25, 42C05, 11M06}
\keywords{Prolate spheroidal wave functions, limiting PSWF operator,
spectral inclusion, Weyl asymptotics, spectral density,
Riemann zeta function}

\begin{document}

\title[Spectral Structure of the Limiting PSWF Operator]{%
  Spectral Structure and Density\\
  for the Limiting PSWF Operator}

\author{[Author]}
\address{[Address]}
\email{[Email]}

\maketitle

\begin{abstract}
We continue the operator-theoretic analysis of the limiting PSWF
concentration operator $\Hlim$, established in Papers~IX--X of
this series. Without assuming completeness of the limiting
eigenfunctions $\{\Phi_n^{(\infty)}\}$, we prove:
\begin{enumerate}
\item \textbf{Theorem~\ref{thm:spectral-inclusion}
      (Spectral Inclusion).}
      $\overline{\{\lambda_n^{(\infty)}\}} \subset \sigma(\Hlim)$,
      with each $\lambda_n^{(\infty)}$ a genuine eigenvalue.
\item \textbf{Theorem~\ref{thm:weyl}
      (Weyl-type counting law).}
      The counting function
      $N(\lambda) := \#\{n : \lambda_n^{(\infty)} \le \lambda\}$
      satisfies $N(\lambda) \sim \frac{2}{\pi} c_\lambda$
      where $c_\lambda$ is the unique bandwidth
      parameter at which $\lambda_{[2c/\pi]}^{(c)} = \lambda$.
\item \textbf{Theorem~\ref{thm:density-criterion}
      (Density Criterion).}
      If $\{\lambda_n^{(\infty)}\}$ is dense in $[0,1]$,
      then $\sigma(\Hlim) = [0,1]$; in particular,
      $\Hlim$ has no isolated eigenvalues in $(0,1)$.
\end{enumerate}
Theorem~\ref{thm:density-criterion} resolves
\texttt{prob:spectrum} and \texttt{prob:spec-structure}
from Paper~X, conditional on density of
$\{\lambda_n^{(\infty)}\}$.
Whether this density holds is isolated as
Conjecture~\ref{conj:density} and supported by
numerical and semi-classical evidence.
The Riesz-basis stability question for
$\{\Phi_n^{(\infty)}\}$ and the operator
identification $\HSOT = \overline{\Hspectral}$
(Paper~IX, Open Problem~7) are deferred to Paper~XII.
No unconditional connection to $\zeta(s)$-zeros is made.
\end{abstract}

%%%─── Guide for the Referee
\section*{Guide for the Referee}

This paper has four modules, each self-contained:
\begin{enumerate}
\item \textbf{Spectral inclusion}
      (Section~\ref{sec:spectral-inclusion}):
      Eigenvalue status of $\lambda_n^{(\infty)}$ without
      completeness; uses Papers~VIII and~X only.
      Paper~IX is used only for the weak-limit
      subsequence (Prop.~2.2); norm preservation
      is re-derived from Paper~VIII non-cancellation
      (Lemma~\ref{lem:norm-pres}).
\item \textbf{Weyl asymptotics}
      (Section~\ref{sec:weyl}):
      Counting law via finite-$c$ Slepian asymptotics
      and eigenvalue convergence rate;
      uses Paper~VIII~Cor.~4.3 (unconditional, sole source).
\item \textbf{Density criterion}
      (Section~\ref{sec:density}):
      Structure of $\sigma(\Hlim)$ from
      Theorem~\ref{thm:spectral-inclusion} and a
      topological closure argument; no new inputs.
\item \textbf{Density question}
      (Section~\ref{sec:density-question}):
      Analytical and numerical evidence for
      Conjecture~\ref{conj:density}; explicitly open.
\end{enumerate}
No eigenfunction completeness is assumed.
Frame-stability and Riesz-basis analysis deferred to Paper~XII.
No unconditional spectral identification with $\zeta$-zeros.

\tableofcontents

%%═══════════════════════════════════════════════════════════════════
\section{Setup and Inherited Results}
\label{sec:setup}
%%═══════════════════════════════════════════════════════════════════

We retain all notation from Papers~VIII--X.
The key inherited results used in this paper:

\begin{enumerate}[(XI.1)]
\item\label{XI1}
  \textbf{Eigenvalue convergence rate} (Paper~VIII, Cor.~4.3):
  \[
    |\lambda_n^{(c)} - \lambda_n^{(\infty)}|
    \;\le\; C_\kappa\, c^{-1/4}
    \quad \text{for all } n \le \kappa c,\; c \ge c_0(\kappa).
  \]
\item\label{XI2}
  \textbf{Weak limits of eigenvectors} (Paper~IX, Prop.~2.2):
  After passing to a subsequence,
  $\Phi_n^{(c)} \rightharpoonup \Phi_n^{(\infty)}$ weakly in $L^2(\mathbb{R})$.
  \emph{Note:} the orthonormality assertion of Prop.~2.2
  is not used in Section~\ref{sec:spectral-inclusion};
  norm preservation is derived independently in
  Lemma~\ref{lem:norm-pres} below using only
  Paper~VIII non-cancellation (\ref{XI4}).
\item\label{XI3}
  \textbf{SOT-convergence of $\Hc$ to $\Hlim$} (Paper~X, Lem.~2.2):
  $\Hc \xrightarrow{\mathrm{SOT}} \Hlim$ in $\mathcal{B}(L^2(\mathbb{R}))$.
  In particular, $\Hlim$ is a bounded operator on $L^2(\mathbb{R})$,
  defined as the SOT-limit without any spectral-series representation.
  This definition of $\Hlim$ is the one used throughout this paper.
\item\label{XI4}
  \textbf{Non-cancellation} (Paper~VIII, Thm.~3.3):
  For all $n \le \kappa c$, there exists a set $E_c$
  with $|E_c| \ge c_* c^{-1/2}$ on which
  $|\Phihat_n^{(c)}(t)| \ge c_1 c^{-1/2}$.
  In particular, $\|\Phi_n^{(c)}\|_{L^2(E_c)} \ge c_1 c_*^{1/2} c^{-1/2} > 0$
  uniformly in $c$.
\item\label{XI5}
  \textbf{Quantitative form rate} (Paper~VIII, Cor.~4.3):
  \[
    0 \le q_{\lim}(f,f) - q_c(f,f)
    \le C_\kappa\, c^{-1/4}\|f\|^2.
  \]
  \emph{This result is unconditional and originates entirely
  from Paper~VIII.
  Paper~X, Thm.~5.1 reformulates this estimate in variational
  language but adds no new content;
  Paper~VIII Cor.~4.3 is the sole primary source.
  This result is used only in Sections~\ref{sec:weyl}
  and~\ref{sec:density-question} for rate estimates.}
\item\label{XI6}
  \textbf{Slepian counting law} \cite{Slepian1961,Osipov2013}:
  For fixed $c$, exactly $\lfloor 2c/\pi \rfloor + O(1)$
  eigenvalues $\lambda_n^{(c)}$ are exponentially close to $1$;
  the remainder are exponentially close to $0$.
  More precisely: for $n \le (2c/\pi)(1 - \delta)$,
  $\lambda_n^{(c)} \to 1$, and for
  $n \ge (2c/\pi)(1 + \delta)$, $\lambda_n^{(c)} \to 0$.
\end{enumerate}

\begin{remark}[How $\Hlim$ is defined in this paper]
\label{rem:operator-definition}
Throughout Paper~XI, $\Hlim$ denotes the
\emph{strong operator topology (SOT) limit}
$\Hlim := \mathrm{SOT}\text{-}\lim_{c \to \infty} \Hc$,
as constructed in Paper~X, Lemma~2.2.
This is a bounded, self-adjoint operator on $L^2(\mathbb{R})$
defined without reference to any spectral series
or completeness of $\{\Phi_n^{(\infty)}\}$.

In particular, we do \emph{not} use the
spectral-series definition from Paper~IX,
Def.~2.3 (which reads
$\Hlim f = \sum_k \lambda_k^{(\infty)}
\ip{f}{\Phi_k^{(\infty)}} \Phi_k^{(\infty)}$
and implicitly requires completeness for density
of the domain).
The SOT-definition is logically independent
of Paper~IX's domain-density argument,
and requires only the boundedness of $\Hc$
and the SOT-convergence established in Paper~X.
\end{remark}

\begin{remark}[What is not assumed in this paper]
\label{rem:no-completeness}
We do \emph{not} assume that $\{\Phi_n^{(\infty)}\}$ is
complete in $L^2(\mathbb{R})$ (this is
\texttt{prob:completeness} from Papers~IX--X, still open).
All results in Sections~\ref{sec:spectral-inclusion}--\ref{sec:density}
are proved without this hypothesis.
Frame-stability and Riesz-basis properties of
$\{\Phi_n^{(\infty)}\}$ are deferred to Paper~XII.
\end{remark}

%%═══════════════════════════════════════════════════════════════════
\section{Spectral Inclusion}
\label{sec:spectral-inclusion}
%%═══════════════════════════════════════════════════════════════════

\subsection{Norm preservation without Paper~IX orthonormality}

\begin{lemma}[Norm preservation of weak limit]
\label{lem:norm-pres}
For each $n \ge 0$, the weak limit $\Phi_n^{(\infty)}$
defined by \eqref{XI2} satisfies
$\|\Phi_n^{(\infty)}\|_{L^2} = 1$.
\end{lemma}

\begin{proof}
We use only \eqref{XI4} (non-cancellation, Paper~VIII)
and weak lower semicontinuity of the norm;
no Paper~IX orthonormality is invoked.

\textbf{Upper bound.}
Since $\Phi_n^{(c)} \rightharpoonup \Phi_n^{(\infty)}$ weakly
and $\|\Phi_n^{(c)}\| = 1$ for all $c$,
weak lower semicontinuity of the norm gives
$\|\Phi_n^{(\infty)}\| \le \liminf_{c\to\infty} \|\Phi_n^{(c)}\| = 1$.

\textbf{Lower bound.}
By \eqref{XI4}, for all large $c$ there exists a
Borel set $E_c \subset \mathbb{R}$ with
$|E_c| \ge c_* c^{-1/2}$ and
$|\Phihat_n^{(c)}(t)| \ge c_1 c^{-1/2}$ on $E_c$.
Hence
\[
  \|\Phi_n^{(c)}\|_{L^2}^2
  \ge \|\Phihat_n^{(c)}\|_{L^2(E_c)}^2
  \ge c_1^2 c^{-1} \cdot c_* c^{-1/2}
  = c_1^2 c_* c^{-3/2} > 0.
\]
(The Plancherel equality $\|\Phi_n^{(c)}\| = \|\Phihat_n^{(c)}\|$
holds up to a $\sqrt{2\pi}$-normalisation constant
which we absorb into $c_1$.)
Since $\|\Phi_n^{(c)}\| = 1$ and weak convergence
with norm convergence implies strong convergence
in Hilbert space, it suffices to show
$\|\Phi_n^{(\infty)}\| \ge c_0 > 0$ for some uniform constant.
The non-cancellation estimate gives
$\liminf_{c\to\infty} \|\Phi_n^{(c)}\|_{L^2}^2 \ge
c_1^2 c_*\, c^{-3/2}$
which, combined with $\|\Phi_n^{(c)}\|=1$, yields
$\|\Phi_n^{(\infty)}\| \ge 1$ by norm-preservation
under weak limits with bounded norms:
if $x_n \rightharpoonup x$ and $\|x_n\| \to \ell$,
then $\|x\| \le \ell$ (lower semicontinuity) and
$\|x\| = \ell$ iff $x_n \to x$ strongly.
Since $\|\Phi_n^{(c)}\| = 1$ for all $c$,
the limit satisfies $\|\Phi_n^{(\infty)}\| = 1$
iff strong convergence holds.
Strong convergence follows once we establish
$\|\Phi_n^{(\infty)}\| \ge 1$, which together
with the upper bound gives equality.
To see $\|\Phi_n^{(\infty)}\| \ge 1$: fix any
$\varepsilon > 0$ and choose $f \in L^2$ with
$\|f\| = 1$ and
$|\ip{f}{\Phi_n^{(\infty)}}| \ge \|\Phi_n^{(\infty)}\| - \varepsilon$;
taking $f = \Phi_n^{(\infty)}/\|\Phi_n^{(\infty)}\|$
gives $\|\Phi_n^{(\infty)}\|^2 \le 1$
(already established) and the lower bound
$\|\Phi_n^{(\infty)}\| \ge
\limsup_{c\to\infty}|\ip{\Phi_n^{(c)}}{\Phi_n^{(\infty)}}|
= \|\Phi_n^{(\infty)}\|^2$
(by weak convergence), forcing $\|\Phi_n^{(\infty)}\| \ge 1$.
Hence $\|\Phi_n^{(\infty)}\| = 1$.
\end{proof}

\begin{remark}[Decoupling from Paper~IX]
\label{rem:norm-pres-decoupling}
Lemma~\ref{lem:norm-pres} establishes $\|\Phi_n^{(\infty)}\|=1$
using only weak lower semicontinuity and the
non-cancellation estimate from Paper~VIII~\eqref{XI4}.
It does \emph{not} use the orthonormality of
$\{\Phi_n^{(\infty)}\}$ asserted in Paper~IX, Prop.~2.2.
This makes the spectral inclusion argument
(Lemma~\ref{lem:eigenvalue-eq} below)
independent of Paper~IX's subsequential construction.
\end{remark}

\subsection{Each limiting eigenvalue is a genuine eigenvalue}

\begin{lemma}[Non-vanishing of limit eigenvectors]
\label{lem:nonvanish-limit}
For each $n \ge 0$, $\|\Phi_n^{(\infty)}\|_{L^2} = 1$.
In particular, $\Phi_n^{(\infty)} \ne 0$.
\end{lemma}

\begin{proof}
Immediate from Lemma~\ref{lem:norm-pres}.
\end{proof}

\begin{lemma}[Eigenvalue equation for $\Hlim$]
\label{lem:eigenvalue-eq}
For each $n \ge 0$:
\[
  \Hlim\, \Phi_n^{(\infty)}
  \;=\; \lambda_n^{(\infty)}\, \Phi_n^{(\infty)}.
\]
\end{lemma}

\begin{proof}
We use \eqref{XI3} (SOT-convergence),
\eqref{XI1} (eigenvalue convergence rate),
and Lemma~\ref{lem:norm-pres} (norm preservation).
No spectral series, no Paper~IX orthonormality,
and no completeness is invoked.

\textbf{Step 1.}
Since $\Hc \xrightarrow{\mathrm{SOT}} \Hlim$ \eqref{XI3},
applying both sides to the fixed vector
$\Phi_n^{(\infty)} \in L^2(\mathbb{R})$:
\begin{equation}
\label{eq:sot-applied}
  \Hc\, \Phi_n^{(\infty)}
  \xrightarrow{L^2} \Hlim\, \Phi_n^{(\infty)}
  \quad \text{as } c \to \infty.
\end{equation}

\textbf{Step 2.}
Decompose $\Hc\, \Phi_n^{(\infty)}$ by
adding and subtracting $\Hc\, \Phi_n^{(c)}$:
\begin{equation}
\label{eq:decomp}
  \Hc\, \Phi_n^{(\infty)}
  = \Hc(\Phi_n^{(\infty)} - \Phi_n^{(c)})
    + \Hc\, \Phi_n^{(c)}.
\end{equation}
Since $\Hc\, \Phi_n^{(c)} = \lambda_n^{(c)}\, \Phi_n^{(c)}$
(eigenvalue equation for $\Hc$), the second
term equals $\lambda_n^{(c)}\, \Phi_n^{(c)}$.

\textbf{Step 3.}
We show $\lambda_n^{(c)}\, \Phi_n^{(c)}
\to \lambda_n^{(\infty)}\, \Phi_n^{(\infty)}$ in $L^2$.
Since $\Phi_n^{(c)} \rightharpoonup \Phi_n^{(\infty)}$
weakly \eqref{XI2} and
$\|\Phi_n^{(c)}\| = 1 = \|\Phi_n^{(\infty)}\|$
(the latter by Lemma~\ref{lem:norm-pres}),
weak convergence plus norm convergence gives
$\Phi_n^{(c)} \to \Phi_n^{(\infty)}$ strongly in $L^2$.
Therefore:
\begin{align}
  \|\lambda_n^{(c)} \Phi_n^{(c)}
  - \lambda_n^{(\infty)} \Phi_n^{(\infty)}\|
  &\le |\lambda_n^{(c)} - \lambda_n^{(\infty)}|\,
       \|\Phi_n^{(c)}\|
     + \lambda_n^{(\infty)}\,
       \|\Phi_n^{(c)} - \Phi_n^{(\infty)}\|
  \to 0,
\end{align}
using \eqref{XI1} for the first term
and strong convergence for the second.

\textbf{Step 4.}
The first term of \eqref{eq:decomp} vanishes:
since $\|\Hc\| \le 1$ (contractivity, standard Slepian),
\[
  \|\Hc(\Phi_n^{(\infty)} - \Phi_n^{(c)})\|
  \le \|\Phi_n^{(\infty)} - \Phi_n^{(c)}\|
  \to 0
\]
by Step~3.

\textbf{Conclusion.}
Combining Steps~1--4:
\[
  \Hlim\, \Phi_n^{(\infty)}
  = \lim_{c \to \infty} \Hc\, \Phi_n^{(\infty)}
  = \lim_{c \to \infty} \lambda_n^{(c)}\, \Phi_n^{(c)}
  = \lambda_n^{(\infty)}\, \Phi_n^{(\infty)}.
\]
Since $\Phi_n^{(\infty)} \ne 0$
(Lemma~\ref{lem:nonvanish-limit}),
this is a genuine eigenvalue equation.
\end{proof}

\begin{remark}[Full independence from Paper~IX spectral structure]
\label{rem:no-specdef}
The proof of Lemma~\ref{lem:eigenvalue-eq} uses:
\begin{itemize}
\item \eqref{XI3}: SOT-convergence $\Hc \to \Hlim$
      (Paper~X, Lem.~2.2)
\item \eqref{XI1}: eigenvalue rate $c^{-1/4}$
      (Paper~VIII, Cor.~4.3)
\item \eqref{XI2}: weak convergence $\Phi_n^{(c)}\rightharpoonup\Phi_n^{(\infty)}$
      (Paper~IX, Prop.~2.2 --- subsequence existence only)
\item Lemma~\ref{lem:norm-pres}: norm preservation
      (Paper~VIII non-cancellation, no Paper~IX orthonormality)
\item $\|\Hc\| \le 1$: contractivity of $\Hc$
      (standard Slepian, Papers~V--VI)
\end{itemize}
It does \emph{not} use:
\begin{itemize}
\item Paper~IX Prop.~2.2 orthonormality assertion
\item Paper~IX Def.~2.3 (spectral series for $\Hlim$)
\item Completeness of $\{\Phi_n^{(\infty)}\}$
\item Any conditional result from Paper~IX
\end{itemize}
The sole Paper~IX dependency is the existence of a
weakly convergent subsequence (Prop.~2.2),
which is a standard compactness argument
(Fr\'echet--Kolmogorov; unit ball weakly compact in Hilbert space)
and does not require the full structure of Paper~IX.
\end{remark}

\begin{remark}[No completeness needed]
\label{rem:no-completeness-lem}
Lemmas~\ref{lem:norm-pres}--\ref{lem:eigenvalue-eq}
use only weak-limit existence (Paper~IX, Prop.~2.2),
SOT-convergence (Paper~X, Lem.~2.2), eigenvalue rates
(Paper~VIII, Cor.~4.3), and non-cancellation
(Paper~VIII, Thm.~3.3).
The completeness hypothesis \texttt{hyp:decay}
from Paper~IX is not invoked.
\end{remark}

\subsection{Main spectral inclusion theorem}

\begin{theorem}[Spectral inclusion]
\label{thm:spectral-inclusion}
\begin{enumerate}[\rm(a)]
\item For each $n \ge 0$,
      $\lambda_n^{(\infty)} \in \sigma_p(\Hlim)$
      (point spectrum; genuine eigenvalue).
\item $\overline{\{\lambda_n^{(\infty)} : n \ge 0\}}
      \;\subset\; \sigma(\Hlim)$.
\end{enumerate}
\end{theorem}

\begin{proof}
\textit{(a)} By Lemma~\ref{lem:eigenvalue-eq},
$\Hlim \Phi_n^{(\infty)} = \lambda_n^{(\infty)} \Phi_n^{(\infty)}$
with $\Phi_n^{(\infty)} \ne 0$. Hence
$\lambda_n^{(\infty)} \in \sigma_p(\Hlim)$.

\textit{(b)} The spectrum $\sigma(\Hlim)$ is closed
(standard; e.g.\ \cite{ReedSimon1975}, Theorem~VII.1).
Since $\sigma_p(\Hlim) \ni \lambda_n^{(\infty)}$
for all $n \ge 0$ (part (a)), the closure follows.
\end{proof}

\begin{remark}[Comparison with Paper~IX approach]
\label{rem:comparison-IX}
Paper~IX proves $\sigapp(\Hlim) = \sigma(\Hlim)$
only \emph{conditionally} on completeness
of $\{\Phi_n^{(\infty)}\}$ (via essential self-adjointness).
Theorem~\ref{thm:spectral-inclusion} gives the spectral
\emph{lower bound} $\overline{\{\lambda_n^{(\infty)}\}}
\subset \sigma(\Hlim)$ \emph{unconditionally},
using only the SOT-limit route which requires
no completeness.
The upper bound $\sigma(\Hlim) \subset [0,1]$
was already established unconditionally in
Paper~X, Cor.~\ref{cor:qlim-op}.
\end{remark}

%%═══════════════════════════════════════════════════════════════════
\section{Weyl-Type Counting Law}
\label{sec:weyl}
%%═══════════════════════════════════════════════════════════════════

\subsection{Definition and setup}

\begin{definition}[Counting function for limit eigenvalues]
\label{def:counting}
For $\lambda \in [0,1]$, define
\[
  N(\lambda)
  := \#\bigl\{n \ge 0 : \lambda_n^{(\infty)} \le \lambda\bigr\}.
\]
\end{definition}

\begin{remark}[Well-definedness]
\label{rem:counting-welldefined}
Since $\lambda_n^{(\infty)} \in (0,1)$ and
$\lambda_n^{(\infty)} \to 0$ as $n \to \infty$
(limit of Slepian eigenvalues, which decay
exponentially past the bandwidth threshold),
$N(\lambda) < \infty$ for all $\lambda \in [0,1)$
and $N(1) = +\infty$.
The function $\lambda \mapsto N(\lambda)$ is
non-decreasing and right-continuous.
\end{remark}

\subsection{Semi-classical approximation}

The classical Slepian--Pollak counting theorem
(see \cite{Slepian1961,Osipov2013,LandauPollak1962})
states that for fixed bandwidth $c$:
\begin{equation}
\label{eq:slepian-count}
  \#\bigl\{n : \lambda_n^{(c)} \in [\varepsilon, 1-\varepsilon]\bigr\}
  \;=\; O(\log c)
  \quad \text{for each fixed } \varepsilon > 0,
\end{equation}
and the transition from $\lambda \approx 1$ to
$\lambda \approx 0$ occurs in a window of width $O(\log c)$
centred at $n = 2c/\pi$.
Equivalently: for $n < (1-\delta)2c/\pi$,
$\lambda_n^{(c)} \ge 1 - e^{-\gamma_\delta c}$;
for $n > (1+\delta)2c/\pi$,
$\lambda_n^{(c)} \le e^{-\gamma_\delta c}$.

\subsection{Counting law for limiting eigenvalues}

\begin{theorem}[Weyl-type counting law]
\label{thm:weyl}
For each $\lambda \in (0,1)$, define
$c(\lambda)$ as the unique value satisfying
\begin{equation}
\label{eq:clambda}
  \lfloor 2c(\lambda)/\pi \rfloor
  \;=\; N(\lambda).
\end{equation}
Then:
\begin{enumerate}[\rm(a)]
\item $N(\lambda) \sim \dfrac{2\,c(\lambda)}{\pi}$
      as $\lambda \to 1^-$.
\item For $\lambda$ in the transition region
      $[\varepsilon, 1-\varepsilon]$:
      $N(\lambda)$ grows at most as $O(\log c(\lambda))$,
      consistent with \eqref{eq:slepian-count}.
\item $N(\lambda) = 0$ for all $\lambda < \lambda_0^{(\infty)}$,
      where $\lambda_0^{(\infty)} = \lim_{c \to \infty} \lambda_0^{(c)}$.
\end{enumerate}
\end{theorem}

\begin{proof}
\textit{(a)} Fix $\lambda \in (0,1)$ and $\delta > 0$.
By \eqref{XI6}, for all sufficiently large $c$,
$\lambda_n^{(c)} \ge \lambda$ iff $n \le (1-\delta)2c/\pi$
(up to the $O(\log c)$ transition window).
By the eigenvalue convergence rate \eqref{XI1},
$|\lambda_n^{(c)} - \lambda_n^{(\infty)}| \le C_\kappa c^{-1/4}$,
so the ordering of eigenvalues is preserved
in the limit:
$\lambda_n^{(\infty)} \ge \lambda$ iff
$n \le 2c(\lambda)/\pi + O(\log c(\lambda))$.
Hence $N(\lambda) = \lfloor 2c(\lambda)/\pi \rfloor
+ O(\log c(\lambda))$, giving part (a).

\textit{(b)} Follows directly from \eqref{eq:slepian-count}
and \eqref{XI1}: the $O(\log c)$ transition
window for $\lambda_n^{(c)}$ transfers to
$\lambda_n^{(\infty)}$ with at most $O(c^{-1/4})$
eigenvalue shift.

\textit{(c)} Since $\lambda_0^{(c)}$
is the largest Slepian eigenvalue and
$\lambda_0^{(c)} \to \lambda_0^{(\infty)}$
(which is $<1$ in the limit regime of interest),
no limiting eigenvalue lies below
$\lambda_0^{(\infty)}$.
\end{proof}

\begin{remark}[Interpretation]
\label{rem:weyl-interp}
Theorem~\ref{thm:weyl} is a semi-classical
statement: it expresses the counting function
of the \emph{limiting} operator via the
bandwidth parameter $c(\lambda)$ of the
approximating finite operators.
It is not a Weyl law in the classical sense
(since $\Hlim$ acts on $L^2(\mathbb{R})$,
which is not compact), but rather a
\emph{discretisation index} result.
A genuine Weyl asymptotic $N(\lambda) \sim C \lambda^\alpha$
for some $\alpha$ would require identifying the
distribution of $\{\lambda_n^{(\infty)}\}$
more precisely; this is related to
Conjecture~\ref{conj:density} below.
\end{remark}

%%═══════════════════════════════════════════════════════════════════
\section{Density Criterion and Spectral Structure}
\label{sec:density}
%%═══════════════════════════════════════════════════════════════════

\begin{theorem}[Density criterion]
\label{thm:density-criterion}
If $\{\lambda_n^{(\infty)} : n \ge 0\}$ is dense in $[0,1]$,
then $\sigma(\Hlim) = [0,1]$.
\end{theorem}

\begin{proof}
By Theorem~\ref{thm:spectral-inclusion}(a),
$\lambda_n^{(\infty)} \in \sigma(\Hlim)$ for all $n$.
The spectrum $\sigma(\Hlim)$ is closed.
If $\{\lambda_n^{(\infty)}\}$ is dense in $[0,1]$,
then $[0,1] \subset \overline{\{\lambda_n^{(\infty)}\}}
\subset \sigma(\Hlim)$.
The reverse inclusion $\sigma(\Hlim) \subset [0,1]$
was established unconditionally in Paper~X,
Cor.~\ref{cor:qlim-op}.
Hence $\sigma(\Hlim) = [0,1]$.
\end{proof}

\begin{corollary}[No isolated eigenvalues in the bulk]
\label{cor:no-isolated}
Under the hypothesis of Theorem~\ref{thm:density-criterion},
$\sigma(\Hlim)$ has no isolated points in $(0,1)$.
\end{corollary}

\begin{proof}
If $\sigma(\Hlim) = [0,1]$, every point of $(0,1)$
is a limit point of $\sigma(\Hlim)$;
hence no point in $(0,1)$ is isolated.
\end{proof}

\begin{remark}[Resolution of open problems from Paper~X]
\label{rem:open-problems-resolved}
Theorem~\ref{thm:density-criterion} and
Corollary~\ref{cor:no-isolated} resolve,
conditional on density of $\{\lambda_n^{(\infty)}\}$:
\begin{itemize}
\item \texttt{prob:spectrum}: Is $\sigma(\Hlim) = [0,1]$?
      \textbf{Yes}, if $\{\lambda_n^{(\infty)}\}$ is dense.
\item \texttt{prob:spec-structure}: Does $\sigma(\Hlim)$
      have isolated points in $(0,1)$?
      \textbf{No}, under the same hypothesis.
\end{itemize}
Both reduce to: is $\{\lambda_n^{(\infty)}\}$ dense in $[0,1]$?
\end{remark}

\begin{remark}[Structure without density]
\label{rem:without-density}
If $\{\lambda_n^{(\infty)}\}$ is \emph{not} dense
in $[0,1]$, $\sigma(\Hlim)$ could be a proper closed
subset with spectral gaps or Cantor-like structure.
Theorem~\ref{thm:spectral-inclusion} still guarantees
$\overline{\{\lambda_n^{(\infty)}\}} \subset \sigma(\Hlim)$.
\end{remark}

%%═══════════════════════════════════════════════════════════════════
\section{The Density Question}
\label{sec:density-question}
%%═══════════════════════════════════════════════════════════════════

\subsection{Statement}

\begin{conjecture}[Density of limiting eigenvalues]
\label{conj:density}
The sequence $\{\lambda_n^{(\infty)}\}_{n \ge 0}$
is dense in $[0,1]$.
\end{conjecture}

\subsection{Analytical evidence}

\subsubsection{From the Slepian transition window}

The classical Slepian--Pollak counting law \eqref{eq:slepian-count}
states that for each bandwidth $c$, there are
$O(\log c)$ eigenvalues $\lambda_n^{(c)}$ in any
fixed subinterval $[\varepsilon, 1-\varepsilon]$.
As $c \to \infty$, these transition eigenvalues
range over all of $[0,1]$ (the transition shifts with $c$).
By \eqref{XI1}, each $\lambda_n^{(c)}$ is within
$C_\kappa c^{-1/4}$ of $\lambda_n^{(\infty)}$.

A natural candidate strategy:
for any $\lambda_0 \in (0,1)$ and $\varepsilon > 0$,
choose $c$ large and $n \approx 2c/\pi$ such that
$|\lambda_n^{(c)} - \lambda_0| < \varepsilon/2$;
then $|\lambda_n^{(\infty)} - \lambda_0|
< C_\kappa c^{-1/4} + \varepsilon/2 < \varepsilon$
for $c$ sufficiently large.
This works if transition eigenvalues cover
$[\varepsilon, 1-\varepsilon]$ densely as $c$ varies.

\subsubsection{From the Airy profile (Paper~VI)}

The Airy-profile result (Paper~VI, Thm.~VI.3) shows
$\lambda_n^{(c)}$ varies continuously in $c$
for fixed $n$ near the transition region.
Since $\lambda_n^{(c)} \to 1$ as $n \to 0$ and
$\lambda_n^{(c)} \to 0$ as $n \to \infty$,
by the intermediate value theorem every value
in $(0,1)$ is attained by $\lambda_n^{(c)}$
for some $(n,c)$.
Whether these become dense in the limit is
the content of Conjecture~\ref{conj:density}.

\subsubsection{Partial result}

\begin{proposition}[Non-emptiness in $[1-\delta, 1)$]
\label{prop:partial-density}
For any $\delta \in (0,1)$,
$\{\lambda_n^{(\infty)}\} \cap [1-\delta, 1) \ne \emptyset$.
\end{proposition}

\begin{proof}
Direct from \eqref{XI6} and \eqref{XI1}.
\end{proof}

\begin{remark}[Gap in the argument]
\label{rem:density-gap}
Proposition~\ref{prop:partial-density} shows
non-emptiness but not density.
A full proof of Conjecture~\ref{conj:density}
requires showing transition eigenvalues
(those in the $O(\log c)$-wide window around $n = 2c/\pi$)
cover $[0,1]$ densely as $c \to \infty$,
with the $c^{-1/4}$ rate surviving the limit.
This requires stronger transition-zone control
than currently available from Papers~VI--VIII.
\end{remark}

\subsection{Numerical evidence}

The density of transition eigenvalues for finite $c$
is well-established numerically
(\cite{Slepian1961,Osipov2013,LandauPollak1962}).
The rate $|\lambda_n^{(c)} - \lambda_n^{(\infty)}|
\le C_\kappa c^{-1/4}$ (\eqref{XI1}) implies numerical
distributions for large $c$ reliably approximate
the limit distribution.
All evidence is consistent with density;
rigorous confirmation requires the analytical
argument outlined above.

%%═══════════════════════════════════════════════════════════════════
\section{Open Problems}
\label{sec:open-problems}
%%═══════════════════════════════════════════════════════════════════

\begin{problem}[Density of limiting eigenvalues]
\label{prob:density}
Prove Conjecture~\ref{conj:density}.
\emph{Approach}: control transition eigenvalues
$\lambda_n^{(c)}$ (in the $O(\log c)$-wide window
around $n = 2c/\pi$) as $c \to \infty$,
showing their limits cover $[0,1]$ densely.
\end{problem}

\begin{problem}[Sharpness of the form rate]
\label{prob:rate-sharp}
Is $\sup_f (q_{\lim}(f,f) - q_c(f,f)) / \|f\|^2
\asymp c^{-1/4}$?
Upper bound $O(c^{-1/4})$ is unconditional
(Paper~VIII, Cor.~4.3).
The matching lower bound would follow from the
sharp eigenvalue-gap lower bound
$|\lambda_n^{(c)} - \lambda_n^{(c+1)}| \ge c_\kappa c^{-3/4}$
(open since Paper~VIII, Remark~4.2).
\end{problem}

\begin{problem}[Completeness of $\{\Phi_n^{(\infty)}\}$]
\label{prob:completeness}
Prove $\{\Phi_n^{(\infty)}\}_{n \ge 0}$
complete in $L^2(\mathbb{R})$
(inherited from Papers~IX--X).
Equivalent to essential self-adjointness of $\Hlim$
and $\sigapp(\Hlim) = \sigma(\Hlim)$
(Paper~IX, Rem.~4.2).
\end{problem}

\begin{problem}[Riesz-basis stability]
\label{prob:riesz}
Is $\{\Phi_n^{(\infty)}\}_{n \ge 0}$ a Riesz basis
for its closed linear span in $L^2(\mathbb{R})$?
A positive answer would imply completeness
(Problem~\ref{prob:completeness}),
provide frame bounds for spectral reconstructions,
and yield the operator identification
$\HSOT = \overline{\Hspectral}$
(Paper~IX, Open Problem~7 / Paper~X, Hyp.~(IX.b)).
The non-cancellation lower bound
(Paper~VIII, Thm.~3.3) provides a
\emph{local frame seed} as starting point.
This problem is the primary target of Paper~XII.
\end{problem}

\begin{remark}[Relations between the open problems]
\label{rem:problems-relation}
Problems~\ref{prob:density} and~\ref{prob:completeness}
are logically independent.
Both are needed for a full unconditional
Hilbert--P\'olya realization:
completeness closes the $\sigapp = \sigma$ gap;
density (with CCM) gives a dense set of zeros
on the critical line.
Problem~\ref{prob:riesz} (Riesz-basis stability)
would resolve both simultaneously and is the
central target of Paper~XII.
\end{remark}

%%═══════════════════════════════════════════════════════════════════
\section{Outlook: Reduction of the Hilbert--P\'olya Problem}
\label{sec:outlook}
%%═══════════════════════════════════════════════════════════════════

The results of Papers~I--XI reduce the
Hilbert--P\'olya question to two independent
spectral-analytic problems:

\medskip
\begin{enumerate}
\item \textbf{Density} (Conjecture~\ref{conj:density}):
      Is $\{\lambda_n^{(\infty)}\}$ dense in $[0,1]$?
      If yes, $\sigma(\Hlim) = [0,1]$ (this paper).
\item \textbf{Completeness} (\texttt{prob:completeness},
      Papers~IX--X):
      Is $\{\Phi_n^{(\infty)}\}$ complete in $L^2(\mathbb{R})$?
      If yes, $\Hlim$ is essentially self-adjoint
      and $\sigapp(\Hlim) = \sigma(\Hlim)$ (Paper~IX).
\end{enumerate}

\medskip\noindent
Under both, combined with Hypothesis~CCM
(Paper~VIII, Hyp.~5.1; \cite{CCM2025}):
all non-trivial zeros $\rho$ of $\zeta(s)$ satisfy
$\mathrm{Re}(\rho) = 1/2$.

The two problems are independent of each other
and of any hypothesis on $\zeta(s)$.

\medskip
A third, deeper layer is the operator identification
$\HSOT = \overline{\Hspectral}$
(Paper~IX, Open Problem~7;
Paper~X, Hyp.~(IX.b)).
This identification --- which would simultaneously
resolve Riesz-basis stability,
completeness, and the Bridge theorem of Paper~X ---
is the primary target of Paper~XII.
Paper~XII will analyse frame-theoretic obstructions
using the non-cancellation seed from Paper~VIII
(Thm.~3.3) as a local lower frame bound,
and isolate the uniform-in-$n$ control problem
as the key gap.

\medskip\noindent
This is the precise state of the
Hilbert--P\'olya programme within the
present framework as of Paper~XI.

%%%─── Bibliography
\begin{thebibliography}{99}

\bibitem{PaperVIII}
[Author],
\textit{Non-cancellation control, the sharp
$c^{-1/2}$ lower bound, and spectral connection
to the zeros of $\zeta(s)$},
Paper~VIII of this series, preprint, 2026.

\bibitem{PaperIX}
[Author],
\textit{Domain identification, deficiency indices,
and essential self-adjointness of the PSWF
concentration limit operator},
Paper~IX of this series, preprint, 2026.

\bibitem{PaperX}
[Author],
\textit{Mosco convergence, resolvent stability,
and the bridge theorem for PSWF concentration operators},
Paper~X of this series, preprint, 2026.

\bibitem{Slepian1961}
D.~Slepian and H.~O.~Pollak,
\textit{Prolate spheroidal wave functions,
Fourier analysis and uncertainty~I},
Bell System Tech.\ J.\ \textbf{40} (1961), 43--63.

\bibitem{LandauPollak1962}
H.~J.~Landau and H.~O.~Pollak,
\textit{Prolate spheroidal wave functions,
Fourier analysis and uncertainty~III:
The dimension of the space of essentially
time- and band-limited signals},
Bell System Tech.\ J.\ \textbf{41} (1962), 1295--1336.

\bibitem{Osipov2013}
A.~Osipov, V.~Rokhlin, and H.~Xiao,
\textit{Prolate Spheroidal Wave Functions of Order Zero},
Springer, New York, 2013.

\bibitem{Kato1966}
T.~Kato,
\textit{Perturbation Theory for Linear Operators},
Springer, Berlin, 1966.

\bibitem{ReedSimon1975}
M.~Reed and B.~Simon,
\textit{Methods of Modern Mathematical Physics,
Vols.~I--II},
Academic Press, New York, 1972--1975.

\bibitem{Mosco1969}
U.~Mosco,
\textit{Convergence of convex sets and of solutions
of variational inequalities},
Adv.\ Math.\ \textbf{3} (1969), 510--585.

\bibitem{KuwaeShioya2003}
K.~Kuwae and T.~Shioya,
\textit{Convergence of spectral structures},
Comm.\ Anal.\ Geom.\ \textbf{11} (2003), 599--673.

\bibitem{CCM2025}
A.~Connes, C.~Consani, and H.~Moscovici,
\textit{Zeta Spectral Triples},
arXiv:2511.22755 [math.NT], preprint, 2025.

\end{thebibliography}

\end{document}
