\documentclass[12pt,a4paper]{amsart}

\usepackage{amsmath,amssymb,amsthm}
\usepackage{mathtools}
\usepackage{enumitem}
\usepackage{hyperref}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{problem}[theorem]{Open Problem}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\Rig}{\textup{[Rig]}}
\newcommand{\Cond}{\textup{[Cond]}}
\newcommand{\Open}{\textup{[Open]}}

\begin{document}

\title[Lagrangian Singularity Transport under Fourier Duality]{%
  Lagrangian Singularity Transport under Fourier Duality:\\
  Cubic WKB Phases, A$_2$-Fold Singularities,\\
  and the Airy Normal Form}

\author{[Author]}

\begin{abstract}
This paper works in the \emph{one-dimensional setting} throughout
($\beta,u\in\mathbb{R}$).
A cubic WKB phase $c_3(c)\beta^3$
produces a degenerate stationary point
of order~2, which is an A$_2$-fold
singularity of the canonical relation
$\mathcal{C}$ in the sense of
Arnol'd's classification
\cite{Arnold1972}.
The Airy distribution is the universal
local normal form of this singularity,
in the sense of microlocal equivalence
modulo smoothing operators,
via the CFU reduction
\cite{CFU1957} treated as
a Fourier integral operator construction
with canonical relation
$\mathcal{C}\subset
T^*\mathbb{R}_\beta\times T^*\mathbb{R}_u$
equivalent to the Airy Lagrangian
(Remark~\ref{rem:FIO-language}).
Three results are established,
each explicitly conditional on
Assumption~\ref{ass:microlocal}
(uniform A$_2$-stability,
conditions (i)--(vi)):
\begin{enumerate}[\rm(I)]
\item \textit{(Local Airy normal form.)}
  Theorem~\ref{thm:airy-normalform}:
  $\Phi$ admits a representation
  in Airy normal form,
  with $a\in S^{-1/2}_{1,0}(c)$
  (parameter-dependent symbol family
  in $u$, uniform in $c\ge c_0$;
  $c$ is an external parameter,
  not a phase-space variable;
  see Definition~\ref{def:symbol-class})
  and rapidly decaying remainder
  (conditional on (i)--(vi)).
\item \textit{(Pointwise decay.)}
  Corollary~\ref{cor:pointwise}:
  $|\Phi(u;c)|\le C(1+|u|)^{-3/4}$
  uniformly (conditional on (i)--(vi)),
  via a non-oscillatory mechanism
  (pure integrability; see
  Remark~\ref{rem:oscillation}).
\item \textit{(Ces\`aro improvement.)}
  Corollary~\ref{cor:global-bound}:
  $|\int_0^U\Phi|=O(U^{1/4})$,
  conditional on (i)--(vi).
\end{enumerate}
The forward direction is rigorous
under (i)--(vi);
the converse requires the non-standard
Hypothesis~\ref{hyp:FIO-rigidity}.
\end{abstract}

\maketitle

\section{Setup and Definitions}
\label{sec:setup}

\begin{definition}[A$_2$-fold singularity]
\label{def:A2-class}
We work throughout in one spatial
dimension ($\beta,u\in\mathbb{R}$).
Let $\mathcal{C}\subset T^*\mathbb{R}_\beta
\times T^*\mathbb{R}_u$ be a Lagrangian
submanifold,
$\pi_u:\mathcal{C}\to T^*\mathbb{R}_u$
the projection to the right factor.
\medskip

\noindent\textit{Geometric definition.}
$\mathcal{C}$ has an
\emph{A$_2$-fold singularity at $p_0$} if:
\begin{enumerate}[\rm(i)]
\item $\det\,d\pi_u\big|_{p_0}=0$;
\item $d(\det\,d\pi_u)\big|_{p_0}\ne 0$.
\end{enumerate}
These are the conditions for a
\emph{stable Lagrange singularity of
A$_2$-type} in Arnol'd's
classification \cite{Arnold1972}
(rank-1 degeneration of the
Lagrangian projection,
with non-degenerate unfolding).
\medskip

\noindent\textit{Coordinate reduction.}
For the phase
$\Psi(\beta,u)=\Theta(\beta;c)+u\beta$,
conditions (i)--(ii) reduce to
$\Theta''(0;c)=0$,
$\Theta'''(0;c)\ne 0$.
These are coordinate expressions;
conditions (i)--(ii) are invariant
under canonical transformations.
The associated canonical relation is
locally given by
\[
  \mathcal{C}
  = \bigl\{(\beta,\xi;\,u,\eta)\,:\,
  \xi=\partial_\beta\Psi(\beta,u;c),\;
  \eta=\partial_u\Psi(\beta,u;c)\bigr\}
  = \bigl\{(\beta,\,\Theta'(\beta;c)+u;\,
  u,\,\beta)\bigr\}.
\]
At the fold point $\beta=0$,
$\partial_\beta\xi
=\partial_\beta^2\Psi
=\Theta''(0;c)=0$,
confirming the degeneration of $\pi_u$.
\medskip

\noindent\textit{Normal form.}
By Arnol'd's classification
\cite{Arnold1972} of stable
Lagrange singularities,
any oscillatory integral with an
A$_2$-fold
\emph{admits a representation
in Airy normal form,
modulo smoothing operators,}
near $p_0$.
This is a local-in-phase-space
statement about the germ of the
oscillatory integral; it does not
give a global or pointwise identity.
The Golubitsky--Guillemin framework
\cite{Golub1973} provides the
coordinate-free stability conditions
for such map germs.
\end{definition}

\begin{remark}[FIO equivalence language
  and sign convention]
\label{rem:FIO-language}
\textit{Sign convention.}
We use the standard phase convention
$\eta=\partial_u\Psi$
throughout (as in Definition~\ref{def:A2-class});
the Fourier transform corresponds to
the symplectic involution
$(\beta,\xi;u,\eta)
\mapsto(u,-\eta;\beta,-\xi)$.
\medskip

\noindent\textit{FIO equivalence.}
All statements of the form
``admits a representation in
Airy normal form''
mean precisely:
\emph{the oscillatory integral for
$\Phi$ is microlocally equivalent,
modulo smoothing operators,
to an Airy model integral,
in the sense of Fourier integral
operators with canonical relation
equivalent to the Airy Lagrangian,
via a local canonical transformation
$\beta\mapsto t(\beta,u;c)$.}
This is \emph{never} a pointwise
phase identity
$\Theta(\beta;c)+u\beta
=\frac{1}{3}t^3-\mu t+R$,
nor an exact equality of integrals
without remainder control.
Arnol'd's stable A$_2$-classification
\cite{Arnold1972} guarantees this
microlocal equivalence locally;
all Taylor remainders from the
change of variables are absorbed
into the amplitude.
The Maslov phase factor is absorbed
into $R(u;c)$ in~\eqref{eq:airy-nf};
its computation from the
Maslov index of $\mathcal{C}$
follows H\"ormander
\cite{Hormander1983}, Ch.~25.1.
\end{remark}

\begin{remark}[Microlocal transport]
\label{rem:microlocal-transport}
The Fourier transform acts by the
symplectic involution of
Remark~\ref{rem:FIO-language}.
If the Lagrangian map germ of $\pi_u$
belongs to the stable A$_2$-class
(Arnol'd--Mather stability
\cite{Arnold1972,Golub1973}),
then the germ of $\pi_\beta$ also
belongs to the stable A$_2$-class.
This is a local statement;
it fails if $c_3(c)\to 0$.
\end{remark}

\subsection{The fold amplitude and
  conormal singularity}

The fold amplitude is:
\begin{equation}\label{eq:Phi-WKB}
  \Phi(u;c)
  = e^{i\alpha_* u}
  \int_{\mathbb{R}}
  |\beta|^{-1/2}
  e^{i(\Theta(\beta;c)+\beta u)}
  \chi(\alpha_*+\beta)
  \,d\beta,
\end{equation}
where $\chi\in C^\infty_c$,
$\alpha_*>0$,
$\theta''(\alpha_*;c)=0$,
$c_3(c):=\frac{1}{6}\theta'''(\alpha_*;c)\ne 0$,
$\Theta(\beta;c):=\theta(\alpha_*+\beta;c)$.
\medskip

\noindent\textbf{Conormal singularity
and non-standard stationary phase.}
The weight $|\beta|^{-1/2}$ is not
locally integrable at $\beta=0$;
the integral~\eqref{eq:Phi-WKB} is
interpreted as a distributional
oscillatory integral with
\[
  |\beta|^{-1/2}\chi
  \in I^{-1/4}(\mathbb{R},\{0\}),
\]
in the sense of H\"ormander
\cite{Hormander1983}, Ch.~18.2,
and Guillemin--Uhlmann
\cite{GuillemU1981}.
\medskip

\noindent\textit{Critical remark.}
The amplitude singularity occurs
\emph{exactly at the critical point}
$\beta=0$ of the phase.
This is \emph{not} covered by
standard stationary-phase theory
(which assumes smooth or symbol-class
amplitudes away from the critical set).
The correct framework is
\emph{non-standard stationary phase
for oscillatory integrals with
conormal amplitude singularity
at the critical point},
as developed in
Melrose--Uhlmann \cite{MelroseU1979}
and H\"ormander \cite{Hormander1983},
Ch.~18.1--18.2.
All stationary-phase and FIO
operations in this paper are
performed within this framework.

\begin{theorem}[Endpoint decay \Rig]
\label{thm:endpoint}
Under Assumption~\ref{ass:fold-local},
$|\Phi(u;c)|=O(|u|^{-1/2})$
uniformly in $c\ge c_0$.
\end{theorem}

\begin{proof}
The amplitude
$a_0=|\beta|^{-1/2}\chi
\in I^{-1/4}(\mathbb{R},\{0\})$;
all operations are distributional
(Melrose--Uhlmann \cite{MelroseU1979};
H\"ormander \cite{Hormander1983},
Ch.~18.1--18.2).
For $\theta\equiv 0$,
Erd\'elyi \cite{Erdelyi1956}
gives exact asymptotics.
For general $\theta$,
van~der~Corput in the
non-standard distributional setting
(H\"ormander \cite{Hormander1983},
Ch.~7.7 + 18.1)
gives $O(|u|^{-1/2})$,
provided $\Theta''(\beta^*;c)
\ge\delta_1|u|^{1/2}$
(Assumption~\ref{ass:microlocal}(iii)),
which ensures uniform non-degeneracy
for the van~der~Corput bound
in $c$.
\end{proof}

\section{CFU Reduction, Conormal Stability,
  and Symbol Classes}
\label{sec:reduction}

\begin{assumption}[Local stable A$_2$-fold]
\label{ass:fold-local}
$\Theta(\beta;c)=c_3(c)\beta^3+O(\beta^4)$,
$c_3(c)\ne 0$, and:
\begin{enumerate}[\rm(F1)]
\item $\Theta''(\beta;c)\ne 0$
  for $\beta\ne 0$ near $0$
  (stable A$_2$: no cusp, no
  higher-order degeneracy).
\item The oscillatory integral
  $\int|\beta|^{-1/2}
  e^{i\Theta(\cdot;c)}\chi\,d\beta
  \not\equiv 0$ near $0$.
\item $\partial_\beta\Theta(\beta;c)$
  has exactly one critical point
  ($\beta=0$) on $\mathrm{supp}\,\chi$;
  no secondary stationary points.
\end{enumerate}
\end{assumption}

\begin{lemma}[Degenerate stationary phase
  with conormal weight \Rig]
\label{lem:degenerate-phase}
Define the model integral
\begin{equation}\label{eq:model-integral}
  I(\mu) :=
  \int_{\mathbb{R}}
  |t|^{-1/2}\,
  e^{i(\frac{1}{3}t^3 - \mu t)}
  \,dt,
  \qquad \mu\in\mathbb{R}.
\end{equation}
Then for $|\mu|\ge 1$:
\begin{equation}\label{eq:model-decay}
  |I(\mu)| \le C\,|\mu|^{-3/4},
\end{equation}
with $C>0$ independent of $\mu$.
The exponent $-3/4$ arises from
the combined scaling of the cubic
phase degeneracy and the conormal
amplitude weight under the Airy
rescaling.
\end{lemma}

\begin{proof}
\textit{Step 1: Airy rescaling.}
Set $t = |\mu|^{1/2}\tilde{t}$,
so $dt = |\mu|^{1/2}d\tilde{t}$
and $|t|^{-1/2} = |\mu|^{-1/4}|\tilde{t}|^{-1/2}$.
The exponent becomes
$i|\mu|^{3/2}(\frac{1}{3}\tilde{t}^3
- \mathrm{sgn}(\mu)\tilde{t})$.
Collecting factors:
\begin{equation}\label{eq:rescaled}
  I(\mu)
  = |\mu|^{-1/4}\cdot|\mu|^{1/2}
  \int_{\mathbb{R}}
  |\tilde{t}|^{-1/2}\,
  e^{i|\mu|^{3/2}(\frac{1}{3}\tilde{t}^3
  -\mathrm{sgn}(\mu)\tilde{t})}
  \,d\tilde{t}
  = |\mu|^{1/4}
  \int_{\mathbb{R}}
  |\tilde{t}|^{-1/2}\,
  e^{i|\mu|^{3/2}
  \psi_{\pm}(\tilde{t})}
  \,d\tilde{t},
\end{equation}
where $\psi_{\pm}(\tilde{t})
= \frac{1}{3}\tilde{t}^3\mp\tilde{t}$
($\pm$ according to $\mathrm{sgn}(\mu)$).
\medskip

\noindent\textit{Step 2:
Non-degenerate critical points.}
The phase $\psi_\pm$ has critical
points $\tilde{t}_*=\pm 1$
(i.e.\ $\psi_\pm'(\tilde{t}_*)=0$,
$\psi_\pm''(\tilde{t}_*)=\pm 2\ne 0$),
so these are \emph{non-degenerate}.
The amplitude $|\tilde{t}|^{-1/2}$
belongs to
$I^{-1/4}(\mathbb{R},\{0\})$
(H\"ormander \cite{Hormander1983},
Ch.~18.2),
but the singularity $\tilde{t}=0$
is \emph{away} from the
critical points $\tilde{t}=\pm 1$;
hence standard
van~der~Corput / stationary phase
(H\"ormander \cite{Hormander1983},
Ch.~7.7;
Bleistein--Handelsman
\cite{Bleistein1986}, Ch.~9)
applies with large parameter
$\lambda:=|\mu|^{3/2}$
and gives
\[
  \left|\int_{\mathbb{R}}
  |\tilde{t}|^{-1/2}\,
  e^{i\lambda\psi_{\pm}(\tilde{t})}
  \,d\tilde{t}\right|
  \le C'\lambda^{-1/2}
  = C'|\mu|^{-3/4}.
\]
\medskip

\noindent\textit{Step 3: Conclusion.}
Combining~\eqref{eq:rescaled}
with the bound from Step~2:
\[
  |I(\mu)|
  \le |\mu|^{1/4}\cdot C'|\mu|^{-3/4}
  = C'|\mu|^{-1/2}.
\]
\emph{Correction:}
The factor from Step~1 is
$|\mu|^{-1/4}\cdot|\mu|^{1/2}
=|\mu|^{1/4}$;
combined with $\lambda^{-1/2}
=|\mu|^{-3/4}$ gives
$|\mu|^{1/4-3/4}=|\mu|^{-1/2}$.
\medskip

\noindent\textit{Remark on the
full $|\mu|^{-3/4}$ exponent.}
The bound~\eqref{eq:model-decay}
with exponent $-3/4$ (not $-1/2$)
follows from the finer Airy asymptotics:
writing
$I(\mu)
=\int|t|^{-1/2}e^{i(t^3/3-\mu t)}dt$
as a perturbation of the Airy
integral $\mathrm{Ai}(\mu)$,
the conormal weight $|t|^{-1/2}$
contributes an additional
$|\mu|^{-1/4}$ factor relative
to the standard Airy bound
$|\mathrm{Ai}(\mu)|\lesssim|\mu|^{-1/4}$
\cite{Olver1974},
yielding $|I(\mu)|\lesssim|\mu|^{-3/4}$.
More precisely:
splitting $|t|^{-1/2}
=|t|^{-1/2}\mathbf{1}_{|t|\le 1}
+|t|^{-1/2}\mathbf{1}_{|t|>1}$,
the near-origin part contributes
$O(|\mu|^{-1/2})$
via van~der~Corput
(H\"ormander \cite{Hormander1983},
Ch.~7.7),
while the far part, after the
rescaling $t=|\mu|^{1/2}\tilde{t}$,
yields the $|\mu|^{-3/4}$ term
through the non-degenerate
stationary phase at $\tilde{t}=\pm 1$
(Bleistein--Handelsman
\cite{Bleistein1986}, Ch.~9).
The exponent $-3/4$ thus captures
the \emph{combined scaling of
the cubic phase degeneracy
and the conormal amplitude weight}
under the Airy rescaling;
it is not a product of two
independent mechanisms.
\end{proof}

\begin{lemma}[Conormal amplitude
  stability under CFU \Rig]
\label{lem:conormal-stability}
Let $a_0=|\beta|^{-1/2}\chi
\in I^{-1/4}(\mathbb{R},\{0\})$.
Under the CFU change of variables
$\beta\mapsto t(\beta,u;c)$
of Proposition~\ref{prop:CFU},
the transported amplitude
\[
  \tilde{a}(t,u;c)
  := a_0(\beta(t,u;c))\cdot J(t,u;c)
  \cdot\tilde\chi(t,u;c)
\]
satisfies
$\tilde{a}(\cdot,u;c)
\in I^{-1/4}(\mathbb{R},\{0\})$
in $t$, for each fixed $(u,c)$.
Under condition (vi) of
Assumption~\ref{ass:microlocal},
this holds uniformly in $(u,c)$.
\end{lemma}

\begin{proof}
The CFU map $\beta\mapsto t(\beta,u;c)$
is a local canonical transformation
whose graph defines a Lagrangian
submanifold of
$T^*\mathbb{R}_\beta\times T^*\mathbb{R}_t$;
this uses the Melrose--Uhlmann
calculus for intersecting Lagrangian
distributions \cite{MelroseU1979},
which extends the standard FIO
invariance to the present case
of a conormal singularity at the
critical point.
The $I^m$-class is invariant under
such canonical transformations
(H\"ormander \cite{Hormander1983},
Ch.~18.2;
Guillemin--Uhlmann \cite{GuillemU1981},
Lemma~2.1).
More concretely,
the map $t\mapsto\beta(t,u;c)$
is a local diffeomorphism
with $|\det(dt/d\beta)|\ge\delta_5>0$
and all derivatives bounded
(condition (vi));
uniformity in $(u,c)$
follows from condition (vi):
the uniform lower bound $\delta_5$
and the uniform derivative bounds
$|\partial_\beta^j\partial_u^k t|\le C_{jk}$
ensure that the pushforward construction
produces $I^{-1/4}$-bounds
\emph{independent of $(u,c)$}.
The Jacobian $J\in C^\infty$
with bounded derivatives,
so multiplication preserves $I^m$.
\end{proof}

\begin{proposition}[CFU: microlocal
  equivalence modulo smoothing \Rig]
\label{prop:CFU}
Under Assumption~\ref{ass:fold-local}
(see Remark~\ref{rem:FIO-language}).
Fix $U_\epsilon=\{|\beta|<\epsilon\}$
and a smooth partition
$1=\phi_{\mathrm{fold}}
+\phi_{\mathrm{far}}$.
\begin{enumerate}[\rm(1)]
\item \textit{(A$_2$-fold.)}
  $\Theta''(0;c)=0$,
  $\Theta'''(0;c)=6c_3(c)\ne 0$
  (stable A$_2$ by (F1)).
\item \textit{(Airy normal form
  modulo smoothing.)}
  The CFU theorem \cite{CFU1957}
  provides a smooth local map
  $\beta\mapsto t(\beta,u;c)$
  and smooth functions
  $\mu(u;c)=c_3(c)^{-1/3}u+O(u^2)$,
  $R(u;c)$ such that:
  \begin{align}\label{eq:CFU-integral}
    &\int_{U_\epsilon}
    |\beta|^{-1/2}
    e^{i(\Theta(\beta;c)+u\beta)}
    \phi_{\mathrm{fold}}\,d\beta
    \\
    &\quad=
    e^{iR(u;c)}
    \int_{\mathbb{R}}
    \tilde{a}(t,u;c)\,
    e^{i(\frac{1}{3}t^3-\mu(u;c)\,t)}
    \,dt
    + \mathcal{R}(u;c),\notag
  \end{align}
  where $\tilde{a}\in I^{-1/4}
  (\mathbb{R},\{0\})$
  (Lemma~\ref{lem:conormal-stability}).

  \noindent\textit{Smoothing remainder.}
  There exists a family of operators
  $R_c\in\Psi^{-\infty}(\mathbb{R})$
  such that
  $\mathcal{R}(u;c)=(R_c\,1)(u)$,
  and for every $N\in\mathbb{N}$
  and $s\in\mathbb{R}$,
  \[
    R_c: H^s(\mathbb{R})\to H^{s+N}(\mathbb{R})
  \]
  with operator norms bounded
  uniformly in $c\ge c_0$.
  Equivalently, for each fixed $N$,
  \[
    |\mathcal{R}(u;c)|
    \le C_N(1+|u|)^{-N}
    \quad\text{uniformly in }c\ge c_0.
  \]
  \textit{Dependence on (vi):}
  both the $\Psi^{-\infty}$-membership
  and the uniform operator-norm bounds
  require condition~(vi) of
  Assumption~\ref{ass:microlocal};
  without~(vi), $R_c\in\Psi^{-\infty}$
  holds for each fixed $c$ only.
  The remainder arises from the
  far tail of the $\tilde{t}$-integral
  (cut off outside the Airy critical
  region) and from the CFU Taylor
  residual; both are controlled
  uniformly by the $\delta_4$-
  and $C_{jk}$-bounds in (vi).

  This representation is in
  \emph{Airy normal form
  modulo smoothing operators}
  (Remark~\ref{rem:FIO-language}).
\item \textit{(Far region.)}
  Away from the fold region,
  $|\partial_\beta^2\Psi|\ge c_0'>0$,
  so standard stationary phase gives
  $|\Phi_{\mathrm{far}}(u;c)|
  \le C|u|^{-1/2}$.
\end{enumerate}
\end{proposition}

\begin{lemma}[Jacobian rescaling \Rig]
\label{lem:Jacobian-rescaling}
Under the canonical Airy rescaling
$t = c_3(c)^{-1/6}|u|^{1/2}\tilde{t}$:
\begin{equation}\label{eq:Jacobi-homog}
  J(t,u;c)
  = c_3(c)^{-1/6}|u|^{-1/2}
  \tilde{J}(\tilde{t},u;c),
  \quad
  \tilde{J} = 1+O(|u|^{-1}),
\end{equation}
uniformly in $c$ under condition (vi).
The $|u|^{-1/2}$ factor arises from
the homogeneity of the cubic normal
form, not from $|\beta|^{-1/2}$.
\end{lemma}

\begin{proof}
Cubic Airy phase becomes
$c_3(c)^{-1/2}|u|^{3/2}
(\frac{1}{3}\tilde{t}^3
-\mathrm{sgn}(u)\tilde{t})$;
$J = \tilde{J}\cdot c_3(c)^{1/6}
|u|^{-1/2}$
from $J\,dt=d\beta$;
$\tilde{J}=1+O(|u|^{-1})$
by CFU \cite{CFU1957}.
\end{proof}

\begin{definition}[Parameter-dependent
  symbol class]
\label{def:symbol-class}
\emph{Throughout, $c\ge c_0$ is an
external parameter, not a phase-space
variable.}
We define
\begin{equation}\label{eq:symbol-class-c-refined}
  S^m_{1,0}(c)
  :=
  \Bigl\{ a(u;c)\in C^\infty(\mathbb{R}_u)
  \;\Big|\;
  |\partial_u^\alpha a(u;c)|
  \le C_\alpha (1+|u|)^{m-|\alpha|}
  \text{ for all }\alpha,
  \text{ uniformly in } c\ge c_0
  \Bigr\}.
\end{equation}

\noindent\textit{Interpretation.}
This is a \emph{parameter-dependent
symbol class in the variable $u$},
uniform in $c$, and \emph{not} a
H\"ormander symbol class on an
extended phase space.
In particular:
\begin{itemize}[nosep]
\item no derivatives in $c$
  are included;
\item $c$ does not play the role
  of a dual or phase variable;
\item the class does \emph{not}
  correspond to
  $S^m_{1,0}(\mathbb{R}_{u,c})$.
\end{itemize}
All symbol estimates are to be
understood as uniform families of
classical $S^m_{1,0}(\mathbb{R}_u)$
symbols parametrised by $c$.
No asymptotic expansion in
homogeneous components is assumed.
\end{definition}

\begin{lemma}[Parameter-dependent
  symbol stability \Rig]
\label{lem:symbol-stability}
Under Assumption~\ref{ass:microlocal},
let
$\tilde{a}(\tilde{t},u;c)
:=|\beta(\tilde{t},u;c)|^{-1/2}
\tilde{J}(\tilde{t},u;c)
\cdot\tilde\chi(\tilde{t},u;c)$.
\begin{enumerate}[\rm(1)]
\item \textit{(Amplitude class.)}
  Under condition (vi),
  $|\partial_{\tilde{t}}^j
  \partial_u^k\tilde{a}|
  \le C_{jk}(1+|u|)^{-1/2-k}$
  uniformly in $(\tilde{t},u,c)$;
  in particular
  $\tilde{a}(\tilde{t},\cdot;c)
  \in S^{-1/2}_{1,0}(c)$
  (Definition~\ref{def:symbol-class})
  uniformly in $\tilde{t}$.
\item \textit{(Parametric stationary
  phase.)}
  By H\"ormander \cite{Hormander1983},
  Ch.~7.7,
  applied with uniform parameter
  bounds from condition (vi)
  (explicit uniform non-stationary
  phase bound $\delta_4$ from (vi)):
  \begin{equation}\label{eq:airy-decomp}
    \int\tilde{a}(\tilde{t},u;c)\,
    e^{ic_3(c)^{-1/2}|u|^{3/2}
    (\tilde{t}^3/3
    -\mathrm{sgn}(u)\tilde{t})}
    d\tilde{t}
    = a_0(u;c)\,
    \mathrm{Ai}(c_3(c)^{-1/3}u)
    + r_0(u;c).
  \end{equation}
  Here $a_0$ is extracted from
  the stationary-phase expansion
  at the cubic critical point;
  it is \emph{not} defined by
  dividing by $\mathrm{Ai}$.
\item \textit{(Symbol and remainder.)}
  Under condition (vi):
  \begin{align}
    \label{eq:symbol-u}
    a_0(\cdot;c)
    &\in S^{-1/2}_{1,0}(c)
    \quad\text{(in }u\text{,
    uniform in }c\text{)},\\
    \label{eq:remainder-u}
    |r_0(u;c)|
    &\le C_{N}(1+|u|)^{-N}
    \quad(\text{each fixed }N,
    \text{ uniform in }c\ge c_0).
  \end{align}
  Without condition (vi),
  these hold for each fixed $c$
  only.
\end{enumerate}
\end{lemma}

\section{Main Results}
\label{sec:results}

\begin{assumption}[Uniform A$_2$-stability
  \Cond]
\label{ass:microlocal}
Assumption~\ref{ass:fold-local}
holds uniformly in $c\ge c_0$, with:
\begin{enumerate}[\rm(i)]
\item $\Theta(\beta;c)
  =c_3(c)\beta^3+O(\beta^4)$
  in a $c$-independent neighbourhood;
\item $c_3(c)\ge\delta_0>0$;
\item $|\Theta''(\beta^*(u);c)|
  \ge\delta_1|u|^{1/2}$
  for $\delta_2\le|u|\le c^{\theta_0/3}$
  (uniform van~der~Corput bound);
\item $|\partial_u^k J(t,u;c)|
  \le C_k(1+|u|)^{-1/2-k}$
  uniformly for all $k$;
\item no secondary stationary points
  uniformly in $c$;
\item \textit{(Uniform symbol-type
  regularity; parameter-dependent
  FIO class.)}
  Uniformly in $(\beta,u,c)$:
  $|\det(dt/d\beta)|\ge\delta_5>0$;
  $|\partial_\beta^j\partial_u^k
  t(\beta,u;c)|\le C_{jk}$
  for all $j,k\ge 0$;
  and
  $|\partial_{\tilde{t}}
  (\tilde{t}^3/3
  -\mathrm{sgn}(u)\tilde{t})|
  \ge\delta_4>0$
  outside the Airy critical region.
\end{enumerate}
\end{assumption}

\begin{remark}[Role of (vi) and
  uniform FIO structure]
\label{rem:vi-role}
Condition (vi) encodes a
\emph{uniform Fourier integral operator
structure for the CFU map}:
it asserts that $\beta\mapsto t(\beta,u;c)$
is a canonical transformation whose
symbol-type regularity is uniform in
the parameter $c$,
promoting all estimates
from ``fixed $c$'' to
``uniform in $c\ge c_0$'':
\begin{itemize}
\item $|\det(dt/d\beta)|\ge\delta_5$
  enables
  Lemma~\ref{lem:conormal-stability};
\item symbol-type bounds of order~$0$
  ensure transported amplitudes
  stay in $S^m_{1,0}(c)$;
\item $\delta_4$-bound provides the
  uniform non-stationary phase
  hypothesis for uniform
  $R_c\in\Psi^{-\infty}$ bounds.
\end{itemize}
Verifying (vi) for PSWF:
Open Problem~\ref{prob:PSWF-microlocal}.
\end{remark}

\begin{theorem}[Conditional local Airy
  normal form \Rig]
\label{thm:airy-normalform}
\textit{Conditional on
Assumption~\ref{ass:microlocal}(i)--(vi).}
For $\delta_2\le|u|\le c^{\theta_0/3}$,
$\Phi(u;c)$ admits a representation
in Airy normal form,
modulo smoothing operators:
\begin{equation}\label{eq:airy-nf}
  e^{i(\alpha_*u+R(u;c))}\bigl[
  a(u;c)\,
  \mathrm{Ai}\!\bigl(c_3(c)^{-1/3}u\bigr)
  + r(u;c)\bigr],
\end{equation}
where
$a(u;c)\in S^{-1/2}_{1,0}(c)$,
$|r(u;c)|\le C_N(1+|u|)^{-N}$
for each fixed $N$,
and $R(u;c)$ absorbs the Maslov
phase factor
(see Remark~\ref{rem:FIO-language}
for the precise FIO sense).
\end{theorem}

\begin{proof}[Proof sketch]
Fold zone:
Proposition~\ref{prop:CFU}(2)
(conormal amplitude stable by
Lemma~\ref{lem:conormal-stability});
Lemma~\ref{lem:Jacobian-rescaling};
Lemma~\ref{lem:symbol-stability}.
Far zone:
Proposition~\ref{prop:CFU}(3).
\end{proof}

\begin{lemma}[Joint stationary-phase
  scaling in the fold zone \Rig]
\label{lem:product-estimate}
\textit{Valid in the fold zone
after canonical Airy rescaling;
not a global statement.}
Under Theorem~\ref{thm:airy-normalform},
for $|u|\ge\delta_2$:
\[
|a(u;c)\cdot\mathrm{Ai}(c_3(c)^{-1/3}u)|
\le C_1 C_2\,|u|^{-3/4}.
\]
This bound follows from
Lemma~\ref{lem:degenerate-phase}
applied to the model integral
with parameter $\mu=c_3(c)^{-1/3}u$:
the $-3/4$ exponent arises from
the combined scaling of the cubic
phase degeneracy and the conormal
amplitude, as established there.
The factorisation
$|a|\lesssim|u|^{-1/2}$,
$|\mathrm{Ai}|\lesssim|u|^{-1/4}$
is only an \emph{a posteriori}
estimate; the mechanism is a
\emph{single} degenerate stationary
phase expansion.
\end{lemma}

\begin{corollary}[Pointwise decay \Cond]
\label{cor:pointwise}
\textit{Conditional on
Assumption~\ref{ass:microlocal}(i)--(vi).}
There exists $C>0$,
uniform in $c\ge c_0$, such that:
\begin{equation}\label{eq:pointwise}
  |\Phi(u;c)|\le C(1+|u|)^{-3/4}
  \quad\forall u\in\mathbb{R}.
\end{equation}
\end{corollary}

\begin{proof}
Set $C:=\max(C_1C_2,\,2^{3/4}C_3,\,C_4)$:
\begin{itemize}
\item \textit{Fold zone
  $\delta_2\le|u|\le c^{\theta_0/3}$:}
  Lemma~\ref{lem:product-estimate}.
\item \textit{Compact zone $|u|\le 1$:}
  Continuity + (vi):
  $|\Phi|\le C_3\le
  2^{3/4}C_3(1+|u|)^{-3/4}$.
\item \textit{Far zone
  $|u|\ge c^{\theta_0/3}$:}
  Theorem~\ref{thm:endpoint}:
  $|\Phi|\le C_4|u|^{-1/2}
  \le C_4|u|^{-3/4}$ for $|u|\ge 1$.
\end{itemize}
\noindent\textit{Transition zone
$c^{\theta_0/3}\le|u|
\le 2c^{\theta_0/3}$.}
Both estimates hold simultaneously:
(a)~$C_1C_2$ is uniform in $c$
by (ii) and (vi);
(b)~$C_4$ via van~der~Corput
with $|\Theta''|\ge\delta_1|u|^{1/2}$
from (iii), uniform in $c$;
(c)~$|u|^{-3/4}\le|u|^{-1/2}$
in the overlap,
so the fold bound dominates.
Hence $C$ is uniform across all
zones and $c\ge c_0$.
\end{proof}

\begin{remark}[Integrability,
  not oscillatory cancellation]
\label{rem:oscillation}
The $O(U^{1/4})$ bound of
Corollary~\ref{cor:global-bound}
is a \emph{non-oscillatory mechanism}:
it uses only the absolute-value
bound $|\Phi|\le C(1+|u|)^{-3/4}$
and integration,
without any oscillatory cancellation
in $u$.
This is the integrability improvement
from the $-3/4$ exponent;
an oscillatory improvement would
require a second stationary-phase
step in $u$.
\end{remark}

\begin{corollary}[Ces\`aro improvement
  \Cond]
\label{cor:global-bound}
\textit{Conditional on
Assumption~\ref{ass:microlocal}(i)--(vi).}
\[
  \left|\int_0^U\Phi(u;c)\,du\right|
  \le C\int_0^U(1+u)^{-3/4}\,du
  = O(U^{1/4}),
\]
improving upon $O(U^{1/2})$
by the non-oscillatory mechanism
of Remark~\ref{rem:oscillation}.
\end{corollary}

\subsection{Converse and FIO rigidity}

\begin{assumption}[Microlocal
  reconstruction (non-standard) \Cond]
\label{hyp:FIO-rigidity}
Suppose $\Phi$ is in the Airy class and:
\begin{enumerate}[\rm(a)]
\item $\sigma(g\mapsto\Phi)
  \big|_\mathcal{C}\ne 0$;
\item $g\mapsto\Phi$ is injective
  modulo smoothing;
\item $\mathrm{WF}(\Phi)
  \subseteq\pi_u(\mathcal{C})$
  on the A$_2$ stratum.
\end{enumerate}
Then $\mathcal{C}$ contains an
A$_2$-fold singularity
uniformly in $c$.
\medskip

\noindent\textit{Status.}
This is an
\emph{inverse microlocal problem /
rigidity conjecture},
analogous in structure to
inverse spectral geometry
and inverse scattering,
and not a consequence of standard
FIO theory
\cite{Zworski2012,Hormander1983}.
Verification for PSWF: Paper~XVII.
\end{assumption}

\begin{corollary}[Conditional equivalence]
\label{cor:conditional-equiv}
Under Assumption~\ref{hyp:FIO-rigidity}:
Airy normal form $\Rightarrow$
Assumption~\ref{ass:microlocal}
(modulo the non-standard hypothesis).
The forward direction holds
rigorously without
Assumption~\ref{hyp:FIO-rigidity}.
\end{corollary}

\section{Residual Open Problem}
\label{sec:open}

\begin{problem}[A$_2$-stability
  for PSWF \Open]
\label{prob:PSWF-microlocal}
Does the PSWF amplitude
$g(\alpha;c)=|\alpha-\alpha_*|^{-1/2}
e^{i\theta(\alpha;c)}$
satisfy
Assumptions~\ref{ass:fold-local}
and~\ref{ass:microlocal}
uniformly in $c\to\infty$?
\begin{enumerate}[\rm(a)]
\item $c_3(c)\ge\delta_0>0$:
  this is expected from Olver's WKB
  analysis \cite{Olver1974},
  which gives asymptotic control
  of the PSWF phase to leading order;
  making it uniform in $c$
  requires quantitative WKB
  remainder bounds.
\item $\theta''(\alpha;c)\ne 0$
  for $\alpha\ne\alpha_*$, uniformly:
  plausible from WKB,
  but not yet established.
\item Conditions (iv) and (vi)
  for the PSWF CFU Jacobian:
  this is the technically hardest
  part, requiring explicit control
  of the CFU canonical transformation
  as a function of $c$;
  no general result is currently
  available.
\end{enumerate}
Outcomes:
(A)~All hold:
$|\int_0^U\Phi|=O(U^{1/4})$.
(B)~$c_3(c)\to 0$:
Ces\`aro sharp.
\end{problem}

\begin{thebibliography}{9}
\bibitem{PaperXV}
[Author], \textit{Paper~XV:
Intermediate-Regime Kernel Estimates},
2026.

\bibitem{Bleistein1986}
N.~Bleistein, R.~A.~Handelsman,
\textit{Asymptotic Expansions
of Integrals},
Dover, 1986
(orig.\ Holt, Rinehart and Winston,
1975).

\bibitem{Erdelyi1956}
A.~Erd\'elyi,
\textit{Asymptotic Expansions},
Dover, 1956.

\bibitem{Olver1974}
F.~W.~J.~Olver,
\textit{Asymptotics and Special
Functions},
Academic Press, 1974.

\bibitem{Zworski2012}
M.~Zworski,
\textit{Semiclassical Analysis},
AMS, 2012.

\bibitem{Hormander1983}
L.~H\"ormander,
\textit{The Analysis of Linear PDE~I},
Springer, 1983.

\bibitem{CFU1957}
C.~Chester, B.~Friedman, F.~Ursell,
\textit{An extension of the method
of steepest descents},
Proc.\ Cambridge Philos.\ Soc.\
\textbf{53} (1957), 599--611.

\bibitem{Arnold1972}
V.~I.~Arnol'd,
\textit{Normal forms for functions
near degenerate critical points},
Funct.\ Anal.\ Appl.\
\textbf{6} (1972), 254--272.

\bibitem{Golub1973}
M.~Golubitsky, V.~Guillemin,
\textit{Stable Mappings and
Their Singularities},
Springer, 1973.

\bibitem{GuillemU1981}
V.~Guillemin, G.~Uhlmann,
\textit{Oscillatory integrals
with singular symbols},
Duke Math.\ J.\ \textbf{48} (1981),
251--267.

\bibitem{MelroseU1979}
R.~Melrose, G.~Uhlmann,
\textit{Lagrangian intersection
and the Cauchy problem},
Comm.\ Pure Appl.\ Math.\
\textbf{32} (1979), 483--519.
\end{thebibliography}

\end{document}
